---
title: "Data Cleaning Part 2"
author: "Data Wrangling in R"
output:
  ioslides_presentation:
    css: ../styles/styles.css
    widescreen: yes
  beamer_presentation: default
---

```{r, include = FALSE}
library(knitr)
library(tidyverse)
library(janitor)
opts_chunk$set(comment = "")
```

# Data Cleaning Part 2




## Example of Cleaning: more complicated

For example, let's say we have a variable about treatment or control conditions coded as treatment, T, treat, Treat, C, Cont, cont, cOnt, Control, and control. Using Excel to find all of these would be a matter of filtering and changing all by hand or using if statements.

Sometimes though, it's not so simple.  That's where functions that find patterns come to be very useful.

<!-- ## UFO data -->

<!-- ```{r} -->
<!-- ufo <-read_csv("https://sisbid.github.io/Data-Wrangling/data/ufo/ufo_data_complete.csv", col_types = cols(`duration (seconds)` = "c")) -->

<!-- ufo <- clean_names(ufo) -->
<!-- ``` -->


<!-- ## UFO data is messy -->

<!-- ```{r} -->

<!-- ufo_short <- ufo %>% head(20) -->
<!-- ufo_short %>% pull(duration_hours_min) -->

<!-- ``` -->

## Some toy data for examples

```{r}
set.seed(4) # random sample below - make sure same every time
status <- sample(c("treatment", "T", "treat", 
                   "Traet", "Treat", "C", "Cont",
                   "cont", "cOnt", "Control", "control"),
                 1000, replace = TRUE)
data_case = tibble(status)
```

## Take a look at the data

```{r gentab}
count(data_case, status)
```

<!-- ## cleaning -->

<!-- ```{r} -->
<!-- #remove all instances of digits of any length or "-" or "/" or "." -->
<!-- ufo_short %>% mutate(unit = str_remove_all(duration_hours_min, "[0-9]+|-|/|\\.")) %>% select(duration_hours_min, unit) -->
<!-- ``` -->


## Example of Cleaning: more complicated
In R, you could use `case_when()`:

```{r}
#case_when way:
data_case <-data_case %>% mutate(status = 
            case_when(status 
          %in% c("C", "cont", "cOnt", "Cont", "control", "Control")
                                ~ "Control",
                           .default =  status)) 
count(data_case, status)
```

Oh dear! This only fixes some values! It is difficult to notice values like `"Traet"`.

# String functions

## The `stringr` package

Like `dplyr`, the `stringr` package:

* Makes some things more intuitive
* Is different than base R
* Is used on forums for answers
* Has a standard format for most functions: `str_`
    * the first argument is a string like first argument is a `data.frame` in `dplyr`


## Useful String Functions

Useful String functions from base R and `stringr`

* `toupper()`, `tolower()` - uppercase or lowercase your data
* `str_sentence()`  - uppercase just the first character (in the `stringr` package)

* `paste()` - paste strings together with a space 
* `paste0` - paste strings together with no space as default

* `str_trim()` (in the `stringr` package) or `trimws` in base 
    - will trim whitespace
* `nchar` - get the number of characters in a string

## recoding with `str_to_sentence()`

```{r}

data_case <-data_case %>%
                mutate(status = str_to_sentence(status))
count(data_case, status)
```


## recoding with `str_to_sentence()`

```{r}
#case_when way:
data_case <-data_case %>%
                mutate(status = str_to_sentence(status)) %>%
                mutate(status = 
                      case_when(status %in% 
                    c("Treatment", "T", "Treat", "Traet", "Treat")
                                ~ "Treatment",
                           .default = status)) 
count(data_case, status)
```

OK, now we are getting somewhere!


## Reading in again

Now we have a chance to keep but clean these values!

```{r}
ufo <-read_csv(
  "https://sisbid.github.io/Data-Wrangling/data/ufo/ufo_data_complete.csv", 
  col_types = cols(`duration (seconds)` = "c"))


```

## Clean names with the `clean_names()` function from the `janitor` package

```{r}
colnames(ufo)
ufo_clean <- clean_names(ufo)
colnames(ufo_clean)
```


    
## str_detect and filter{.codesmall}

Now let's fix our ufo data and remove those pesky backticks in the `duration_seconds` variable. First let's find them with `str_detect`.
```{r}
ufo_clean %>% 
  filter(str_detect( 
    string = duration_seconds, 
    pattern = "`"))

```

## str_remove

Now, let's remove the backticks.


```{r}
ufo_clean  <- ufo_clean %>% 
  mutate(duration_seconds = 
           str_remove(string = duration_seconds,
                     pattern = "`"))
```

And let's check it:

```{r}
ufo_clean %>% 
  filter(str_detect( 
    string = duration_seconds, 
    pattern = "`"))
```

## Lets also mutate to be numeric again {.codesmall}

There are many class functions to change from one class to another start with "as.".

```{r}
ufo_clean <- ufo_clean %>% 
  mutate(duration_seconds = as.numeric(duration_seconds))
ufo_clean 
```




## Substringing

`stringr`

* `str_sub(x, start, end)` - substrings from position start to position end

## Substringing

Examples: 

```{r}
str_sub("I like friesian horses", 8,12)
#123456789101112
#I like fries
str_sub(c("Site A", "Site B", "Site C"), 6,6)
```

## Splitting/Find/Replace and Regular Expressions

* R can do much more than find exact matches for a whole string
* Like Perl and other languages, it can use regular expressions.
* What are regular expressions?
    * Ways to search for specific strings 
    * Can be very complicated or simple
    * Highly Useful - think "Find" on steroids


## A bit on Regular Expressions

* https://rstudio.github.io/cheatsheets/strings.pdf 
* They can use to match a large number of strings in one statement
* `.` matches any single character
* `*` means repeat as many (even if 0) more times the last character
* `?` makes a pattern optional (i.e. it matches 0 or 1 times)
* `^` matches start of vector `^a` - starts with "a"
* `$` matches end of vector `b$` - ends with "b"

## 'Find' functions: `stringr`

`str_detect`, `str_subset`, `str_replace`, and `str_replace_all` search for matches to argument pattern within each element of a character vector: they differ in the format of and amount of detail in the results. 

* `str_detect` - returns `TRUE` if `pattern` is found
* `str_subset` - returns only the strings where the pattern were detected
* `str_extract` - returns only the pattern that was detected
* `str_replace` - replaces `pattern` with `replacement` the first time
* `str_replace_all` - replaces `pattern` with `replacement` as many times matched


<!-- ## Let's look at modifier for `stringr` -->

<!-- `?modifiers` -->

<!-- * `fixed` - match everything exactly -->
<!-- * `regexp` - default - uses **reg**ular **exp**ressions -->
<!-- * `ignore_case` is an option to not have to use `tolower` -->



## 'Find' functions: Finding Indices{.codesmall}

These are the indices where the pattern match occurs:

```{r alienMatch}
ufo_clean %>% 
  filter(str_detect(comments, "two aliens")) %>% 
  head()
```

## To Take a look at comments... need to select it first{.codesmall}

```{r}
ufo_clean %>%
  filter(str_detect(comments, "two aliens")) %>% 
  select(comments)
```



## 'Find' functions: `str_subset()` is easier {.smaller}

`str_subset()` gives the values that match the pattern: 
(or if we used negate = TRUE it would find the opposite)

```{r ggrep}
ufo_clean %>% pull(comments) %>%
  str_subset( "two aliens")

```

## Showing difference in `str_extract`

`str_extract` extracts just the matched string

```{r ggrep2}
ufo_clean %>% 
  mutate(aliens  = str_extract(comments, "two aliens")) %>%
  count(aliens)

```


* Look for any comment that starts with "aliens" 
  
```{r grepstar}
ufo_clean %>% pull(comments) %>% str_subset( "^aliens")
```

## Using Regular Expressions{.smaller}

That contains space then ship maybe with stuff in between
```{r grepstar2}
ufo_clean %>% pull(comments) %>%
  str_subset("space.?ship") %>% head(4) # gets "spaceship" or "space ship" or...
ufo_clean %>% pull(comments) %>% 
  str_subset("space.ship") %>% head(4) # no "spaceship" must have character in between

```

## time information

```{r}

pull(ufo_clean, duration_hours_min) %>% head(n = 20)
```

## str_replace()


Let's say we wanted to make the time information more consistent. 
Using `case_when()` could be very tedious and error-prone!

We can use `str_replace()` to do so.

```{r classSal}
ufo_clean %>% mutate(duration_hours_min = 
                 str_replace(string = duration_hours_min, 
                             pattern = "minutes", 
                             replacement ="mins")) %>%
  pull(duration_hours_min) %>%
  head(8)

```

## Separating columns

Better yet, you might notice that this data isn't tidy- there are more than two entries for each value - amount of time and unit. We could separate this using `separate()` from the `tidyr` package.

```{r, warning=FALSE }
ufo_clean %>% separate(duration_hours_min,
                 into = c("duration_amount", "duration_unit"),
                 sep = " ") %>%
  select(duration_amount, duration_unit) %>% head()
```
As you can see there is still plenty of cleaning to do!


## more seperating

```{r, warning=FALSE }
ufo_clean <- ufo_clean %>% separate(datetime,
                 into = c("date", "time"),
                 sep = " ")
ufo_clean %>% select(date, time) %>% head()
```

## Dates and times{.smaller}

The [`lubridate`](https://lubridate.tidyverse.org/) package is amazing for dates.
Most important functions are those that look like ymd or mdy etc. They specify how a date should be interpreted.

```{r, message = FALSE}
library(lubridate)#need to load this one!

ufo_clean <- ufo_clean %>% mutate(date = mdy(date))
head(ufo_clean)
```


## str_*functions

```{r}
str_detect(string = c("abcdd", "two"), pattern = "dd")
str_subset(string = c("abcdd", "two"), pattern = "dd")
str_extract(string = c("abcdd", "two"), pattern = "dd")
str_sub(string = c("abcdd", "two"), start = 1, end = 3)
```

## Summary 

- `stringr` package has lots of helpful functions that work on vectors or variables in a data frame
- `str_detect` helps find patterns
- `str_detect` and `filter` can help you filter data based on patterns within value
- `str_extract` helps extract a pattern
- `str_sub` extracts pieces of strings based on the position of the the characters
- `str_subset` gives the values that match a pattern
- `separate` can separate columns into two
- `^` indicates the start of a string
- `$` indicates the end of a string
- the `lubridate` package is useful for dates and times

## Lab

https://sisbid.github.io/Data-Wrangling/09_Data_Cleaning/lab/data-cleaning-lab-part2.Rmd
