---
title: "R - Big Data Tricks"
author: "Andrew Jaffe"
date: "July 13, 2016"
output:
  ioslides_presentation:
    css: styles.css
  beamer_presentation: default
---

```{r, echo = FALSE}
library(knitr)
opts_chunk$set(comment = "")
```

## Overview

There are several different tricks and approaches for handling "big" datasets in R. Most of these tricks either improve the importing of large datasets and/or speed up tasks applied to large datasets.

## R is memory intensive

R is pretty much limited by the amount of memory on your computer (or cluster), and potentially the number of cores on your computer (or cluster).

# Getting around memory limits

## Using Disk

The package `ff` can "trick" R into using disk space as virtual memory instead of relying on physical memory

https://cran.r-project.org/web/packages/ff/index.html

## Using Disk

There are `read.csv` and `read.table` functions for leveraging these `ff` objects.

These functions are `read.csv.ffdf` and `read.table.ffdf` respectively.

```{r}
suppressPackageStartupMessages(library(ff))
mon_ff= read.csv.ffdf(file="../data/Monuments.csv")
class(mon_ff)
```

## Using Disk

New columns must be converted to the `ff` class: 

```{r}
mon_ff$newCol = ff(1:nrow(mon_ff))
mon_ff[1:4,1:7]
```

## Using Disk 

More info, with examples, can be found here:

* http://hsinay.blogspot.com/p/big-data-analysis-using-ff-and-12-nov.html
* http://ff.r-forge.r-project.org/bit&ff2.1-2_WU_Vienna2010.pdf
* https://www.r-project.org/conferences/useR-2007/program/presentations/adler.pdf

## Leveraging sparsity

If your data is very sparse, e.g. contains a lot of 0s, then the `sparseMatrix` command in the `Matrix` package might be helpful

https://stat.ethz.ch/R-manual/R-devel/library/Matrix/html/sparseMatrix.html

You need to provide which rows and columns have non-zero values (and the actual values themselves)

## Leveraging sparsity

We've had to do this for genomics projects:

* http://biorxiv.org/content/early/2016/01/29/038224 with corresponding code:
* https://github.com/nellore/runs/blob/master/sra/v2/countJunctions.R

## Leveraging sparsity

Run length encoding (RLE) is a popular approach for handling sparse vectors...the`DataFrame()` function in the `IRanges` package creates a more flexible extension of the `data.frame` class

```{r, rleExample}
x= rep(c(0,1,0,1,0), times=c(10000,4,10000,6,10000))
length(x)
rle(x) # base r
```

## Leveraging sparsity

```{r, rleExample2}
suppressPackageStartupMessages(library(IRanges))
df= data.frame(x=x)
DF = DataFrame(df)
Rle(x) # from IRanges, s4-methods
DF$x = Rle(DF$x) # note the capital R
head(DF)
```
## Leveraging sparsity

How much space did we save in memory?

```{r, sparseSize}
s1=object.size(df)
s1
s2=object.size(DF)
s2
as.numeric(round(s1/s2))
```
You can see this will can add up for large datasets. Note this doesn't rely on sparity per se, but rather many repeated fields that cluster together.

## Leveraging Sparsity

We've also used these file encoding for genomics projects:

* http://biorxiv.org/content/early/2016/05/19/015370 with package:
* https://www.bioconductor.org/packages/release/bioc/html/derfinder.html

## Using Databases

Jeff covered this yesterday. If you can store the really large dataset in a SQL database, then it's easy to create your own SQL queries to read in only a subset of the dataset. 

# Speeding up computing

## Parallel libraries

Almost all R functions use only a single core on your computer but you can leverage more cores. You can usually speed up code by using multiple cores on your computer. 

* [doParallel](https://cran.r-project.org/web/packages/doParallel/vignettes/gettingstartedParallel.pdf)
* [Parallel Basics](http://www.r-bloggers.com/how-to-go-parallel-in-r-basics-tips/)
* [mclapply](http://www.inside-r.org/packages/cran/multicore/docs/mclapply)
* [HPC](https://cran.r-project.org/web/views/HighPerformanceComputing.html)

## Parallel libraries

You can parallelize `lapply()` (or `sapply()`) statements with the `parallel::mclapply()` function. 

You can parallel `for` loops with the `foreach()` with `%dopar%` syntax. [`foreach` with `%do%` are analogous to `for` loops]

## mclapply

Let's make some data

```{r}
set.seed(34)
x = rnorm(10000)
datList = replicate(10000, sample(1:10000),simplify = FALSE)
class(datList)
lapply(datList[1:5], head)
head(x)
```

## mclapply 

Let's try to run `mclapply()` and compare to the output of `lapply()`

```{r}
system.time(lapply(datList,cor,x))
system.time(mclapply(datList, cor, x))
```

???

## mclapply

This only works on Linux/Mac, and not Windows `:(`

```
> system.time(mclapply(datList,function(y) cor(x,y),mc.cores=4))
Error in mclapply(datList, sort, mc.cores = 4) : 
  'mc.cores' > 1 is not supported on Windows
```

## doParallel

The `doParallel` library works better on Windows:

```{r}
suppressPackageStartupMessages(library(doParallel))
system.time(foreach(i=1:10000) %do% cor(datList[[i]],x))
system.time(foreach(i=1:10000) %dopar% cor(datList[[i]],x))
```

Oops! You need to initiate a parallel backend first!

## doParallel

Implementing parallel backend:

```{r}
registerDoParallel(cores=2)
system.time(foreach(i=1:10000) %dopar% cor(datList[[i]], x))
```

This task was not particularly amenable to parallelization, but you get the idea of the code.

This is a good quick-start guide: http://blog.aicry.com/r-parallel-computing-in-5-minutes/

## Vectorization

Some tasks can be computed quickly just using vector or matrix-based calculations instead of iterating:

```{r}
datMat = simplify2array(datList)
class(datMat)
dim(datMat)
system.time(cor(x,datMat))
```

## Rcpp

The `Rcpp` package allows you to write C++ code and run it within R.

http://gallery.rcpp.org/
http://dirk.eddelbuettel.com/code/rcpp/Rcpp-introduction.pdf
https://cran.r-project.org/web/packages/Rcpp/index.html

[Note: I am not very good at C++, and it usually takes me a lot less time to just let code run longer than figure out how to do something in C]

## GPU computing

There are packages that can rely on "graphics processing units" aka video cards that can greatly speed up linear computations

http://www.r-tutor.com/gpu-computing
http://www.r-tutor.com/gpu-computing/clustering/distance-matrix

[Note: I have never actually used this, and it usually takes me a lot less time to just let code run longer than figure out how to do something using the GPU]

# Modifying data prior to importing

## Slicing and dicing large data

Say you have a large tab-delimited matrix that has dimensions: 100,000,000 (rows) x 500 (columns) called `foo.txt`

The easiest tricks are for Mac or Linux computers since you can utilize very fast linux file processing on text files prior to importing data into R

## (Linux) Piping 

R can read in the output of Linux commands using the `pipe()` function.

```
dat = read.delim(pipe([some command]))
```

The most useful two linux functions are `awk` (for selecting rows) and `cut` (for selecting columns)

## awk

Here are a collection of useful "one liners" using `awk`

http://www.pement.org/awk/awk1line.txt

Note, I almost always have to google how to use `awk` to remember the exact syntax for specific tasks

## awk

Let's say we want to select all rows where the second column has the values "chr21"

The awk command looks like:

`awk -F"\t" '$2 == "chr21" { print $0 }' foo.txt`

Note in linux you would probably redirect this to a file, but R can handle this "on the fly"

## awk {.smaller}

```{r, awkImport,eval=FALSE}
> awkCall = "awk -F\"\t\" '$2 == \"chr21\" { print $0 }' foo.txt"
> read.delim(pipe(awkCall),header=FALSE)
     V1    V2         V3        V4        V5        V6        V7         V8
1  row6 chr21 0.06470508 0.2104643 0.2630865 0.6748644 0.5838493 0.59531516
2  row7 chr21 0.69930753 0.9533248 0.7929459 0.7084059 0.5305247 0.37651206
3  row8 chr21 0.31752770 0.6364067 0.7774075 0.6607039 0.5074746 0.43583281
4  row9 chr21 0.18445757 0.1926704 0.5522359 0.4889265 0.2885484 0.36124031
5 row10 chr21 0.96333352 0.9455556 0.2251288 0.7133460 0.2673244 0.01864215
         V9       V10        V11
1 0.5832685 0.8670779 0.08756482
2 0.9471309 0.6032699 0.37931679
3 0.2296118 0.4037352 0.37954743
4 0.3595676 0.6895200 0.09194656
5 0.1601413 0.4742028 0.07500326
```

## awk {.smaller}

Maybe we don't want all of the columns

```{r, awkImport2,eval=FALSE}
> awkCall2 = "awk -F\"\t\" '$2 == \"chr21\" { print $2\"\t\"$3\"\t\"$4 }' foo.txt"
> read.delim(pipe(awkCall2),header=FALSE)
     V1         V2        V3
1 chr21 0.06470508 0.2104643
2 chr21 0.69930753 0.9533248
3 chr21 0.31752770 0.6364067
4 chr21 0.18445757 0.1926704
5 chr21 0.96333352 0.9455556
```

## awk

Note you have to "escape" all of the quotation marks that are within the actual awk call so that they are not treated by quotation marks in R


## cut

`cut` selects certain columns 

Let's take columns 2, 3, and 4 for all rows:

```{r, cutImport,eval=FALSE}
> cutCall = "cut -f2,3,4 foo.txt"
> read.delim(pipe(cutCall),header=FALSE)
      V1         V2         V3
1   chr1 0.55127556 0.06161779
2   chr1 0.97378813 0.84200893
3   chr1 0.67479999 0.54580091
4   chr1 0.90762683 0.62246401
5   chr1 0.27251173 0.09567016
6  chr21 0.06470508 0.21046433
7  chr21 0.69930753 0.95332484
8  chr21 0.31752770 0.63640675
9  chr21 0.18445757 0.19267041
10 chr21 0.96333352 0.94555563
```

## cut

The `-d` option specifies that string that separates columns

## Combining awk + cut {.smaller}

You can also use the linux pipe "|" within the R `pipe()` call:

```{r, bothRun, eval=FALSE}
> bothCall = paste("awk -F\"\t\" '$2 == \"chr21\" { print $0 }' foo.txt | cut -f2,3,4")
> read.delim(pipe(bothCall),header=FALSE)
     V1         V2        V3
1 chr21 0.06470508 0.2104643
2 chr21 0.69930753 0.9533248
3 chr21 0.31752770 0.6364067
4 chr21 0.18445757 0.1926704
5 chr21 0.96333352 0.9455556
```

## Windows users

Doing all of this in Windows takes a lot more setup, and there are a more limited set of Linux commands:
http://stackoverflow.com/questions/23963106/read-columns-of-a-csv-file-using-shell-or-pipe-inside-r-windows
