---
title: "Data I/O + Structure"
author: "Data Wrangling in R"
output:
  ioslides_presentation:
    css: styles.css
    widescreen: yes
  beamer_presentation: default
---

```{r, include = FALSE}
library(knitr)
library(tidyverse)
library(visdat)
# suppressPackageStartupMessages(library(dplyr))
knitr::opts_chunk$set(comment = "")
ufo <-read_csv("../data/ufo/ufo_data_complete.csv")
```

# Data Input

## Outline 

* Part 0: A little bit of set up!
* Part 1: reading in manually (point and click)
* Part 2: reading in directly & working directories
* Part 3: checking data & multiple file formats


## Data Input: `readr`

`read_delim()` and `read_csv()` from the`readr` package

```{r eval=FALSE}
# example:
read_delim(file = "file.txt", delim = "\t")
read_csv("file.csv")
```


## Data Input

* The filename is the path to your file, in quotes
* The function will look in your "working directory" if no absolute file path is given
* Note that the filename can also be a path to a file on a website (e.g. 'www.someurl.com/table1.txt')


## Example

https://sisbid.github.io/Data-Wrangling/data/

```{r eval=FALSE}
# From URL
ufo <- read_csv("https://sisbid.github.io/Data-Wrangling/data/ufo/ufo_data_complete.csv")
# From your 'data-wrangling' directory
ufo <- read_csv("ufo_data_complete.csv")
```


## Data Input

The `read_delim()` and related functions return a "tibble" is a `data.frame` with special printing, which is the primary data format for most data cleaning and analyses.

```{r viewInput, message=FALSE}
class(ufo)
```

Check to make sure you see the new object in the Environment pane.


## Data Input

There are also data importing functions provided in base R (rather than the `readr` package), like `read.delim` and `read.csv`.

These functions have slightly different syntax for reading in data, like `header` and `as.is`.

However, while many online resources use the base R tools, recent versions of RStudio switched to use these new `readr` data import tools, so we will use them here. They are also up to two times faster for reading in large datasets, and have a progress bar which is nice.


## Data Input: `readr`

`read_table()` from the`readr` package, allows any number of whitespace characters between columns, and the lines can be of different lengths.

```{r eval=FALSE}
# example:
read_table(file = "file.txt")
```


## Clean the data while you read it in!

Some data have different values for `NA`. Maybe we need to correct the data   

```{r}
vacc <- 
  read_csv("https://sisbid.github.io/Data-Wrangling/data/vaccinations_1.csv")
vacc_na <- 
  read_csv("https://sisbid.github.io/Data-Wrangling/data/vaccinations_1.csv",
           na = '"NaN"')
vacc[1:3,1:4]
vacc_na[1:3,1:4]
```


## Clean the data while you read it in!

The argument `trim_ws` removes trailing and leading spaces around your data.

```{r eval=FALSE}
# example:
read_csv(file = "file.txt", trim_ws = TRUE)
```


## Data Input - working directories

What if your file is in the "Home" directory?

```{r, out.width="80%", echo=FALSE, fig.alt='.'}
ottrpal::include_slide("https://docs.google.com/presentation/d/104LQkFTsC5R9vAC4HHj4mK7IjCTH55X6BkrAb3aDd6E/edit#slide=id.g13ea04b780d_1_121")
```

## Data Input

Backtrack using the relative path with `../` like:

```{r, eval = FALSE}
ufo <- read_csv("../ufo_data_complete.csv.gz")
```

## Data Input

Or, read in from a subfolder:

```{r, eval=FALSE}
ufo <- read_csv("data/ufo/ufo_data_complete.csv")
```

```{r, echo=FALSE}
ufo <- read_csv("../data/ufo/ufo_data_complete.csv")
```


# Check the data + other formats

## Check the data out

- Some functions to look at a data frame:
  - `head()` shows first few rows
  - `tail()` shows the last few rows
  - `View()` shows the data as a spreadsheet
  - `spec()` gives specification of column types
  - `str()` gives the column types and specs
  - `glimpse()` similar to `str` (dplyr package)


## What did I just read in?

* `nrow()` displays the number of rows of a data frame
* `ncol()` displays the number of columns 
* `dim()` displays a vector of length 2: # rows, # columns

```{r, dims}
nrow(ufo)
ncol(ufo)
dim(ufo)
```


## All Column Names

* `colnames()` displays the column names 

```{r, colnames}
colnames(ufo)
```


## Structure using `str()`

```{r, str}
str(ufo)
```


## Data Input

* Sometimes you get weird messages when reading in data.
* The problems()` function shows you any issues with the data read-in.

```{r stop_problems}
head(problems(ufo))
```

```{r}
dim(problems(ufo))
```


## Data input: other file types

* For reading Excel files, you can do one of:
  - use `read_excel()` function from `readxl` package
  - use other packages: `xlsx`, `openxlsx`

* `haven` package has functions to read SAS, SPSS, Stata formats


## Selecting Excel sheets

Use the `sheet` argument to indicate which sheet to pull from. It can refer to the sheet's index or name.

```{r eval=FALSE}
# example:
read_excel(file = "file.xlsx", sheet = 2)
read_excel(file = "file.xlsx", sheet = "data")
```


# After hours of cleaning... output!

## Data Output

While its nice to be able to read in a variety of data formats, it's equally important to be able to output data somewhere.

`write_delim()`: Write a data frame to a delimited file
`write_csv()`:  Write a data frame to a comma-delimited file

**This is about twice as fast as `write.csv()`, and never writes row names.**


## Data Output

For example, we can write back out just the first 100 lines of the `ufo` dataset:

```{r writecsv}
first_100 <- ufo[1:100,]
write_delim(first_100, file = "ufo_first100.csv", delim = ",")
write_csv(first_100, file = "ufo_first100.csv")
```


## More ways to save: `write_rds`

If you want to save **one** object, you can use `readr::write_rds` to save to a compressed `rds` file:

```{r eval=FALSE}
write_rds(ufo, file = "ufo_dataset.rds", compress = "xz")
```

Read it back in: 

```{r eval=FALSE}
ufo_new <- read_rds(file = "ufo_dataset.rds")
```


## More ways to save: `save`

The `save` command can save a set of `R` objects into an "R data file", with the extension `.rda` or `.RData`.  

```{r, message = FALSE}
x = 5
save(ufo, x, file = "ufo_data.rda")
```

The opposite of `save` is `load`.

```{r, eval = FALSE}
load(file = "ufo_data.rda")
```


## Summary & Lab

- Use `read_delim()`, `read_csv()`, `read_table()` for common data types
- These have helpful `trim_ws` and `na` arguments!
- `read_excel()` has the `sheet` argument for reading from different sheets of the Excel file
- Many functions like `str()`, `View()`, and `glimpse()` can help you understand your data better
- Save your data with `write_delim()` and `write_csv()`

https://sisbid.github.io/Data-Wrangling/labs/data-io-lab-part2.Rmd
