---
title: "Subsetting Data in R"
author: "Data Wrangling in R"
output:
  beamer_presentation: default
  ioslides_presentation:
    css: styles.css
    widescreen: yes
---

```{r, include=FALSE}
knitr::opts_chunk$set(comment = "")
library(tidyverse)
```

# Dealing with Missing Data

## Missing data types
One of the most important aspects of data cleaning is missing values.  

Types of "missing" data:

* `NA` - general missing data
* `NaN` - stands for "**N**ot **a** **N**umber", happens when you do 0/0.
* `Inf` and `-Inf` - Infinity, happens when you take a positive number (or negative number) by 0.

## Finding Missing data

Each missing data type has a function that returns `TRUE` if the data is missing:

* `NA` - `is.na()`
* `NaN` - `is.nan()`
* `Inf` and `-Inf` - `is.infinite()`
* `is.finite()` returns `FALSE` for all missing data and `TRUE` for non-missing

## Working with `is.na()`

```{r}
t <- c(1,0,3,0)
x <-c(0,0,3,NA)
t / x
```
```{r}
is.na(t/x)
```

## Missing Data with Logicals

One important aspect (esp with subsetting) is that logical operations return `NA` for `NA` values.  The data could be `> 2` or not. We don't know, 
so `R` says there is no `TRUE` or `FALSE`, and instead that is missing:
```{r}
x <- c(0, NA, 2, 3, 4)
x > 2
```

## Missing Data with Logicals

What to do?  What if we want if `x > 2` and `x` isn't `NA`?  

```{r}
x
!is.na(x)
x > 2 & !is.na(x)
```

<!-- ## Missing Data with Logicals: `%in%` -->

<!-- What about seeing if a value is equal to multiple values?   -->

<!-- You can do `(x == 1 | x == 2) & !is.na(x)`, but that is not efficient.   -->

<!-- ```{r} -->
<!-- x -->
<!-- (x == 0 | x == 2) & !is.na(x) # No NA -->
<!-- x %in% c(0, 2) # NEVER has NA and returns logical -->
<!-- ``` -->

<!-- reads "return `TRUE` if `x` is in 0 or 2". (Like `inlist` in Stata). -->


## Missing Data with Operations

Similarly with logicals, operations/arithmetic with `NA` will result in `NA`s:

```{r}
x + 2
x * 2
```
 

## UFO data again

```{r, eval = FALSE}
ufo = read_csv("../data/ufo/ufo_data_complete.csv", col_types = 
                 cols(
                   .default = col_character(),
                   `duration (seconds)` = col_double(),
                   longitude = col_double()
                 ))
head(ufo)
```

```{r, echo = FALSE, warning = FALSE}
ufo = read_csv("../data/ufo/ufo_data_complete.csv", col_types = 
                 cols(
                   .default = col_character(),
                   `duration (seconds)` = col_double(),
                   longitude = col_double()
                 ))
head(ufo)
```


## Filtering and tibbles

The filter function automatically removes `NA` values.

```{r}
count(ufo, country)
ufo %>%filter(country == "de") %>% count(country)
```

## Filtering for missing data

Missing value and filter can be powerful 

```{r}
ufo %>% 
  filter(is.na(state) & is.na(country)) %>% 
  head()
```

## Filtering and tibbles

Group logical statements with parentheses

```{r}
ufo %>% 
  filter( 
    (!is.na(state) & is.na(country)) | city == "seattle") %>% 
  head()
```

# Renaming Columns

## Renaming Columns of a `data.frame`

To rename columns in `dplyr`, you use the `rename` command (NEW=old)

```{r}
ufo = ufo %>% rename(City = city, duration_s = `duration (seconds)`)
glimpse(ufo)
```

## Renaming All Columns of a `data.frame`: dplyr

To rename all columns you use the `rename_with` command (with a function)

```{r}
ufo_upper = ufo %>% rename_with(toupper)
glimpse(ufo_upper)
```

## Adding columns to a `data.frame`

 `mutate` - allows you to add or replace columns (IMPORTANT: need to reassign for it to stick!)
 
```{r}
ufo2 <- ufo %>% mutate(State = toupper(state))
ufo2 %>% glimpse()
```


<!-- ## Adding new columns to a `data.frame`: base R -->

<!-- Can also use `$` to add columns, but only one column at a time -->
<!-- ```{r} -->
<!-- ufo2$State2 = tolower(ufo2$State) -->
<!-- ufo2 %>% select(state, State, State2) %>% head -->
<!-- ``` -->



## Creating conditional variables

One frequently-used tool is creating variables with conditions.

A general function for creating new variables based on existing variables is the `ifelse()` function, which "returns a value with the same shape as test which is filled with elements selected from either yes or no depending on whether the element of test is `TRUE` or `FALSE`."

```
ifelse(test, yes, no)

# test: an object which can be coerced
    to logical mode.
# yes: return values for true elements of test.
# no: return values for false elements of test.
```

## Recoding to missing

Sometimes people code missing data in weird or inconsistent ways.

```{r codeNA}
ages <-  data.frame(age = c(23,-999,21,44,32,57,65,54))
range(ages$age)
```


## Recoding to create new column
Say we want to make a new column about if the age was over 30?

```{r}
pull(ages, age)
ages <- ages %>% mutate(over_20 = ifelse(age > 30, "Yes", "No"))
ages
```



## Recoding value to missing

How do we change the `-999` to be treated as missing for the age column?

```{r codeNA2}
pull(ages, age)
ages <- ages %>% mutate(age = ifelse(age == -999, NA, age))
range(ages$age)
range(ages$age, na.rm=TRUE)
pull(ages, age)
```

<!-- ## Recoding from missing -->

<!-- What if you were the person that coded the `-999` -->

<!-- ```{r codeFromNA} -->
<!-- ages = ages %>% mutate(age = ifelse(is.na(age), -999, age)) -->
<!-- ages -->
<!-- ``` -->

## Adding columns to a `data.frame`: dplyr

```{r}
ufo <- ufo %>% mutate( 
              region = ifelse(
              country %in% c("us", "ca"),
              "North America",
              "Not North America")
            )
ufo %>% select(country, region) %>% head()
```

## `case_when` provides a more general way

```
casewhen(test ~ value if test is true,
         test2 ~ vlue if test2 is true,
         TRUE ~ value if all above tests are not true) # defaults to NA
```
```{r}
ufo <- ufo %>% mutate( 
            region = case_when(
              country %in% c("us", "ca") ~ "North America",
              country %in% c("de") ~ "Europe",
              country %in% "gb" ~ "Great Britain",
            ))

ufo %>% select(country, region) %>% head()
```
## `case_when` defaults to NA when all tests fail

```{r}
ufo <- ufo %>% mutate( 
            region = case_when(
              country %in% c("us", "ca") ~ "North America",
              country %in% c("de") ~ "Europe"
              #country %in% "gb" ~ "Great Britain",
            ))

ufo %>% select(country, region) %>% head()
```


## `case_when()` with value if all tests fail

```{r}
ufo <- ufo %>% mutate( 
            region = case_when(
              country %in% c("us", "ca") ~ "North America",
              country %in% c("de") ~ "Europe",
              country %in% "gb" ~ "Great Britain",
              TRUE ~ "Other"
            ))
ufo %>% select(country, region) %>% head()
```




## `case_when()` with value if all tests fail (use a variable!)

```{r}
ufo <- ufo %>% mutate( 
            region = case_when(
              country %in% "gb" ~ "Great Brit.",
              TRUE ~ region
            ))
ufo %>% select(country, region) %>% head()
```




## Ordering the rows of a `data.frame`: dplyr

The `arrange` function can reorder rows  By default, `arrange` orders in ascending order:
```{r}
ufo %>% arrange(duration_s)
```

## Ordering the rows of a `data.frame`: dplyr
Use the `desc` to arrange the rows in descending order:
```{r}
ufo %>% arrange(desc(duration_s))
```

## Ordering the rows of a `data.frame`: dplyr
It is a bit more straightforward to mix increasing and decreasing orderings:
```{r}
ufo %>% arrange(country, desc(duration_s))
```

## Lab

[Link to Lab](http://sisbid.github.io/Data-Wrangling/labs/data-subsetting-lab-part2.Rmd)