---
title: "Data Cleaning Part 2"
author: "Data Wrangling in R"
output:
  ioslides_presentation:
    css: styles.css
    widescreen: yes
---

```{r, include = FALSE}
library(knitr)
library(tidyverse)
library(janitor)
opts_chunk$set(comment = "")
```

# Data Cleaning Part 2


## Reading in again

Now we have a chance to keep but clean these values!

```{r}
ufo <- read_delim("../data/ufo/ufo_data_complete.txt", 
                  col_types = cols("duration (seconds)" = col_character()))

ufo %>% glimpse()

```

## Clean names with the `clean_names()` function from the `janitor` package

```{r}
colnames(ufo)
ufo = clean_names(ufo)
colnames(ufo)
```


## The `stringr` package

Like `dplyr`, the `stringr` package:

* Makes some things more intuitive
* Is different than base R
* Is used on forums for answers
* Has a standard format for most functions: `str_`
    * the first argument is a string like first argument is a `data.frame` in `dplyr`
    
    
## str_remove

Now let's fix our ufo data and remove those pesky backticks in the `duration_seconds` variable.

```{r}
ufo <- ufo %>% mutate(duration_seconds = str_remove(string = duration_seconds,
                                             pattern = "`"))
ufo <- ufo %>% mutate(duration_seconds = as.numeric(duration_seconds))
ufo[(ufo$row-1),]
```

## Paste can add things back to variables

```{r}
head(Orange)
Orange %>% mutate(Tree = paste(Tree, "Tree", sep = "_"))

```
## Paste0 doesn't need a separator

```{r}
head(Orange)
Orange %>% mutate(Tree = paste0(Tree, "Tree"))

```

## Substringing

`stringr`

* `str_sub(x, start, end)` - substrings from position start to position end
* `str_split(string, pattern)` - splits strings up - returns list! [we'll revisit in "Functional Programming"]

## Substringing

Examples: 

```{r}
str_sub("I like friesian horses", 8,12)
#123456789101112
#I like fries
str_sub(c("Site A", "Site B", "Site C"), 6,6)
```

## Splitting/Find/Replace and Regular Expressions

* R can do much more than find exact matches for a whole string
* Like Perl and other languages, it can use regular expressions.
* What are regular expressions?
    * Ways to search for specific strings 
    * Can be very complicated or simple
    * Highly Useful - think "Find" on steroids


## A bit on Regular Expressions

* http://www.regular-expressions.info/reference.html
* They can use to match a large number of strings in one statement
* `.` matches any single character
* `*` means repeat as many (even if 0) more times the last character
* `?` makes the last thing optional
* `^` matches start of vector `^a` - starts with "a"
* `$` matches end of vector `b$` - ends with "b"

## 'Find' functions: `stringr`

`str_detect`, `str_subset`, `str_replace`, and `str_replace_all` search for matches to argument pattern within each element of a character vector: they differ in the format of and amount of detail in the results. 

* `str_detect` - returns `TRUE` if `pattern` is found
* `str_subset` - returns only the strings which pattern were detected
    * convenient wrapper around `x[str_detect(x, pattern)]`
* `str_extract` - returns only strings which pattern were detected, but ONLY the pattern
* `str_replace` - replaces `pattern` with `replacement` the first time
* `str_replace_all` - replaces `pattern` with `replacement` as many times matched


<!-- ## Let's look at modifier for `stringr` -->

<!-- `?modifiers` -->

<!-- * `fixed` - match everything exactly -->
<!-- * `regexp` - default - uses **reg**ular **exp**ressions -->
<!-- * `ignore_case` is an option to not have to use `tolower` -->



## 'Find' functions: Finding Indices

These are the indices where the pattern match occurs:

```{r alienMatch}
str_detect(ufo$comments, "two aliens") %>% head()
str_detect(ufo$comments, "two aliens") %>% table()
which(str_detect(ufo$comments, "two aliens"))
```

## 'Find' functions: Finding Logicals {.smaller}

`filter()` using `str_detect()` gives a tibble:

```{r}
filter(ufo, str_detect(comments, "two aliens"))
filter(ufo, str_detect(comments, "two aliens")) %>% select(comments)

```

## 'Find' functions: `str_subset()` is easier {.smaller}

`str_subset()` gives the values that match the pattern:

```{r ggrep}
str_subset(ufo$comments, "two aliens")

```

## Showing difference in `str_extract`

`str_extract` extracts just the matched string

```{r ggrep2}
ss = str_extract(ufo$comments, "two aliens")
head(ss)
ss[ !is.na(ss)]
```


* Look for any comment that starts with "aliens"
  
```{r grepstar}
str_subset(ufo$comments, "^aliens.*")
```

## Using Regular Expressions{.smaller}

That contains space then ship maybe with stuff in between
```{r grepstar2}
str_subset(ufo$comments, "space.?ship") %>% head(4) # gets "spaceship" or "space ship" or...
str_subset(ufo$comments, "space.ship") %>% head(4) # no "spaceship" must have character in bw

```

## str_replace()


Let's say we wanted to make the time information more consistent. 
Using `case_when()` would be very tedious and error-prone!

We can use `str_replace()` to do so.

```{r classSal}
head(ufo$duration_hours_min, 8)

ufo %>% mutate(duration_hours_min = 
                 str_replace(string = duration_hours_min, 
                             pattern = "minutes", 
                             replacement ="mins")) %>%
  pull(duration_hours_min) %>%
  head(8)

```

<!-- ## Separating columns -->

<!-- Better yet, you might notice that this data isn't tidy- there are more than two entries for each value - amount of time and unit. We could separate this using `separate()` from the `tidyr` package. -->
<!-- ```{r, warning=FALSE } -->
<!-- ufo %>% separate(duration_hours_min,  -->
<!--                  into = c("duration_amount", "duration_unit"), -->
<!--                  sep = " ") %>% -->
<!--   select(duration_amount, duration_unit) %>% head() -->
<!-- ``` -->
<!-- As you can see there is still plenty of cleaning to do! -->

## Dates and times{.smaller}

The `[lubridate`](https://lubridate.tidyverse.org/) package is amazing, there's no reason to use anything else.

```{r, message = FALSE}
library(lubridate)#need to load this one!
head(ufo$datetime)
ufo$date_posted = mdy(ufo$date_posted)
head(ufo$date_posted)
```

<!-- ## Special characters -->

<!-- ```{r} -->
<!-- money = tibble(group = letters[1:5],  -->
<!--   amount = c("$12.32", "$43.64", "$765.43", "$93.31", "$12.13")) -->
<!-- money %>% arrange(amount) -->
<!-- as.numeric(money$amount) -->
<!-- ``` -->

<!-- ## Special characters -->

<!-- In the past, we would recommend just replacing the `$` sign with an empty string and convert to numeric: -->

<!-- ```{r}  -->
<!-- money$amountNum = as.numeric(str_replace(money$amount, fixed("$"), "")) -->
<!-- money %>% arrange(amountNum) -->
<!-- ``` -->

<!-- ## Special characters -->

<!-- But now there are better helper functions for this: -->

<!-- ```{r} -->
<!-- money$amount = parse_number(money$amount) -->
<!-- money %>% arrange(amount) -->

<!-- ``` -->

<!-- ## Special characters -->

<!-- Also works for internal commas: -->

<!-- ```{r} -->
<!-- parse_number(c("12,123,123.00", "12,465.10")) -->
<!-- ``` -->




