---
title: "Data I/O + Structure"
author: "Data Wrangling in R"
output:
  ioslides_presentation:
    css: styles.css
    widescreen: yes
  beamer_presentation: default
---

```{r, include = FALSE}
library(knitr)
library(tidyverse)
library(visdat)
# suppressPackageStartupMessages(library(dplyr))
knitr::opts_chunk$set(comment = "")
ufo <-read_csv("../data/ufo/ufo_data_complete.csv")
```




<!-- ## Data Input: `readr` -->

<!-- `read_delim()` and `read_csv()` from the`readr` package -->

<!-- ```{r eval=FALSE}  -->
<!-- # example:  -->
<!-- read_delim(file = "file.txt", delim = "\t")  -->
<!-- read_csv("file.csv") -->
<!-- ``` -->

<!-- ## Data Input -->

<!-- * The filename is the path to your file, in quotes -->
<!-- * The function will look in your "working directory" if no absolute file path is given -->
<!-- * Note that the filename can also be a path to a file on a website (e.g. 'www.someurl.com/table1.txt') -->

<!-- ## Data Input -->

<!-- What if your file is in the "Home" directory? -->

<!-- ```{r, out.width="80%", echo=FALSE, fig.alt='.'} -->
<!-- ottrpal::include_slide("https://docs.google.com/presentation/d/104LQkFTsC5R9vAC4HHj4mK7IjCTH55X6BkrAb3aDd6E/edit#slide=id.g13ea04b780d_1_121") -->
<!-- ``` -->

<!-- ## Data Input -->

<!-- Backtrack using the relative path with `../` like: -->

<!-- ```{r, eval = FALSE} -->
<!-- ufo <- read_csv("../ufo_data_complete.csv.gz") -->
<!-- ``` -->

<!-- ## Data Input -->

<!-- Or, read in from a subfolder: -->

<!-- ```{r, eval=FALSE} -->
<!-- ufo <- read_csv("data/ufo/ufo_data_complete.csv") -->
<!-- ``` -->

<!-- ```{r, echo=FALSE} -->
<!-- ufo <- read_csv("../data/ufo/ufo_data_complete.csv") -->
<!-- ``` -->

<!-- ## Data Input -->

<!-- The `read_delim()` and related functions return a "tibble" is a `data.frame` with special printing, which is the primary data format for most data cleaning and analyses.  -->

<!-- ```{r viewInput, message=FALSE} -->
<!-- class(ufo) -->
<!-- ``` -->

<!-- ## Data Input -->

<!-- ```{r, viewTibble} -->
<!-- ufo -->
<!-- ``` -->

<!-- ## Data Input -->

<!-- There are also data importing functions provided in base R (rather than the `readr` package), like `read.delim` and `read.csv`.  -->

<!-- These functions have slightly different syntax for reading in data, like `header` and `as.is`.  -->

<!-- However, while many online resources use the base R tools, recent versions of RStudio switched to use these new `readr` data import tools, so we will use them in the class for slides. They are also up to two times faster for reading in large datasets, and have a progress bar which is nice.  -->

<!-- ## Data input: `readr` highlights -->

<!-- - Modern, improved tools from `readr` R package: `read_delim()`, `read_csv()` -->
<!--   - needs a file path to be provided  -->
<!--   - parses the file into rows/columns, determines column type  -->
<!--   - returns a data frame  -->

<!-- - Some functions to look at a data frame:  -->
<!--   - `head()` shows first few rows -->
<!--   - `tail()` shows the last few rows -->
<!--   - `View()` shows the data as a spreadsheet -->
<!--   - `spec()` gives  specification of column types -->

<!-- ## Data input: other file types -->

<!-- * From `readr` package:  -->
<!--   - `read_delim()`: general delimited files  -->
<!--   - `read_csv()`: comma separated (CSV) files -->
<!--   - `read_tsv()`: tab separated files -->
<!--   - others  -->

<!-- * For reading Excel files, you can do one of:  -->
<!--   - use `read_excel()` function from `readxl` package -->
<!--   - use other packages: `xlsx`, `openxlsx` -->

<!-- ## Data input: other file types -->

<!-- * `haven` package has functions to read SAS, SPSS, Stata formats -->

<!-- ```{r, eval = FALSE} -->
<!-- library(haven) -->

<!-- # SAS -->
<!-- read_sas(file = "mtcars.sas7bdat") -->

<!-- # SPSS -->
<!-- read_sav(file = "mtcars.sav") -->

<!-- # Stata -->
<!-- read_dta(file = "mtcars.dta") -->
<!-- ``` -->


## What did I just read in?

* `nrow()` displays the number of rows of a data frame
* `ncol()` displays the number of columns 
* `dim()` displays a vector of length 2: # rows, # columns

```{r, dims}
nrow(ufo)
ncol(ufo)
dim(ufo)
```

## All Column Names

* `colnames()` displays the column names 

```{r, colnames}
colnames(ufo)
```

## Structure using `str()`

```{r, str}
str(ufo)
```

## Data Input

* Sometimes you get weird messages when reading in data.
* The problems()` function shows you any issues with the data read-in.

```{r stop_problems}
head(problems(ufo))
```

```{r}
dim(problems(ufo))
```


# After hours of cleaning...

## More ways to save: write_rds

If you want to save **one** object, you can use `readr::write_rds` to save to a compressed `rds` file:

```{r}
write_rds(ufo, file = "ufo_dataset.rds", compress = "xz")
```

## More ways to save: `read_rds`

To read this back in to R, you need to use `read_rds`, but **need to assign it**:

```{r}
ufo3 <- read_rds(file = "ufo_dataset.rds")
identical(ufo, ufo3) # test if they are the same 
```

## More ways to save: `save`

The `save` command can save a set of `R` objects into an "R data file", with the extension `.rda` or `.RData`.  

```{r, message = FALSE}
x = 5
save(ufo, x, file = "ufo_data.rda")
```

## More ways to save: load

The opposite of `save` is `load`.  The `ls()` command lists the items in the workspace/environment and `rm()` removes them

```{r, eval = FALSE}
load(file = "ufo_data.rda")
```

## Data Output

While its nice to be able to read in a variety of data formats, it's equally important to be able to output data somewhere.

`write_delim()`: Write a data frame to a delimited file
"This is about twice as fast as write.csv(), and never writes row names."

## Data Output

`x`: A data frame to write to disk

`file`: the file name where you want to R object written. It can be an absolute path, or a filename (which writes the file to your working directory)

`delim`: what character separates the columns?  
  
  * `","` = .csv - Note there is also a `write_csv()` function
  * `"\t"` = tab delimited

## Data Output

For example, we can write back out just the first 100 lines of the `ufo` dataset:

```{r writecsv}
write_delim(ufo[1:100,], file = "ufo_first100.csv", delim = ",")
```

```{r, include = FALSE}
file.remove("ufo_dataset.rds")
file.remove("ufo_data.RData")
file.remove("ufo_data.rda")
file.remove("ufo_first100.csv")
```