---
title: "Data Cleaning"
author: "Andrew Jaffe"
date: "July 12, 2016"
output:
  ioslides_presentation:
    css: styles.css
  beamer_presentation: default
---


```{r, echo = FALSE}
library(knitr)
opts_chunk$set(comment = "")
```

## Data Cleaning

In general, data cleaning is a process of investigating your data for inaccuracies, or recoding it in a way that makes it more manageable.

MOST IMPORTANT RULE - LOOK AT YOUR DATA!

## Useful checking functions

- `is.na` - is `TRUE` if the data is `FALSE` otherwise
- `!` - negation (NOT) 
    - if `is.na(x)` is `TRUE`, then `!is.na(x)` is `FALSE`

- `all` takes in a `logical` and will be `TRUE` if ALL are `TRUE`
    - `all(!is.na(x))` - are all values of `x` NOT NA
- `any` will be `TRUE` if ANY are true
    - `any(is.na(x))` - do we have any `NA`'s in `x`?
- `complete.cases` - returns `TRUE` if EVERY value of a row is NOT `NA`
    - very stringent condition 
    - `FALSE` missing one value (even if not important)
 
# Dealing with Missing Data

## Missing data types
One of the most important aspects of data cleaning is missing values.  

Types of "missing" data:

* `NA` - general missing data
* `NaN` - stands for "**N**ot **a** **N**umber", happens when you do 0/0.
* `Inf` and `-Inf` - Infinity, happens when you take a positive number (or negative number) by 0.

## Finding Missing data

Each missing data type has a function that returns `TRUE` if the data is missing:

* `NA` - `is.na`
* `NaN` - `is.nan`
* `Inf` and `-Inf` - `is.infinite`
* `is.finite` returns `FALSE` for all missing data and `TRUE` for non-missing

## Missing Data with Logicals

One important aspect (esp with subsetting) is that logical operations return `NA` for `NA` values.  Think about it, the data could be `> 2` or not we don't know, 
so `R` says there is no `TRUE` or `FALSE`, so that is missing:
```{r}
x = c(0, NA, 2, 3, 4)
x > 2
```

## Missing Data with Logicals

What to do?  What if we want if `x > 2` and `x` isn't `NA`?  
Don't do `x != NA`, do `x > 2` and `x` is NOT `NA`:

```{r}
x != NA
x > 2 & !is.na(x)
```

## Missing Data with Logicals

What about seeing if a value is equal to multiple values?  You can do `(x == 1 | x == 2) & !is.na(x)`, but that is not efficient.  

```{r}
(x == 0 | x == 2) # has NA
(x == 0 | x == 2) & !is.na(x) # No NA
```

what to do?

## Missing Data with Logicals: `%in%`

Introduce the `%in%` operator:
```{r}
x %in% c(0, 2) # NEVER has NA and returns logical
```

reads "return `TRUE` if `x` is in 0 or 2". (Like `inlist` in Stata).

## Missing Data with Logicals: `%in%`

NEVER has NA, even if you put it there (BUT DON'T DO THIS):

```{r}
x %in% c(0, 2, NA) # NEVER has NA and returns logical
x %in% c(0, 2) | is.na(x)
```


## Missing Data with Operations

Similarly with logicals, operations/arithmetic with `NA` will result in `NA`s:

```{r}
x + 2
x * 2
```
 
# Tables and Tabulations

## Useful checking functions

- `unique` - gives you the unique values of a variable
- `table(x)` - will give a one-way table of `x`
    - `table(x, useNA = "ifany")` - will have row NA 
- `table(x, y)` - will give a cross-tab of `x` and `y`

## Creating One-way Tables {.smaller}

Here we will use `table` to make tabulations of the data.   Look at `?table` to see options for missing data.

```{r table}
unique(x)
table(x)
table(x, useNA = "ifany") # will not 
```


## Creating One-way Tables {.smaller}

`useNA = "ifany"` will not have NA in table heading if no `NA`:
```{r onetab_ifany}
table(c(0, 1, 2, 3, 2, 3, 3, 2,2, 3), 
        useNA = "ifany")
```

## Creating One-way Tables {.smaller}

You can set `useNA = "always"` to have it always have a column for `NA`
```{r onetab}
table(c(0, 1, 2, 3, 2, 3, 3, 2,2, 3), 
        useNA = "always")
```

## Tables with Factors {.smaller}

If you use a `factor`, all levels will be given even if no exist! 
  - (May be wanted or not):

```{r onetab_fact}
fac = factor(c(0, 1, 2, 3, 2, 3, 3, 2,2, 3),
             levels = 1:4)
tab = table(fac)
tab
tab[ tab > 0 ]
```

## Creating Two-way Tables {.smaller}

A two-way table.  If you pass in 2 vectors, `table` creates a 2-dimensional table.

```{r}
tab <- table(c(0, 1, 2, 3, 2, 3, 3, 2,2, 3), 
             c(0, 1, 2, 3, 2, 3, 3, 4, 4, 3), 
              useNA = "always")
```

## Recoding to missing

Sometimes people code missing data in weird or inconsistent ways.

```{r codeNA}
ages = c(23,21,44,32,57,65,-999,54)
range(ages)
```


## Recoding to missing

How do we change the `-999` to be treated as missing?

```{r codeNA2}
ages[ages == -999] = NA
range(ages)
range(ages,na.rm=TRUE)
```

## Recoding from missing

What if you were the person that coded the `-999`

```{r codeFromNA}
is.na(ages)
ages[is.na(ages)] = -999
ages
```

## Download Salary FY2014 Data

From https://data.baltimorecity.gov/City-Government/Baltimore-City-Employee-Salaries-FY2014/2j28-xzd7
http://www.aejaffe.com/summerR_2016/data/Baltimore_City_Employee_Salaries_FY2014.csv

Read the CSV into R `Sal`:

```{r readSal}
Sal = read.csv("../data/Baltimore_City_Employee_Salaries_FY2014.csv",
               as.is = TRUE)
```

## Checking for logical conditions {.smaller}
* `any()` - checks if there are any `TRUE`s
* `all()` - checks if ALL are true

```{r isna}
head(Sal,2)
any(is.na(Sal$Name)) # are there any NAs?
```

# Recoding Variables

## Example of Recoding: base R

For example, let's say gender was coded as Male, M, m, Female, F, f.  Using Excel to find all of these would be a matter of filtering and changing all by hand or using if statements.

In R, you can simply do something like:

```{r, eval = FALSE}
data$gender[data$gender %in% 
    c("Male", "M", "m")] <- "Male"
```

## Example of Cleaning: more complicated

Sometimes though, it's not so simple.  That's where functions that find patterns come in very useful.

```{r gender, echo=FALSE}
set.seed(4) # random sample below - make sure same every time
gender <- sample(c("Male", "mAle", "MaLe", "M", "MALE", "Ma", "FeMAle", "F", "Woman", "Man", "Fm", "FEMALE"), 1000, replace = TRUE)
```

```{r gentab}
table(gender)
```

# String functions

## Useful String Functions

Useful String functions

* `toupper()`, `tolower()` - uppercase or lowercase your data:
* `str_trim()` (in the `stringr` package) or `trimws` in base 
    - will trim whitespace
* `nchar` - get the number of characters in a string
* `paste()` - paste strings together with a space 
* `paste0` - paste strings together with no space as default

## Pasting strings with `paste` and `paste0`

Paste can be very useful for joining vectors together:

```{r Paste}
paste("Visit", 1:5, sep = "_")
paste("Visit", 1:5, sep = "_", collapse = " ")
paste("To", "is going be the ", "we go to the store!", sep = "day ")
# and paste0 can be even simpler see ?paste0 
paste0("Visit",1:5)
```

## Paste Depicting How Collapse Works

```{r Paste2}
paste(1:5)
paste(1:5, collapse = " ")
```

## The `stringr` package

Like `dplyr`, the `stringr` package:

* Makes some things more intuitive
* Is different than base R
* Is used on forums for answers
* Has a standard format for most functions
    * the first argument is a string like first argument is a `data.frame` in `dplyr`



## Splitting/Find/Replace and Regular Expressions

* R can do much more than find exact matches for a whole string
* Like Perl and other languages, it can use regular expressions.
* What are regular expressions?
    * Ways to search for specific strings 
    * Can be very complicated or simple
    * Highly Useful - think "Find" on steroids


## A bit on Regular Expressions

* http://www.regular-expressions.info/reference.html
* They can use to match a large number of strings in one statement
* `.` matches any single character
* `*` means repeat as many (even if 0) more times the last character
* `?` makes the last thing optional
* `^` matches start of vector `^a` - starts with "a"
* `$` matches end of vector `b$` - ends with "b"

# Splitting Strings

## Substringing

Very similar:

Base R

* `substr(x, start, stop)` - substrings from position start to position stop
* `strsplit(x, split)` - splits strings up - returns list!

`stringr`

* `str_sub(x, start, end)` - substrings from position start to position end
* `str_split(string, pattern)` - splits strings up - returns list!


## Splitting String: base R

In base R, `strsplit` splits a vector on a string into a list
```{r strsplit}
x <- c("I really", "like writing", "R code programs")
y <- strsplit(x, split = " ") # returns a list
y
```

## Splitting String: `stringr`

`stringr::str_split` does the same thing:

```{r str_split}
library(stringr)
y2 <- str_split(x, " ") # returns a list
y2
```

## Using a fixed expression

One example case is when you want to split on a period "`.`".  In regular expressions `.` means **ANY** character, so

```{r}
str_split("I.like.strings", ".")
str_split("I.like.strings", fixed("."))
```

## Let's extract from `y`

```{r stsplit2}
suppressPackageStartupMessages(library(dplyr)) # must be loaded AFTER plyr
y[[2]]
sapply(y, dplyr::first) # on the fly
sapply(y, nth, 2) # on the fly
sapply(y, last) # on the fly
```


## 'Find' functions: base R

`grep`: `grep`, `grepl`, `regexpr` and `gregexpr` search for matches to argument pattern within each element of a character vector: they differ in the format of and amount of detail in the results. 

`grep(pattern, x, fixed=FALSE)`, where:

* pattern = character string containing a regular expression to be matched in the given character vector.

* x = a character vector where matches are sought, or an object which can be coerced by as.character to a character vector.

* If fixed=TRUE, it will do exact matching for the phrase anywhere in the vector (regular find)


## 'Find' functions: `stringr`

`str_detect`, `str_subset`, `str_replace`, and `str_replace_all` search for matches to argument pattern within each element of a character vector: they differ in the format of and amount of detail in the results. 

* `str_detect` - returns `TRUE` if `pattern` is found
* `str_subset` - returns only the strings which pattern were detected
    * convenient wrapper around `x[str_detect(x, pattern)]`
* `str_extract` - returns only strings which pattern were detected, but ONLY the pattern
* `str_replace` - replaces `pattern` with `replacement` the first time
* `str_replace_all` - replaces `pattern` with `replacement` as many times matched

## 'Find' functions: stringr compared to base R

Base R does not use these functions.  Here is a "translator" of the `stringr` function to base R functions

* `str_detect` - similar to `grepl` (return logical)
* `grep(value = FALSE)` is similar to `which(str_detect())`
* `str_subset` - similar to `grep(value = TRUE)` - return value of matched
* `str_replace` - similar to `sub` - replace one time 
* `str_replace_all` - similar to `gsub` - replace many times

## Let's look at modifier for `stringr`

`?modifiers`

* `fixed` - match everything exactly
* `regexp` - default - uses **reg**ular **exp**ressions
* `ignore_case` is an option to not have to use `tolower`


## Important Comparisons

Base R:

* Argument order is `(pattern, x)`
* Uses option `(fixed = TRUE)`

`stringr`

* Argument order is `(string, pattern)` aka `(x, pattern)`
* Uses function `fixed(pattern)`



## 'Find' functions: Finding Indices

These are the indices where the pattern match occurs:

```{r RawlMatch}
grep("Rawlings",Sal$Name)
which(grepl("Rawlings", Sal$Name))
which(str_detect(Sal$Name, "Rawlings"))
```

## 'Find' functions: Finding Logicals

These are the indices where the pattern match occurs:

```{r RawlMatch_log}
head(grepl("Rawlings",Sal$Name))
head(str_detect(Sal$Name, "Rawlings"))
```


## 'Find' functions: finding values, base R {.smaller}

```{r grepl}
grep("Rawlings",Sal$Name,value=TRUE)
Sal[grep("Rawlings",Sal$Name),]
```

## 'Find' functions: finding values, `stringr` and `dplyr` {.smaller}

```{r ggrep}
str_subset(Sal$Name, "Rawlings")
Sal %>% filter(str_detect(Name, "Rawlings"))
```

## Showing differnce in `str_extract`

`str_extract` extracts just the matched string

```{r ggrep2}
ss = str_extract(Sal$Name, "Rawling")
head(ss)
ss[ !is.na(ss)]
```

## Showing differnce in `str_extract` and  `str_extract_all`

`str_extract_all` extracts all the matched strings

```{r }
head(str_extract(Sal$AgencyID, "\\d"))
head(str_extract_all(Sal$AgencyID, "\\d"), 2)
```



## Using Regular Expressions
* Look for any name that starts with:
    * Payne at the beginning, 
    * Leonard and then an S
    * Spence then capital C

```{r grepstar}
head(grep("^Payne.*", x = Sal$Name, value = TRUE), 3)
```

```{r grepstar2}
head(grep("Leonard.?S", x = Sal$Name, value = TRUE))
head(grep("Spence.*C.*", x = Sal$Name, value = TRUE))
```



## Using Regular Expressions: `stringr`
```{r grepstar_stringr}
head(str_subset( Sal$Name, "^Payne.*"), 3)
```

```{r grepstar2_stringr}
head(str_subset( Sal$Name, "Leonard.?S"))
head(str_subset( Sal$Name, "Spence.*C.*"))
```

## Replace

Let's say we wanted to sort the data set by Annual Salary:

```{r classSal}
class(Sal$AnnualSalary)
```

```{r orderstring}
sort(c("1", "2", "10")) #  not sort correctly (order simply ranks the data)
order(c("1", "2", "10"))
```

## Replace

So we must change the annual pay into a numeric:
```{r destringSal}
head(Sal$AnnualSalary, 4)
head(as.numeric(Sal$AnnualSalary), 4)
```

R didn't like the `$` so it thought turned them all to `NA`.

`sub()` and `gsub()` can do the replacing part in base R.

## Replacing and subbing {.smaller}

Now we can replace the `$` with nothing (used `fixed=TRUE` because `$` means ending):

```{r orderSal}
Sal$AnnualSalary <- as.numeric(gsub(pattern = "$", replacement="", 
                              Sal$AnnualSalary, fixed=TRUE))
Sal <- Sal[order(Sal$AnnualSalary, decreasing=TRUE), ] 
Sal[1:5, c("Name", "AnnualSalary", "JobTitle")]
```

## Replacing and subbing: `stringr` {.smaller}

We can do the same thing (with 2 piping operations!) in dplyr
```{r orderSal_stringr}
dplyr_sal = Sal
dplyr_sal = dplyr_sal %>% mutate( 
  AnnualSalary = AnnualSalary %>%
    str_replace(
      fixed("$"), 
      "") %>%
    as.numeric) %>%
  arrange(desc(AnnualSalary))
check_Sal = Sal
rownames(check_Sal) = NULL
all.equal(check_Sal, dplyr_sal)
```

