---
title: "Data Summarization"
author: "Data Wrangling in R"
output:
  ioslides_presentation:
    css: styles.css
    widescreen: yes
---


```{r, echo = FALSE, message=FALSE, error = FALSE}
library(knitr)
opts_chunk$set(comment = "", message = FALSE)
suppressWarnings({library(dplyr)})
library(readr)
library(tidyverse)
```

<style type="text/css">
pre { /* Code block - slightly smaller in this lecture */
    font-size: 20px;
}
</style>

## Data Summarization

* Basic statistical summarization
    * `mean(x)`: takes the mean of x
    * `sd(x)`: takes the standard deviation of x
    * `median(x)`: takes the median of x
    * `quantile(x)`: displays sample quantiles of x. Default is min, IQR, max
    * `range(x)`: displays the range. Same as `c(min(x), max(x))`
    * `sum(x)`: sum of x
    * `max(x)`: maximum value in x
    * `min(x)`: minimum value in x
    
* **all have the ** `na.rm =` **argument for missing data**

## Statistical summarization

These functions work on **vectors**:

```{r}
x <- c(1, 5, 7, 4, 2, 8)
mean(x)
mean(x, na.rm = TRUE) # Remove NAs if needed
```

## Statistical summarization

Summarization on a `data.frame`/`tibble`:

```{r}
mtcars %>% pull(hp) %>% mean() # alt: pull(mtcars, hp) %>% mean()
mean(mtcars$hp)
```

## Youth Tobacco Survey

Here we will be using the Youth Tobacco Survey data:

```{r}
yts <- 
  read_csv("http://sisbid.github.io/Data-Wrangling/data/Youth_Tobacco_Survey_YTS_Data.csv")
head(yts)
```

## Column to vector

Let's work with one column as a vector using `pull()`.

```{r, message = FALSE}
locations <- yts %>% pull(LocationDesc)
locations
```

## Length and unique

`unique(x)` will return the unique elements of `x`

```{r, message = FALSE}
unique(locations)
```

## Length and unique

`length` will tell you the length of a vector. Combined with `unique`, tells you the number of unique elements:

```{r}
length(unique(locations))
```

## Counting NAs

use `sum(is.na())`:

```{r}
sum(is.na(locations))
```

## `dplyr`: `count` 

Use `count` directly on a data.frame and column: count the number of rows in each group.

```{r, message = FALSE}
yts %>% count(LocationDesc)
```

## `dplyr`: `count` 

Multiple columns listed further subdivides the count.

```{r, message = FALSE}
yts %>% count(LocationDesc, TopicDesc)
```

## `dplyr`: `count` 

Option to sort the results with `sort = TRUE`

```{r, message = FALSE}
yts %>% count(LocationDesc, sort = TRUE)
```

## `dplyr`: `count` 

Instead of counting the number of rows in each group, `wt` computes sum(wt) for each group.

```{r, message = FALSE}
# Add up "Data_Value" for each LocationDesc category
yts %>% count(LocationDesc, wt = Data_Value)
```

# Grouping

## Perform Operations By Groups: dplyr

`group_by` allows you group the data set by variables/columns you specify:

```{r}
# Regular data
yts
```

## Perform Operations By Groups: dplyr

`group_by` allows you group the data set by variables/columns you specify:

```{r}
yts_grouped <- yts %>% group_by(Response)
yts_grouped
```

## Summarize the data: `dplyr` `summarize()` function

`summarize` is a helpful function to use after `group_by()`. It creates a summary table of a column you're interested in.

```{r}
yts %>% 
  summarize(mean_value = mean(Data_Value, na.rm = TRUE))
```

## Summarize the grouped data

It's grouped! Grouping doesn't change the data in any way, but how **functions operate on it**. Now we can summarize `Data_Value` (percent of respondents) by group:

```{r}
yts_grouped %>% 
  summarize(mean_value = mean(Data_Value, na.rm = TRUE))
```

## Use the `pipe` to string these together!

Pipe `yts` into `group_by`, then pipe that into `summarize`:

```{r}
yts %>%
  group_by(Response) %>%
  summarize(mean_value = mean(Data_Value, na.rm = TRUE),
            max_value = max(Data_Value, na.rm = TRUE))
```

## `group_by` with `mutate` - just add data

We can also use `mutate` to calculate the mean value for each year and add it as a column:

```{r}
yts %>%
  group_by(YEAR) %>%
  mutate(year_avg = mean(Data_Value, na.rm = TRUE)) %>%
  select(LocationDesc, Data_Value, year_avg)
```

## Use `n()` for sample size by group

There are other functions, such as `n()` count the number of observations.

```{r}
yts %>%
  group_by(YEAR) %>%
  summarize(n = n(),
            mean = mean(Data_Value, na.rm = TRUE))
```

# Iterative summaries

## Iterative summaries: `dplyr` `summarize()` and `across()` functions

Use the [`across`](https://dplyr.tidyverse.org/reference/across.html) function with `summarize()` to summarize across multiple columns of your data.

```{r, eval = FALSE}
# General format - Not the code!
across({ columns to go across }, ~ { summarization_function(.x, na.rm = ..) }) 
```
</div>

```{r}
yts %>%
  group_by(YEAR) %>%
  summarize(across( c(Data_Value, Data_Value_Std_Err, Sample_Size), 
                    ~ mean(.x, na.rm = TRUE)))
```

## Iterative summaries: `dplyr` `summarize()` and `across()` functions

Another example using select helpers (`??tidyr_tidy_select`):

```{r}
yts %>% 
  summarize(across( where(is.numeric), ~ mean(.x, na.rm = TRUE)))
```

## Data Summarization on data frames

* Basic statistical summarization for numeric data
    * `rowMeans(x)`: takes the means of each row of x
    * `colMeans(x)`: takes the means of each column of x
    * `rowSums(x)`: takes the sum of each row of x
    * `colSums(x)`: takes the sum of each column of x
    * `summary(x)`: for data frames, displays the quantile information

## `summary()` Function

Using `summary()` can give you rough snapshots of each numeric column (character columns are skipped):

```{r}
summary(yts)
```

## Summary

- summary stats (`mean()`) work with `pull()`
- `count(x)`: what unique values do you have? 
  - `pull()` to get vectors
  - `unique()` combined with `length()`
- `group_by()`: changes all subsequent functions
  - combine with `summarize()` to get statistics per group
  - combine with `across()` to programmatically select columns
- `summary(x)`: quantile information
