---
title: "Data I/O + Structure"
author: "Data Wrangling in R"
output:
  ioslides_presentation:
    css: styles.css
    widescreen: yes
  beamer_presentation: default
---

```{r, include = FALSE}
library(knitr)
library(tidyverse)
library(visdat)
# suppressPackageStartupMessages(library(dplyr))
knitr::opts_chunk$set(comment = "")
ufo = read_csv("../data/ufo/ufo_data_complete.csv")
```


## What did I just read in?

* `nrow()` displays the number of rows of a data frame
* `ncol()` displays the number of columns 
* `dim()` displays a vector of length 2: # rows, # columns

```{r, dims}
dim(ufo)
nrow(ufo)
ncol(ufo)
```

## All Column Names

* `colnames()` displays the column names 

```{r, colnames}
colnames(ufo)
```


## Data Input

* Sometimes you get weird messages when reading in data.
* The problems()` function shows you any issues with the data read-in.

```{r stop_problems}
problems(ufo)
dim(problems(ufo))
```

## Data Input

* The `spec()` functions show you the specification of how the data was read in, `cols_condense` can help with this.

```{r spec}
spec(ufo)
cols_condense(spec(ufo))
```

## Data Input

This specification is passed to `readr` functions:

```{r col_types}
ufo_char = read_csv("../data/ufo/ufo_data_complete.csv", col_types = cols(
  .default = col_character(),
  longitude = col_double()
))
dim(problems(ufo))
dim(problems(ufo_char))
```


## Data Input: Checking for problems

* The `stop_for_problems()` function will stop if your data had an error when reading in.  If this occurs, you can either use `col_types` (from `spec()`) for the problematic columns, or set `guess_max = Inf` (takes much longer):

```{r stop_problems2, eval=FALSE}
stop_for_problems(ufo)
```

## Missing data with the `vizdat`/`naniar` packages

The `vis_dat` function can give you an overview 

```{r vis_dat}
library(visdat) # only so many rows can be visualized are good
ufo_samp = ufo %>% sample_n(size = 1000)
vis_dat(ufo_samp)
```

## Missing data with the `vizdat`/`naniar` packages


```{r vis_miss}
vis_miss(ufo_samp)
```

## Missing data with the `vizdat`/`naniar` packages

```{r}
library(naniar)
gg_miss_var(ufo)
```


## Missing data with the `vizdat`/`naniar` packages

`miss_case_summary` which rows have missing data
```{r}
miss_case_summary(ufo)
```

## Missing data with the `vizdat`/`naniar` packages

`miss_var_summary` which variables have missing data

```{r}
miss_var_summary(ufo)
```

# After hours of cleaning...

## More ways to save: write_rds

If you want to save **one** object, you can use `readr::write_rds` to save to a compressed `rds` file:

```{r}
write_rds(ufo, path = "ufo_dataset.rds", compress = "xz")
```


## More ways to save: `read_rds`

To read this back in to R, you need to use `read_rds`, but **need to assign it**:

```{r}
ufo3 = read_rds(path = "ufo_dataset.rds")
identical(ufo, ufo3) # test if they are the same 
```

## More ways to save: `save`

The `save` command can save a set of `R` objects into an "R data file", with the extension `.rda` or `.RData`.  

```{r, message = FALSE}
x = 5
save(ufo, x, file = "ufo_data.rda")
```


## More ways to save: load

The opposite of `save` is `load`.  The `ls()` command lists the items in the workspace/environment and `rm` removes them


## Data Output

While its nice to be able to read in a variety of data formats, it's equally important to be able to output data somewhere.

`write_delim()`: Write a data frame to a delimited file
"This is about twice as fast as write.csv(), and never writes row names."

```{r}
args(readr::write_delim)
```

## Data Output

`x`: A data frame to write to disk

`path`: the file name where you want to R object written. It can be an absolute path, or a filename (which writes the file to your working directory)

`delim`: what character separates the columns? 

* "," = .csv - Note there is also a `write_csv()` function
* "\t" = tab delimited


## Data Output

For example, we can write back out the `ufo` dataset with the new column name:

```{r writecsv}
write_csv(ufo[1:100,], path = "ufo_first100.csv")
```

## Lab
[Link to Lab](http://sisbid.github.io/Module1/labs/data-io-lab-part2.Rmd)


```{r, include = FALSE}
file.remove("ufo_dataset.rds")
file.remove("ufo_data.RData")
file.remove("ufo_data.rda")
file.remove("ufo_first100.csv")
```