---
title: "Data Reshaping"
author: "Data Wrangling in R"
output:
  ioslides_presentation:
    css: styles.css
    widescreen: yes
---


```{r, echo = FALSE, include=FALSE}
library(knitr)
library(readr)
opts_chunk$set(comment = "")
library(tidyverse)
```


## Reshaping Data 

```{r, echo = FALSE, message=FALSE}
library(tidyverse)
```

In this module, we will show you how to:

1. Reshaping data from wide (fat) to long (tall)
2. Reshaping data from long (tall) to wide (fat)
3. Merging Data/Joins
4. Perform operations by a grouping variable

## Setup

We will show you how to do each operation in base R then show you how to use the `dplyr` or `tidyr` package to do the same operation (if applicable).  

See the "Data Wrangling Cheat Sheet using `dplyr` and `tidyr`":

* https://www.rstudio.com/wp-content/uploads/2015/02/data-wrangling-cheatsheet.pdf

## What is wide/long data?


```{r, echo = FALSE}
ex_wide = tibble(id = 1:2,
                     visit1 = c(10, 5),
                     visit2 = c(4, 6),
                     visit3 = c(3, NA)
                     )
ex_long = tibble(id = c(rep(1, 3), rep(2, 2)),
                     visit = c(1:3, 1:2),
                     value = c(10, 4, 3, 5, 6))
```

See http://www.cookbook-r.com/Manipulating_data/Converting_data_between_wide_and_long_format/

- Wide - multiple columns per observation 
    - e.g. visit1, visit2, visit3
```{r, echo = FALSE}
ex_wide
```

- Long - multiple rows per observation
```{r, echo = FALSE}
ex_long
```

## What is wide/long data?

<div style="font-size:40pt">
More accurately, data is wide or long **with respect** to certain variables.
</div>

## Data used: Charm City Circulator

http://sisbid.github.io/Module1/data/Charm_City_Circulator_Ridership.csv

```{r, message = FALSE}
circ = read_csv(
  paste0("http://sisbid.github.io/Module1/",
         "data/Charm_City_Circulator_Ridership.csv"))
head(circ, 2)
class(circ$date)
```

## Creating a Date class from a character date

```{r, message = FALSE}
library(lubridate) # great for dates!
```

```{r, message= FALSE}
sum(is.na(circ$date))
sum( circ$date == "")
circ = mutate(circ, date = mdy(date))
sum( is.na(circ$date) ) # all converted correctly
head(circ$date, 3)
class(circ$date)
```

## Reshaping data from wide (fat) to long (tall): base R

The `reshape` command exists.  It is a **confusing** function.  Don't use it.

## tidyr package

`tidyr` allows you to "tidy" your data.  We will be talking about:

- `gather` - make multiple columns into variables, (wide to long)
- `spread` - make a variable into multiple columns, (long to wide)
- `separate` - string into multiple columns
- `unite` - multiple columns into one string

- All the "join" functions for mergin are in `dplyr`


## Reshaping data from wide (fat) to long (tall): tidyr 

`tidyr::gather` - puts column data into rows.

We want the column names into "`var`" variable in the output dataset and the value in "`number`" variable. We then describe which columns we want to "gather:" 

```{r}
long = gather(circ, key = "var", value = "number", 
              -day, -date, -daily)
head(long, 4)
```


## Reshaping data from wide (fat) to long (tall): tidyr 

- Could be explicit on what we want to gather

```{r}
long = gather(circ, key = "var", value = "number", 
              starts_with("orange"), starts_with("purple"),
              starts_with("green"), starts_with("banner"))
long
```

## Reshaping data from wide (fat) to long (tall): tidyr 

```{r}
long %>% count(var)
```


## Making a separator

We will use `str_replace` from `stringr` to put `_` in the names

```{r}
long = long %>% mutate(
  var = var %>% 
    str_replace("Board", "_Board") %>% 
    str_replace("Alight", "_Alight") %>% 
    str_replace("Average", "_Average") 
)
long %>% count(var)
```

## Reshaping data from wide (fat) to long (tall): tidyr

Now each `var` is boardings, averages, or alightings.  We want to separate these so we can have these by route. Remember "." is special character:

```{r}
long = separate(long, var, into = c("route", "type"), sep = "_")
head(long, 2)
unique(long$route)
unique(long$type)
```

## Re-uniting all the routes

If we had the opposite problem, we could use the `unite` function:

```{r}
reunited = long %>% 
  unite(col = var, route, type, sep = "_")  
reunited %>% select(day, var) %>% head(3) %>% print
```
We could also use `paste`/`paste0`.



## Reshaping data from long (tall) to wide (fat): tidyr

In `tidyr`, the `spread` function spreads rows into columns.  Now we have a long data set, but we want to separate the Average, Alightings and Boardings into different columns:

```{r}
# have to remove missing days
wide = long %>% filter(!is.na(date))
wide = wide %>% spread(type, number)
head(wide)
```
## Pivoting Functions

`pivot_longer` and `pivot_wider` are new (as of late 2019) `tidyr` functions.

See link below:

https://tidyr.tidyverse.org/dev/articles/pivot.html

## Pivoting Functions

```{r}
long2 = circ %>% 
  rename_all(function(var) {
    var %>% 
    str_replace("Board", "_Board") %>% 
    str_replace("Alight", "_Alight") %>% 
    str_replace("Average", "_Average") 
  })
longer =long2 %>% pivot_longer(
  cols = matches("orange|purple|green|banner"),
  names_to = c("route", "type"),
  names_sep = "_"
)
head(longer)
```

## Pivoting Functions

```{r}
longer %>% 
  filter(!is.na(value)) %>% # keep where there is data
  pivot_wider(
    names_from = type,
    values_from = value
  )
```
