---
title: "Data I/O + Structure"
author: "Data Wrangling in R"
output:
  ioslides_presentation:
    css: styles.css
    widescreen: yes
  beamer_presentation: default
---

```{r, include = FALSE}
library(knitr)
library(tidyverse)
library(visdat)
# suppressPackageStartupMessages(library(dplyr))
knitr::opts_chunk$set(comment = "")
ufo <-read_csv("../data/ufo/ufo_data_complete.csv")
```

## What did I just read in?

* `nrow()` displays the number of rows of a data frame
* `ncol()` displays the number of columns 
* `dim()` displays a vector of length 2: # rows, # columns

```{r, dims}
nrow(ufo)
ncol(ufo)
dim(ufo)
```

## All Column Names

* `colnames()` displays the column names 

```{r, colnames}
colnames(ufo)
```

## Structure using `str()`

```{r, str}
str(ufo)
```

## Data Input

* Sometimes you get weird messages when reading in data.
* The problems()` function shows you any issues with the data read-in.

```{r stop_problems}
head(problems(ufo))
```

```{r}
dim(problems(ufo))
```


# After hours of cleaning...

## More ways to save: write_rds

If you want to save **one** object, you can use `readr::write_rds` to save to a compressed `rds` file:

```{r}
write_rds(ufo, file = "ufo_dataset.rds", compress = "xz")
```

## More ways to save: `read_rds`

To read this back in to R, you need to use `read_rds`, but **need to assign it**:

```{r}
ufo3 <- read_rds(file = "ufo_dataset.rds")
identical(ufo, ufo3) # test if they are the same 
```

## More ways to save: `save`

The `save` command can save a set of `R` objects into an "R data file", with the extension `.rda` or `.RData`.  

```{r, message = FALSE}
x = 5
save(ufo, x, file = "ufo_data.rda")
```

## More ways to save: load

The opposite of `save` is `load`.  The `ls()` command lists the items in the workspace/environment and `rm()` removes them

```{r, eval = FALSE}
load(file = "ufo_data.rda")
```

## Data Output

While its nice to be able to read in a variety of data formats, it's equally important to be able to output data somewhere.

`write_delim()`: Write a data frame to a delimited file
"This is about twice as fast as write.csv(), and never writes row names."

## Data Output

`x`: A data frame to write to disk

`file`: the file name where you want to R object written. It can be an absolute path, or a filename (which writes the file to your working directory)

`delim`: what character separates the columns?  
  
  * `","` = .csv - Note there is also a `write_csv()` function
  * `"\t"` = tab delimited

## Data Output

For example, we can write back out just the first 100 lines of the `ufo` dataset:

```{r writecsv}
write_delim(ufo[1:100,], file = "ufo_first100.csv", delim = ",")
```

```{r, include = FALSE}
file.remove("ufo_dataset.rds")
file.remove("ufo_data.RData")
file.remove("ufo_data.rda")
file.remove("ufo_first100.csv")
```