---
title: "Subsetting Data in R"
author: "Data Wrangling in R"
output:
  ioslides_presentation:
    css: styles.css
    widescreen: yes
  beamer_presentation: default
---

```{r, echo = FALSE}
library(knitr)
opts_chunk$set(comment = "")
suppressPackageStartupMessages(library(dplyr))
library(dplyr)
library(ggplot2)
```

## Overview

```{css, echo=FALSE}
.watch-out {
  #background-color: lightpink;
  border: 3px solid red;
  font-weight: bold;
}
# from: https://bookdown.org/yihui/rmarkdown-cookbook/chunk-styling.html
```



We showed different ways to read data into R using:  

```{r, eval = FALSE}
readr::read_csv()
readr::read_delim()
readxl::read_excel()
```

In this module, we will show you how select rows and columns of datasets. 

## Setup


We will be using the $\color{red}{\text{dplyr}}$  package in the tidyverse.

Here are several resources on how to use `dplyr`: 

* https://dplyr.tidyverse.org/
* https://r4ds.had.co.nz/
* https://cran.rstudio.com/web/packages/dplyr/vignettes/dplyr.html
* https://stat545.com/dplyr-intro.html

The `dplyr` package also interfaces well with tibbles.

## Dataset

We will be using the `diamonds` dataset in the `ggplot2` package as an example (so make sure you initiate the `ggplot2` package if you are following along on your own). 

## Selecting a single column of a `data.frame`: 

If you were just selecting a single column, and you wanted the output to be a vector (and not a `tibble`), you need to use `pull` function. Since this is a long vector we will just show the first 6 values using the `head` function around the output of the `pull` function.
```{r, class.source="watch-out"}
head(pull(diamonds,carat))
```

## Selecting a single column of a `data.frame:

The `pull` function is equivalent to using the `$` method (in base R). 

We can grab the `carat` column using the `$` operator. 

```{r}
head(pull(diamonds,carat))
head(diamonds$carat)
```

Note this does *not* return a `tibble` (or `data.frame`) but rather a vector. 

## Selecting a single column of a `data.frame`: 

The `select` function extracts one or more columns from a `tibble` or `data.frame` and returns a `tibble` (not a vector).

```{r}
select(diamonds, carat)
```


## Selecting multiple columns of a `data.frame`: 

The `select` command from `dplyr` is very flexible. You just need to list all columns you want to extract separated by commas.

```{r}
select(diamonds, carat, depth)
```


## Using the `pipe` (comes with `dplyr`):

That was a lot of typing and nested functions, which can be confusing. Recently, the pipe `%>%` makes things such as this much more readable.  It reads left side "pipes" into right side.  RStudio `CMD/Ctrl + Shift + M` shortcut. 

## Using the `pipe` (comes with `dplyr`):

Pipe `diamonds` into `select`, then pipe that into `pull`, and then show the `head`:

```{r}
diamonds %>% pull(carat) %>% head()
```

## See the Select "helpers"

Run the command:

```{r, eval = FALSE}
??tidyselect::select_helpers
```

Here are a few:

```{r, eval = FALSE}
last_col()
ends_with()
contains() # like searching
matches() # Matches a regular expression - cover later
```

## Tidyselect helpers

For example, we can take all columns that start with a "c":

```{r}
diamonds %>% select(starts_with("c"))
```

## Tidyselect helpers

Or we can take all columns that end with an "e":

```{r}
diamonds %>% select(ends_with("e"))
```

## Tidyselect helpers

We are going to cover "fancier" ways of matching column names (and strings more generally) in the data cleaning lecture. 

## Subset rows of a `data.frame`:

The command in `dplyr` for subsetting rows is `filter`.  Try `?filter`.

The easiest way to filter is by testing whether numeric observations are greater than or less than some cutoff:

```{r}
filter(diamonds, depth > 60)
```

Note, no `$` or subsetting is necessary.  R "knows" `depth` refers to a column of `diamonds`.

## Subset rows of a `data.frame`:

You can also using piping here:

```{r}
diamonds %>% filter(depth > 60)
```

## Subset rows of a `data.frame`:

You can combine filtering on multiple columns by separating the filter arguments with commas:

```{r}
diamonds %>% filter(depth > 60, table > 60, price > 2775)
```

## Subset rows of a `data.frame`:

You can also filter character strings by a single value or category:

```{r}
diamonds %>% filter(color == "I", 
            clarity == "SI2", cut == "Premium")
```

## Subset rows of a `data.frame`:

Sometimes you want to be able to filter on matching several values or categories. The `%in%` operator is useful here:

```{r}
diamonds %>% filter(clarity %in% c("SI1", "SI2"))
```

## Subset rows of a `data.frame`:

You can mix and match filtering on numeric and categorical/character columns in the same `filter()` command:

```{r}
diamonds %>% filter(clarity %in% c("SI1", "SI2"),
                    cut == "Premium", price > 3000)
```

## Subset rows of a `data.frame`:

Other useful logical tests:

`&` : AND

`|` : OR

`<=` : less than or equals

`>=` : greater than or equals

`!=` : not equals

## Subset rows of a `data.frame`:

The AND operator (`&`) is the what is being performed "behind the scenes" when chaining together filter statements with commas: 

```{r}
diamonds %>% filter(depth > 60 & table > 60 & price > 2775)
```

You can use either syntax.

## Subset rows of a `data.frame`:

The OR operator (`|`) is more permissive than the AND operator:

```{r}
diamonds %>% filter(depth > 60 | table > 60 | price > 2775)
```

## Subset rows of a `data.frame`:

The OR operator (`|`) can be a substitute for `%in%` (although it might take more typing):

```{r}
diamonds %>% filter(clarity =="SI1" | clarity == "SI2")
```

## Combining `filter` and `select`:

You can combine `filter` and `select` to subset the rows and columns, respectively, of a `data.frame`:

```{r}
diamonds %>% 
  filter(clarity == "SI2") %>% 
  select(starts_with("c"))
```

## Combining `filter` and `select`:

The order of these functions matters though, since you can remove columns that you might want to filter on. 

```{r,eval=FALSE}
diamonds %>% 
  select(starts_with("c") %>%
  filter(table > 60))
```