---
title: "Data Summarization"
author: "Data Wrangling in R"
output:
  ioslides_presentation:
    css: styles.css
    widescreen: yes
---

```{r, include = FALSE}
library(knitr)
opts_chunk$set(comment = "", message = FALSE)
library(tidyverse)
```


## Quick Data read in

We can use the Charm City Circulator Dataset from "http://sisbid.github.io/Data-Wrangling/data/Charm_City_Circulator_Ridership.csv".  
```{r}
circ = read_csv(paste0("http://sisbid.github.io/Data-Wrangling/",
                       "data/Charm_City_Circulator_Ridership.csv"))
# or circ = read_csv("../data/Charm_City_Circulator_Ridership.csv")
```

## Head and Tail Commands

The `head`/`tail` commands displays the first/last `6` (default) rows:

```{r}
head(circ)
```

## Head and Tail Commands

The `head`/`tail` commands displays the first/last `6` (default) rows:

```{r}
tail(circ, 10)
```

## Data Summarization

* Basic statistical summarization is key after cleaning data!
    * `mean(x)`: takes the mean of x
    * `sd(x)`: takes the standard deviation of x
    * `median(x)`: takes the median of x
    * `quantile(x)`: displays sample quantiles of x. Default is min, IQR, max
    * `range(x)`: displays the range. Same as `c(min(x), max(x))`
    * `sum(x)`: sum of x
    * **all have a **`na.rm` for missing data 
* Transformations
    * `log`, `log2`, `log10` - log transformation
    * `sqrt` - square root

## Statistical summarization

Remember `NA` is "missing" so it's unknown what the mean or sum of something is (by default). `na.rm` argument ("remove NAs"). 

```{r}
mean(circ$daily)
sum(circ$daily)
mean(circ$daily, na.rm = TRUE)
sum(circ$daily, na.rm = TRUE)
```

## Statistical summarization

```{r}
quantile(circ$daily, na.rm = TRUE)
quantile(circ$daily, na.rm = TRUE, probs = c(0.6, 0.84))
median(circ$daily, na.rm = TRUE)
```

## Length and unique

`unique(x)` will return the unique elements of `x`

```{r, message = FALSE}
unique(circ$day)
```

`length` will tell you the length of a vector. Combined with `unique`, tells you the number of unique elements:

```{r}
length(unique(circ$date))
```

## Table

`table(x)` will return a frequency table of unique elements of `x`

```{r, message = FALSE}
table(circ$day)
```

## The `tidy` way: `dplyr`: `count` 

```{r, message = FALSE}
circ %>% count(day)
```

## The `tidy` way: `dplyr`: `count` 

```{r, message = FALSE}
circ %>% mutate(many_riders = daily > 1000) %>% count(many_riders, day)
```

## Summarize the data: `dplyr` `summarize`/`summarise` function

`dplyr::summarise` will allow you to summarize data.  Format is `new_column = SUMMARY`. If you don't set a `new` name, it will be a messy output:

```{r}
circ %>% 
  summarize(mean_purple = mean(purpleAverage, na.rm = TRUE),
            mean(bannerAverage, na.rm = TRUE))
```

## `across` - summarize multiple columns!

If you would like to a bunch of columns, you can use `across` and pass in a function (with other arguments) with select helpers:

```{r}
circ %>% summarise(across(ends_with("Boardings"), mean, na.rm = TRUE))
```

## Perform Operations By Groups: dplyr

`group_by` allows you group the data set by grouping variables:

```{r}
sub_circ = circ %>% group_by(day)
head(sub_circ)
```

- doesn't change the data in any way, but how **functions operate on it**

## Summarize the data

It's grouped!

```{r}
sub_circ %>% summarize(avg_daily = mean(daily, na.rm = TRUE))
```

## Using the `pipe` 

Pipe `sub_circ` into `group_by`, then pipe that into `summarise`:

```{r}
day_avgs = circ %>% 
  group_by(day) %>% 
  summarize(avg_daily = mean(daily, na.rm = TRUE))
head(day_avgs)
```

## Ungroup the data

You usually want to perform operations on groups and may want to redefine the groups.  The `ungroup` function will allow you to clear the groups from the data: 

```{r}
sub_circ = ungroup(sub_circ)
sub_circ
```

## `group_by` with `mutate`

We can use `mutate` instead of `summarise` to add the summary back in to the original data!

```{r}
circ %>% 
  group_by(day) %>% 
  mutate(mean = mean(daily, na.rm = TRUE)) %>% 
  select(day, date, mean, daily)
```

## Counting with `n()`

Standard statistics can be calculated: `n()` counts the number of observations.

```{r}
circ %>% 
  group_by(day) %>% 
  summarize(n = n(),
            mean = mean(daily, na.rm = TRUE)) %>% 
  head
```

## Statistical summarization

`t.test` is good for t-tests, but also gives a mean and 95\% CI:

```{r}
t.test(circ$daily)
broom::tidy(t.test(circ$daily))
```

## Data Summarization on matrices/data frames

* Basic statistical summarization
    * `rowMeans(x)`: takes the means of each row of x
    * `colMeans(x)`: takes the means of each column of x
    * `rowSums(x)`: takes the sum of each row of x
    * `colSums(x)`: takes the sum of each column of x
    * `summary(x)`: for data frames, displays the quantile information
* The `matrixStats` package has additional `row*` and `col*` functions 
    * Like `rowSds`, `colQuantiles`

## Column and Row means

`colMeans` and `rowMeans` must work on **all numeric data**.  We will subset the boardings

```{r colMeans}
avgs = circ %>% select(ends_with("Boardings"))
colMeans(avgs, na.rm = TRUE)
```

```{r}
circ = circ %>% mutate(mean_boarding = rowMeans(avgs, na.rm = TRUE))
head(circ %>% select(day, mean_boarding))
```
