---
title: "Data I/O + Structure"
author: "Data Wrangling in R"
output:
  beamer_presentation: default
  ioslides_presentation:
    css: styles.css
---

```{r, echo = FALSE, message=FALSE}
library(knitr)
library(readr)
# suppressPackageStartupMessages(library(dplyr))
opts_chunk$set(comment = "")
```

## Data Input

* Sometimes you get weird messages when reading in data:
* The `spec()` and `problems()` functions show you the specification of how the data was read in.  

```{r stop_problems}
dim(problems(ufo))
spec(ufo)
```


## Data Input: Checking for problems

* The `stop_for_problems()` function will stop if your data had an error when reading in.  If this occurs, you can either use `col_types` (from `spec()`) for the problematic columns, or set `guess_max = Inf` (takes much longer):

```{r stop_problems2, eval=FALSE}
stop_for_problems(ufo)
```


## More ways to save: write_rds

If you want to save **one** object, you can use `readr::write_rds` to save to an `rds` file:

```{r}
write_rds(ufo, path = "ufo_dataset.rds")
```


## More ways to save: read_rds

To read this back in to R, you need to use `read_rds`, but **need to assign it**:

```{r}
ufo3 = read_rds(path = "ufo_dataset.rds")
identical(ufo, ufo3) # test if they are the same 
```

## More ways to save: save

The `save` command can save a set of `R` objects into an "R data file", with the extension `.rda` or `.RData`.  

```{r, message = FALSE}
x = 5
save(ufo, x, file = "ufo_data.RData")
```


## More ways to save: load

The opposite of `save` is `load`.  The `ls()` command lists the items in the workspace/environment and `rm` removes them:

## What did I just read in?

* `nrow()` displays the number of rows of a data frame
* `ncol()` displays the number of columns 
* `dim()` displays a vector of length 2: # rows, # columns

```{r, dims}
dim(ufo)
nrow(ufo)
ncol(ufo)
```

## Data Summaries

* `colnames()` displays the column names (if any) and `rownames()` displays the row names (if any)
* Note that tibbles do not have row names

```{r, colnames}
colnames(ufo)
```


## Data Output

While its nice to be able to read in a variety of data formats, it's equally important to be able to output data somewhere.

`write_delim()`: Write a data frame to a delimited file
"This is about twice as fast as write.csv(), and never writes row names."

```{r}
args(readr::write_delim)
```

## Data Output

`x`: A data frame to write to disk

`path`: the file name where you want to R object written. It can be an absolute path, or a filename (which writes the file to your working directory)

`delim`: what character separates the columns? 

* "," = .csv - Note there is also a `write_csv()` function
* "\t" = tab delimited


## Data Output

For example, we can write back out the Monuments dataset with the new column name:

```{r writecsv}
write_csv(ufo[1:100,], path="ufo_first100.csv")
```
