---
title: "Data Cleaning"
author: "Data Wrangling in R"
output:
  ioslides_presentation:
    css: styles.css
    widescreen: yes
  beamer_presentation: default
---


```{r, include = FALSE}
library(knitr)
library(tidyverse)
opts_chunk$set(comment = "")
```

# Dealing with Missing Data

## Missing data types
One of the most important aspects of data cleaning is missing values.  

Types of "missing" data:

* `NA` - general missing data
* `NaN` - stands for "**N**ot **a** **N**umber", happens when you do 0/0.
* `Inf` and `-Inf` - Infinity, happens when you take a positive number (or negative number) by 0.

## Finding Missing data

Each missing data type has a function that returns `TRUE` if the data is missing:

* `NA` - `is.na`
* `NaN` - `is.nan`
* `Inf` and `-Inf` - `is.infinite`
* `is.finite` returns `FALSE` for all missing data and `TRUE` for non-missing

## Missing Data with Logicals

One important aspect (esp with subsetting) is that logical operations return `NA` for `NA` values.  Think about it, the data could be `> 2` or not we don't know, 
so `R` says there is no `TRUE` or `FALSE`, so that is missing:
```{r}
x = c(0, NA, 2, 3, 4)
x > 2
```

## Missing Data with Logicals

What to do?  What if we want if `x > 2` and `x` isn't `NA`?  
Don't do `x != NA`, do `x > 2` and `x` is NOT `NA`:

```{r}
x != NA
x > 2 & !is.na(x)
```

## Missing Data with Logicals

What about seeing if a value is equal to multiple values?  You can do `(x == 1 | x == 2) & !is.na(x)`, but that is not efficient.  

```{r}
(x == 0 | x == 2) # has NA
(x == 0 | x == 2) & !is.na(x) # No NA
```

what to do?

## Missing Data with Logicals: `%in%`

Introduce the `%in%` operator:
```{r}
x %in% c(0, 2) # NEVER has NA and returns logical
```

reads "return `TRUE` if `x` is in 0 or 2". (Like `inlist` in Stata).

## Missing Data with Logicals: `%in%`

NEVER has NA, even if you put it there (BUT DON'T DO THIS):

```{r}
x %in% c(0, 2, NA) # NEVER has NA and returns logical
x %in% c(0, 2) | is.na(x)
```

## Filtering and tibbles

Filter removes missing values, have to keep them if you want them:

```{r}
df = tibble(x = x)
df %>% filter(x > 2)
filter(df, between(x, -1, 3) | is.na(x))
```


## Missing Data with Operations

Similarly with logicals, operations/arithmetic with `NA` will result in `NA`s:

```{r}
x + 2
x * 2
```
 
 
## Recoding to missing

Sometimes people code missing data in weird or inconsistent ways.

```{r codeNA}
ages = c(23,21,44,32,57,65,-999,54)
range(ages)
```


## Recoding to missing

How do we change the `-999` to be treated as missing?

```{r codeNA2}
ages[ages == -999] = NA
range(ages)
range(ages,na.rm=TRUE)
```

## Recoding from missing

What if you were the person that coded the `-999`

```{r codeFromNA}
is.na(ages)
ages[is.na(ages)] = -999
ages
```
