---
title: "Data Cleaning"
author: "Data Wrangling in R"
output:
  beamer_presentation: default
  ioslides_presentation:
    css: styles.css
---


```{r, echo = FALSE,message=FALSE}
library(knitr)
library(tidyverse)
opts_chunk$set(comment = "")
```

## Data Cleaning

In general, data cleaning is a process of investigating your data for inaccuracies, or recoding it in a way that makes it more manageable.

MOST IMPORTANT RULE - LOOK AT YOUR DATA!

## Useful checking functions

- `is.na` - is `TRUE` if the data is `FALSE` otherwise
- `!` - negation (NOT) 
    - if `is.na(x)` is `TRUE`, then `!is.na(x)` is `FALSE`

- `all` takes in a `logical` and will be `TRUE` if ALL are `TRUE`
    - `all(!is.na(x))` - are all values of `x` NOT NA
- `any` will be `TRUE` if ANY are true
    - `any(is.na(x))` - do we have any `NA`'s in `x`?
- `complete.cases` - returns `TRUE` if EVERY value of a row is NOT `NA`
    - very stringent condition 
    - `FALSE` missing one value (even if not important)
 
# Dealing with Missing Data

## Missing data types
One of the most important aspects of data cleaning is missing values.  

Types of "missing" data:

* `NA` - general missing data
* `NaN` - stands for "**N**ot **a** **N**umber", happens when you do 0/0.
* `Inf` and `-Inf` - Infinity, happens when you take a positive number (or negative number) by 0.

## Finding Missing data

Each missing data type has a function that returns `TRUE` if the data is missing:

* `NA` - `is.na`
* `NaN` - `is.nan`
* `Inf` and `-Inf` - `is.infinite`
* `is.finite` returns `FALSE` for all missing data and `TRUE` for non-missing

## Missing Data with Logicals

One important aspect (esp with subsetting) is that logical operations return `NA` for `NA` values.  Think about it, the data could be `> 2` or not we don't know, 
so `R` says there is no `TRUE` or `FALSE`, so that is missing:
```{r}
x = c(0, NA, 2, 3, 4)
x > 2
```

## Missing Data with Logicals

What to do?  What if we want if `x > 2` and `x` isn't `NA`?  
Don't do `x != NA`, do `x > 2` and `x` is NOT `NA`:

```{r}
x != NA
x > 2 & !is.na(x)
```

## Missing Data with Logicals

What about seeing if a value is equal to multiple values?  You can do `(x == 1 | x == 2) & !is.na(x)`, but that is not efficient.  

```{r}
(x == 0 | x == 2) # has NA
(x == 0 | x == 2) & !is.na(x) # No NA
```

what to do?

## Missing Data with Logicals: `%in%`

Introduce the `%in%` operator:
```{r}
x %in% c(0, 2) # NEVER has NA and returns logical
```

reads "return `TRUE` if `x` is in 0 or 2". (Like `inlist` in Stata).

## Missing Data with Logicals: `%in%`

NEVER has NA, even if you put it there (BUT DON'T DO THIS):

```{r}
x %in% c(0, 2, NA) # NEVER has NA and returns logical
x %in% c(0, 2) | is.na(x)
```

## Filtering and tibbles

Filter removes missing values, have to keep them if you want them:

```{r}
df = tibble(x = x)
df %>% filter(x > 2)
filter(df, between(x, -1, 3) | is.na(x))
```


## Missing Data with Operations

Similarly with logicals, operations/arithmetic with `NA` will result in `NA`s:

```{r}
x + 2
x * 2
```
 
# Tables and Tabulations

## Useful checking functions

- `unique` - gives you the unique values of a variable
- `table(x)` - will give a one-way table of `x`
    - `table(x, useNA = "ifany")` - will have row NA 
- `table(x, y)` - will give a cross-tab of `x` and `y`

## Creating One-way Tables {.smaller}

Here we will use `table` to make tabulations of the data.   Look at `?table` to see options for missing data.

```{r table}
unique(x)
table(x)
table(x, useNA = "ifany") # will not 
```


## Creating One-way Tables {.smaller}

`useNA = "ifany"` will not have NA in table heading if no `NA`:
```{r onetab_ifany}
table(c(0, 1, 2, 3, 2, 3, 3, 2,2, 3), 
        useNA = "ifany")
```

## Creating One-way Tables {.smaller}

You can set `useNA = "always"` to have it always have a column for `NA`
```{r onetab}
table(c(0, 1, 2, 3, 2, 3, 3, 2,2, 3), 
        useNA = "always")
```

## Tables with Factors {.smaller}

If you use a `factor`, all levels will be given even if no exist! 
  - (May be wanted or not):

```{r onetab_fact}
fac = factor(c(0, 1, 2, 3, 2, 3, 3, 2,2, 3),
             levels = 1:4)
tab = table(fac)
tab
tab[ tab > 0 ]
```

## Creating Two-way Tables {.smaller}

A two-way table.  If you pass in 2 vectors, `table` creates a 2-dimensional table.

```{r}
tab <- table(c(0, 0, 0, 1, 1,2, 3, 2, 3, 3, 2,2, 3,3,2,3,3), 
             c(0, 0,1, 1, 2, 3, 3, 2, 3, 3, 4, 4, 3,2,2,1,1), 
              useNA = "always")
```

## Finding Row or Column Totals

`margin.table` finds the marginal sums of the table.  `margin` is 1 for rows, 2 for columns in general in `R`.  Here is the column sums of the table:

```{r margin}
margin.table(tab, 2)
```

## Proportion Tables 

`prop.table` finds the marginal proportions of the table. Think of it dividing the table by it's respective marginal totals.  If `margin` not set, divides by overall total.

```{r table2}
prop.table(tab)
prop.table(tab,1) * 100
```

## Recoding to missing

Sometimes people code missing data in weird or inconsistent ways.

```{r codeNA}
ages = c(23,21,44,32,57,65,-999,54)
range(ages)
```


## Recoding to missing

How do we change the `-999` to be treated as missing?

```{r codeNA2}
ages[ages == -999] = NA
range(ages)
range(ages,na.rm=TRUE)
```

## Recoding from missing

What if you were the person that coded the `-999`

```{r codeFromNA}
is.na(ages)
ages[is.na(ages)] = -999
ages
```

## Read in the UFO dataset

* Read in data from RStudio Cloud or download from: http://sisbid.github.io/Module1/data/ufo/ufo_data_complete.csv.gz

```{r, importUfo}
ufo = read_csv("../data/ufo/ufo_data_complete.csv")
```

## Data cleaning "before" R

You saw warning messages when reading in this dataset. 

```{r}
p = problems(ufo)
p
```

We might want to check these out manually

## Data cleaning "before" R

```{r}
ufo[p$row, ]
```

## Data cleaning "before" R

Let's just drop those rows for now

```{r}
ufo = ufo[-p$row,]
```

## Checking for logical conditions {.smaller}
* `any()` - checks if there are any `TRUE`s
* `all()` - checks if ALL are true

```{r isna}
any(is.na(ufo$state)) # are there any NAs?
table(is.na(ufo$state)) # are there any NAs?
```

# Recoding Variables

## Example of Recoding: base R

For example, let's say gender was coded as Male, M, m, Female, F, f.  Using Excel to find all of these would be a matter of filtering and changing all by hand or using if statements.

In R, you can simply do something like:

```{r, eval = FALSE}
data$gender[data$gender %in% 
    c("Male", "M", "m")] <- "Male"
```

## Example of Cleaning: more complicated

Sometimes though, it's not so simple.  That's where functions that find patterns come in very useful.

```{r gender, echo=FALSE}
set.seed(4) # random sample below - make sure same every time
gender <- sample(c("Male", "mAle", "MaLe", "M", "MALE", "Ma", "FeMAle", "F", "Woman", "Man", "Fm", "FEMALE"), 1000, replace = TRUE)
```

```{r gentab}
table(gender)
```

# String functions

## Useful String Functions

Useful String functions

* `toupper()`, `tolower()` - uppercase or lowercase your data:
* `str_trim()` (in the `stringr` package) or `trimws` in base 
    - will trim whitespace
* `nchar` - get the number of characters in a string
* `paste()` - paste strings together with a space 
* `paste0` - paste strings together with no space as default

## Pasting strings with `paste` and `paste0`

Paste can be very useful for joining vectors together:

```{r Paste}
paste("Visit", 1:5, sep = "_")
paste("Visit", 1:5, sep = "_", collapse = " ")
paste("To", "is going be the ", "we go to the store!", sep = "day ")
# and paste0 can be even simpler see ?paste0 
paste0("Visit",1:5)
```

## Paste Depicting How Collapse Works

```{r Paste2}
paste(1:5)
paste(1:5, collapse = " ")
```

## The `stringr` package

Like `dplyr`, the `stringr` package:

* Makes some things more intuitive
* Is different than base R
* Is used on forums for answers
* Has a standard format for most functions
    * the first argument is a string like first argument is a `data.frame` in `dplyr`



## Splitting/Find/Replace and Regular Expressions

* R can do much more than find exact matches for a whole string
* Like Perl and other languages, it can use regular expressions.
* What are regular expressions?
    * Ways to search for specific strings 
    * Can be very complicated or simple
    * Highly Useful - think "Find" on steroids


## A bit on Regular Expressions

* http://www.regular-expressions.info/reference.html
* They can use to match a large number of strings in one statement
* `.` matches any single character
* `*` means repeat as many (even if 0) more times the last character
* `?` makes the last thing optional
* `^` matches start of vector `^a` - starts with "a"
* `$` matches end of vector `b$` - ends with "b"

## Substringing

`stringr`

* `str_sub(x, start, end)` - substrings from position start to position end
* `str_split(string, pattern)` - splits strings up - returns list!

## Splitting String: `stringr`

In base R, `strsplit` splits a vector on a string into a list
```{r strsplit_base}
x <- c("I really", "like writing", "R code programs")
y <- str_split(x, pattern = " ") # returns a list
y
```



## Using a fixed expression

One example case is when you want to split on a period "`.`".  In regular expressions `.` means **ANY** character, so

```{r}
str_split("I.like.strings", ".")
str_split("I.like.strings", fixed("."))
```

## Let's extract from `y`

The `purrr` package allows you to more easily interface with lists.

https://purrr.tidyverse.org/

The main function family for this is `map()`

## Let's extract from `y`

`map_chr()` takes a list and returns a character vector 

```{r stsplit2}
map_chr(y, first) # on the fly
map_chr(y, nth, 2) # on the fly
map_chr(y, last) # on the fly
```


## 'Find' functions: `stringr`

`str_detect`, `str_subset`, `str_replace`, and `str_replace_all` search for matches to argument pattern within each element of a character vector: they differ in the format of and amount of detail in the results. 

* `str_detect` - returns `TRUE` if `pattern` is found
* `str_subset` - returns only the strings which pattern were detected
    * convenient wrapper around `x[str_detect(x, pattern)]`
* `str_extract` - returns only strings which pattern were detected, but ONLY the pattern
* `str_replace` - replaces `pattern` with `replacement` the first time
* `str_replace_all` - replaces `pattern` with `replacement` as many times matched


## Let's look at modifier for `stringr`

`?modifiers`

* `fixed` - match everything exactly
* `regexp` - default - uses **reg**ular **exp**ressions
* `ignore_case` is an option to not have to use `tolower`



## 'Find' functions: Finding Indices

These are the indices where the pattern match occurs:

```{r alienMatch}
which(str_detect(ufo$comments, "two aliens"))
```

## 'Find' functions: Finding Logicals

These are the indices where the pattern match occurs:

```{r alienMatch_log}
str_detect(ufo$comments, "two aliens") %>% head()
```


## 'Find' functions: finding values, `stringr` and `dplyr` {.smaller}

```{r ggrep}
str_subset(ufo$comments, "two aliens")
ufo %>% filter(str_detect(comments, "two aliens"))
```

## Showing differnce in `str_extract`

`str_extract` extracts just the matched string

```{r ggrep2}
ss = str_extract(ufo$comments, "two aliens")
head(ss)
ss[ !is.na(ss)]
```



* Look for anycomment that starts with "aliens"
  
```{r grepstar}
str_subset(ufo$comments, "^aliens.*")
```

## Using Regular Expressions

That contains space then ship maybe with stuff in between
```{r grepstar2}
str_subset(ufo$comments, "space.?ship") %>% head(7)
```

## Replace

Let's say we wanted to sort the data set by latitude and longitude:

```{r classSal}
class(ufo$latitude)
```

```{r orderstring}
sort(c("1", "2", "10")) #  not sort correctly (order simply ranks the data)
order(c("1", "2", "10"))
```

## Replace

So we must change the coordinates into a numeric:
```{r destringSal}
head(ufo$latitude, 4)
head(as.numeric(ufo$latitude), 4)
```

## Dropping bad observations

```{r, dropIndex}
dropIndex = which(is.na(as.numeric(ufo$latitude)) | 
                      is.na(as.numeric(ufo$longitude)))
ufo_clean = ufo[-dropIndex,]
dim(ufo_clean)
```


## Ordering

```{r order}
ufo2 = ufo_clean
ufo2$latitude = as.numeric(ufo2$latitude)
ufo2$longitude = as.numeric(ufo2$longitude)
ufo2 <- ufo2[order(ufo2$latitude, ufo2$longitude), ] 
ufo2[1:5, c("datetime", "latitude", "longitude")]
```

## Replacing and subbing: `stringr` {.smaller}

We can do the same thing (with 2 piping operations!) in dplyr

```{r orderSal_stringr}
ufo_dplyr = ufo_clean
ufo_dplyr = ufo_dplyr %>% mutate( 
  latitude = latitude %>% as.numeric,
  longitude = longitude %>% as.numeric) %>% 
    arrange(latitude,longitude)
ufo_dplyr[1:5, c("datetime", "latitude", "longitude")]
```

## Special characters like money/$

```{r}
money = tibble(group = letters[1:5], 
  amount = c("$12.32", "$43.64", "$765.43", "$93.31", "$12.13"))
money %>% arrange(amount)
as.numeric(money$amount)
```

## Special characters like money/$

In the past, we would recommend just replacing the `$` sign with an empty string and convert to numeric:

```{r} 
money$amountNum = as.numeric(str_replace(money$amount, fixed("$"), ""))
money %>% arrange(amountNum)
```

## Special characters like money/$

But now there are better helper functions for this:

```{r}
money$amount = parse_number(money$amount)
```

This is way easier..."This drops any non-numeric characters before or after the first number. The grouping mark specified by the locale is ignored inside the number."

# Base R versions

## Substrings 

Base R

* `substr(x, start, stop)` - substrings from position start to position stop
* `strsplit(x, split)` - splits strings up - returns list!


## Splitting String: base R

In base R, `strsplit` splits a vector on a string into a list
```{r strsplit}
x <- c("I really", "like writing", "R code programs")
y <- strsplit(x, split = " ") # returns a list
y
```


## 'Find' functions: base R

`grep`: `grep`, `grepl`, `regexpr` and `gregexpr` search for matches to argument pattern within each element of a character vector: they differ in the format of and amount of detail in the results. 

`grep(pattern, x, fixed=FALSE)`, where:

* pattern = character string containing a regular expression to be matched in the given character vector.

* x = a character vector where matches are sought, or an object which can be coerced by as.character to a character vector.

* If fixed=TRUE, it will do exact matching for the phrase anywhere in the vector (regular find)

## 'Find' functions: stringr compared to base R

Base R does not use these functions.  Here is a "translator" of the `stringr` function to base R functions

* `str_detect` - similar to `grepl` (return logical)
* `grep(value = FALSE)` is similar to `which(str_detect())`
* `str_subset` - similar to `grep(value = TRUE)` - return value of matched
* `str_replace` - similar to `sub` - replace one time 
* `str_replace_all` - similar to `gsub` - replace many times


## Important Comparisons

Base R:

* Argument order is `(pattern, x)`
* Uses option `(fixed = TRUE)`

`stringr`

* Argument order is `(string, pattern)` aka `(x, pattern)`
* Uses function `fixed(pattern)`

