---
title: "Data Cleaning"
author: "Data Wrangling in R"
output:
  ioslides_presentation:
    css: styles.css
    widescreen: yes
  beamer_presentation: default
---


```{r, include = FALSE}
library(knitr)
library(tidyverse)
library(janitor)
opts_chunk$set(comment = "")
```

## Data Cleaning

In general, data cleaning is a process of investigating your data for inaccuracies, or recoding it in a way that makes it more manageable.

MOST IMPORTANT RULE - LOOK AT YOUR DATA!

## Useful checking functions

- `is.na` - is `TRUE` if the data is `FALSE` otherwise
- `!` - negation (NOT) 
    - if `is.na(x)` is `TRUE`, then `!is.na(x)` is `FALSE`

- `all` takes in a `logical` and will be `TRUE` if ALL are `TRUE`
    - `all(!is.na(x))` - are all values of `x` NOT NA
- `any` will be `TRUE` if ANY are true
    - `any(is.na(x))` - do we have any `NA`'s in `x`?
- `complete.cases()` - returns `TRUE` if EVERY value of a row is NOT `NA`
    - very stringent condition 
    - `FALSE` missing one value (even if not important)
 

## Read in the UFO dataset

* Read in data from RStudio Cloud or download from: http://sisbid.github.io/Module1/data/ufo/ufo_data_complete.csv.gz

```{r, importUfo}
ufo = read_csv("../data/ufo/ufo_data_complete.csv")
ufo = clean_names(ufo)
```

## Data cleaning "before" R

You saw warning messages when reading in this dataset. Let's just drop those rows for now

```{r}
p = problems(ufo)
ufo = ufo[-p$row,] # brackets can also be used for subsetting
```

## Checking for logical conditions {.smaller}
* `any()` - checks if there are any `TRUE`s
* `all()` - checks if ALL are true

```{r isna}
any(is.na(ufo$state)) # are there any NAs?
table(is.na(ufo$state)) # are there any NAs?
```

# Recoding Variables

## Example of Recoding: base R

For example, let's say gender was coded as Male, M, m, Female, F, f.  Using Excel to find all of these would be a matter of filtering and changing all by hand or using if statements.

In R, you can simply do something like:

```{r, eval = FALSE}
data$gender[data$gender %in% 
    c("Male", "M", "m")] <- "Male"
```

## Example of Cleaning: more complicated

Sometimes though, it's not so simple.  That's where functions that find patterns come in very useful.

```{r gender, echo=FALSE}
set.seed(4) # random sample below - make sure same every time
gender <- sample(c("Male", "mAle", "MaLe", "M", "MALE", "Ma", "FeMAle", "F", "Woman", "Man", "Fm", "FEMALE"), 1000, replace = TRUE)
```

```{r gentab}
table(gender)
```

# String functions

## Useful String Functions

Useful String functions

* `toupper()`, `tolower()` - uppercase or lowercase your data:
* `str_trim()` (in the `stringr` package) or `trimws` in base 
    - will trim whitespace
* `nchar` - get the number of characters in a string
* `paste()` - paste strings together with a space 
* `paste0` - paste strings together with no space as default

## Pasting strings with `paste` and `paste0`

Paste can be very useful for joining vectors together:

```{r Paste}
paste("Visit", 1:5, sep = "_")
paste("Visit", 1:5, sep = "_", collapse = " ")
paste("To", "is going be the ", "we go to the store!", sep = "day ")
# and paste0 can be even simpler see ?paste0 
paste0("Visit",1:5)
```

## Paste Depicting How Collapse Works

```{r Paste2}
paste(1:5)
paste(1:5, collapse = " ")
```

## The `stringr` package

Like `dplyr`, the `stringr` package:

* Makes some things more intuitive
* Is different than base R
* Is used on forums for answers
* Has a standard format for most functions
    * the first argument is a string like first argument is a `data.frame` in `dplyr`


## Substringing

`stringr`

* `str_sub(x, start, end)` - substrings from position start to position end
* `str_split(string, pattern)` - splits strings up - returns list! [we'll revisit in "Functional Programming"]

## Substringing

Examples: 

```{r}
str_sub("I like pizza", 8,12)
str_sub(c("Site A", "Site B", "Site C"), 6,6)
```

## Splitting/Find/Replace and Regular Expressions

* R can do much more than find exact matches for a whole string
* Like Perl and other languages, it can use regular expressions.
* What are regular expressions?
    * Ways to search for specific strings 
    * Can be very complicated or simple
    * Highly Useful - think "Find" on steroids


## A bit on Regular Expressions

* http://www.regular-expressions.info/reference.html
* They can use to match a large number of strings in one statement
* `.` matches any single character
* `*` means repeat as many (even if 0) more times the last character
* `?` makes the last thing optional
* `^` matches start of vector `^a` - starts with "a"
* `$` matches end of vector `b$` - ends with "b"

## 'Find' functions: `stringr`

`str_detect`, `str_subset`, `str_replace`, and `str_replace_all` search for matches to argument pattern within each element of a character vector: they differ in the format of and amount of detail in the results. 

* `str_detect` - returns `TRUE` if `pattern` is found
* `str_subset` - returns only the strings which pattern were detected
    * convenient wrapper around `x[str_detect(x, pattern)]`
* `str_extract` - returns only strings which pattern were detected, but ONLY the pattern
* `str_replace` - replaces `pattern` with `replacement` the first time
* `str_replace_all` - replaces `pattern` with `replacement` as many times matched


## Let's look at modifier for `stringr`

`?modifiers`

* `fixed` - match everything exactly
* `regexp` - default - uses **reg**ular **exp**ressions
* `ignore_case` is an option to not have to use `tolower`



## 'Find' functions: Finding Indices

These are the indices where the pattern match occurs:

```{r alienMatch}
which(str_detect(ufo$comments, "two aliens"))
```

## 'Find' functions: Finding Logicals

These are the indices where the pattern match occurs:

```{r alienMatch_log}
str_detect(ufo$comments, "two aliens") %>% head()
```


## 'Find' functions: finding values, `stringr` and `dplyr` {.smaller}

```{r ggrep}
str_subset(ufo$comments, "two aliens")
ufo %>% filter(str_detect(comments, "two aliens"))
```

## Showing differnce in `str_extract`

`str_extract` extracts just the matched string

```{r ggrep2}
ss = str_extract(ufo$comments, "two aliens")
head(ss)
ss[ !is.na(ss)]
```


* Look for any comment that starts with "aliens"
  
```{r grepstar}
str_subset(ufo$comments, "^aliens.*")
```

## Using Regular Expressions

That contains space then ship maybe with stuff in between
```{r grepstar2}
str_subset(ufo$comments, "space.?ship") %>% head(7)
```

## Replace

Let's say we wanted to sort the data set by latitude and longitude:

```{r classSal}
class(ufo$latitude)
```

```{r orderstring}
sort(c("1", "2", "10")) #  not sort correctly (order simply ranks the data)
order(c("1", "2", "10"))
```

## Replace

So we must change the coordinates into a numeric:
```{r destringSal}
head(ufo$latitude, 4)
head(as.numeric(ufo$latitude), 4)
```

## Dropping bad observations

```{r, dropIndex}
dropIndex = which(is.na(as.numeric(ufo$latitude)) | 
                      is.na(as.numeric(ufo$longitude)))
ufo_clean = ufo[-dropIndex,]
dim(ufo_clean)
```


## Replacing and subbing: `stringr` {.smaller}

We can do the same thing (with 2 piping operations!) in dplyr

```{r orderSal_stringr}
ufo_dplyr = ufo_clean
ufo_dplyr = ufo_dplyr %>% mutate( 
  latitude = latitude %>% as.numeric,
  longitude = longitude %>% as.numeric) %>% 
    arrange(latitude,longitude)
ufo_dplyr[1:5, c("datetime", "latitude", "longitude")]
```

## Special characters

```{r}
money = tibble(group = letters[1:5], 
  amount = c("$12.32", "$43.64", "$765.43", "$93.31", "$12.13"))
money %>% arrange(amount)
as.numeric(money$amount)
```

## Special characters

In the past, we would recommend just replacing the `$` sign with an empty string and convert to numeric:

```{r} 
money$amountNum = as.numeric(str_replace(money$amount, fixed("$"), ""))
money %>% arrange(amountNum)
```

## Special characters

But now there are better helper functions for this:

```{r}
money$amount = parse_number(money$amount)
money %>% arrange(amount)

```

## Special characters

Also works for internal commas:

```{r}
parse_number(c("12,123,123.00", "12,465.10"))
```

## Dates and times

The `lubridate` package is amazing, there's no reason to use anything else.

https://lubridate.tidyverse.org/

```{r}
library(lubridate)
ufo$timestamp = mdy_hm(ufo$datetime)
ufo$date_posted = mdy(ufo$date_posted)
head(ufo$timestamp)
```
