---
title: "Functional Programming"
author: "Data Wrangling in R"
output:
  ioslides_presentation:
    css: styles.css
    widescreen: yes
  beamer_presentation: default
---

```{r, include=FALSE}
library(tidyverse)
library(repurrrsive)
library(tidyr)
```

## Functional Programming

Allows you to flexibly iterate functions to each element of a list or vector

* https://purrr.tidyverse.org/

* https://github.com/rstudio/cheatsheets/raw/master/purrr.pdf

## Examples we will use

* https://jennybc.github.io/purrr-tutorial/

* https://cran.r-project.org/web/packages/repurrrsive/index.html

* https://tidyr.tidyverse.org/articles/rectangle.html

## Why do this at all?

https://jennybc.github.io/purrr-tutorial/bk01_base-functions.html

You need a way to iterate in R in a data-structure-informed way. What does that mean?

* Iterate over elements of a list
* Iterate over rows or columns of a 2-dimensional object
* Iterate over sub data frames induced by one or more factors
* Iterate over tuples formed from the i-th element of several vectors of equal length

## What is a 'list'?

* Lists are the most flexible/"generic" data class in R
* Can be created using list()
* Can hold vectors, strings, matrices, models, list of other list, lists upon lists!
* Can reference data using $ (if the elements are named), or using [], or [[]]

```{r makeList, comment="", prompt=TRUE}
mylist <- list(letters=c("A", "b", "c"), 
        numbers=1:3, matrix(1:25, ncol=5))
```

## List Structure

```{r Lists, comment="", prompt=TRUE}
head(mylist)
```

## List referencing

```{r Listsref1, comment="", prompt=TRUE}
mylist[1] # returns a list
mylist["letters"] # returns a list
```

## List referencing
  
```{r Listsrefvec, comment="", prompt=TRUE}  
mylist[[1]] # returns the vector 'letters'
mylist$letters # returns vector
mylist[["letters"]] # returns the vector 'letters'
```

## List referencing

You can also select multiple lists with the single brackets. 

```{r Listsref2, comment="", prompt=TRUE}
mylist[1:2] # returns a list
```

## List referencing

You can also select down several levels of a list at once

```{r Listsref3, comment="", prompt=TRUE}
mylist$letters[1]
mylist[[2]][1]
mylist[[3]][1:2,1:2]
```

## How would I encounter lists?

This comes up a lot in data cleaning (although many tasks can be accomplished by `separate`)

```{r} 
h = c("I like performing", "much data wrangling in R", "it is oh so fun")
ll = str_split(h, " ")
ll
```

## Why do this at all?

```{r}
class(got_chars)
length(got_chars)
lengths(got_chars)
names(got_chars[[1]])
```



## How would we get the names of each GoT character?

The really old way involved `for` loops:

```{r}
char_names = vector("character", length(got_chars)) # initiate
for(i in seq(along=got_chars)) { # iterate
    char_names[i] = got_chars[[i]]$name
}
char_names[1:5] # examine
```


## How would we get the names of each GoT character?

The kinda old way involved `apply` statements:

```{r}
sapply(got_chars[1:5], function(x) x$name)
```


## How would we get the names of each GoT character?

This can still be pretty efficient:

```{r}
sapply(got_chars[1:5], "[[", "name")
```

The user interface of the “apply” functions is not as consistent as it could be, which slows down learning. The return objects frequently require further checking and massage to use downstream. In particular, there’s a tendency to return a vector (atomic or otherwise) or array, instead of data frame, with the original factor levels appearing in a names attribute. [https://jennybc.github.io/purrr-tutorial/bk01_base-functions.html]


## How would we get the names of each GoT character?

This can still be pretty efficient:

```{r}
map_chr(got_chars[1:5], "name")
```

## `sapply` versus `map_chr`

Potentially confusing output:
```{r}
sapply(got_chars[2:3], "[[", "aliases")
```

Compared to error: 
```{r,eval=FALSE}
map_chr(got_chars[2:3], "aliases")
Error: Result 1 must be a single string, not a character vector of length 11

```

## Example using string split output

```{r}
map_chr(ll, first)
map_chr(ll, last)
map_chr(ll, nth, 3)
```


## More extensive examples

You can create tibbles where each observation is a list:

https://tidyr.tidyverse.org/articles/rectangle.html

```{r}
chars <- tibble(char = got_chars)
chars
```

## More extensive examples

```{r}
chars2 <- chars %>% unnest_wider(char)
chars2
```

## Say you wanted all characters and their titles:

```{r}
chars2 %>%
  select(name, title = titles) %>%
  unnest_longer(title)
```
