---
title: "Manipulating Data in R"
author: "Andrew Jaffe"
output:
  ioslides_presentation:
    css: styles.css
  beamer_presentation: default
---

```{r, echo = FALSE, include=FALSE}
library(knitr)
library(readr)
opts_chunk$set(comment = "")
library(tidyverse)
```


## Reshaping Data 

```{r, echo = FALSE, message=FALSE}
library(tidyverse)
```

In this module, we will show you how to:

1. Reshaping data from wide (fat) to long (tall)
2. Reshaping data from long (tall) to wide (fat)

## Setup

We will show you how to do each operation in base R then show you how to use the `dplyr` or `tidyr` package to do the same operation (if applicable).  

See the "Data Wrangling Cheat Sheet using `dplyr` and `tidyr`":

* https://www.rstudio.com/wp-content/uploads/2015/02/data-wrangling-cheatsheet.pdf

## What is wide/long data?


```{r, echo = FALSE}
ex_wide = tibble(id = 1:2,
                     visit1 = c(10, 5),
                     visit2 = c(4, 6),
                     visit3 = c(3, NA)
                     )
ex_long = tibble(id = c(rep(1, 3), rep(2, 2)),
                     visit = c(1:3, 1:2),
                     value = c(10, 4, 3, 5, 6))
```

See http://www.cookbook-r.com/Manipulating_data/Converting_data_between_wide_and_long_format/

- Wide - multiple columns per observation 
    - e.g. visit1, visit2, visit3
```{r, echo = FALSE}
ex_wide
```

- Long - multiple rows per observation
```{r, echo = FALSE}
ex_long
```

## What is wide/long data?

More accurately, data is wide or long **with respect** to certain variables.

## Data used: Charm City Circulator

```{r, message = FALSE}
circ = read_csv("../data/Charm_City_Circulator_Ridership.csv")
head(circ, 2)
class(circ$date)
```

## Creating a Date class from a character date

```{r, message = FALSE}
library(lubridate) # great for dates!
```

```{r, message= FALSE}
sum(is.na(circ$date))
sum( circ$date == "")
circ = mutate(circ, date = mdy(date))
sum( is.na(circ$date) ) # all converted correctly
head(circ$date, 3)
class(circ$date)
```

## Reshaping data from wide (fat) to long (tall): base R

The `reshape` command exists.  It is a **confusing** function.  Don't use it.

## tidyr package

`tidyr` allows you to "tidy" your data.  We will be talking about:

- `gather` - make multiple columns into variables, (wide to long)
- `spread` - make a variable into multiple columns, (long to wide)
- `separate` - string into multiple columns
- `unite` - multiple columns into one string


## Reshaping data from wide (fat) to long (tall): tidyr 

`tidyr::gather` - puts column data into rows.

We want the column names into "`var`" variable in the output dataset and the value in "`number`" variable. We then describe which columns we want to "gather:" 

```{r}
long = gather(circ, key = "var", value = "number", 
              -day, -date, -daily)
head(long, 4)
```


## Reshaping data from wide (fat) to long (tall): tidyr 

- Could be explicit on what we want to gather

```{r}
long = gather(circ, key = "var", value = "number", 
              starts_with("orange"), starts_with("purple"),
              starts_with("green"), starts_with("banner"))
long
```

## Reshaping data from wide (fat) to long (tall): tidyr 

```{r}
long %>% count(var)
```


## Lab Part 1

[Website](https://sisbid.github.io/Module1/)


## Making a separator

We will use `str_replace` from `stringr` to put periods in the names (periods are **not** special when in a replacement)

```{r}
long = long %>% mutate(
  var = var %>% 
    str_replace("Board", ".Board") %>% 
    str_replace("Alight", ".Alight") %>% 
    str_replace("Average", ".Average") 
)
long %>% count(var)
```

## Reshaping data from wide (fat) to long (tall): tidyr

Now each `var` is boardings, averages, or alightings.  We want to separate these so we can have these by line. Remember "." is special character:

```{r}
long = separate(long, var, into = c("line", "type"), 
                 sep = "[.]")
head(long, 2)
unique(long$line)
unique(long$type)
```

## Re-uniting all the lines

If we had the opposite problem, we could use the `unite` function:

```{r}
reunited = long %>% 
  unite(col = var, line, type, sep = ".")  
reunited %>% select(day, var) %>% head(3) %>% print
```
We could also use `paste`/`paste0`.


<!-- ## Making column names a little more separated -->

<!-- Alternative: We could have replaced the column names first **then** reshaped: -->

<!-- ```{r, eval = FALSE} -->
<!-- cn = colnames(circ) -->
<!-- cn = cn %>%  -->
<!--   str_replace("Board", ".Board") %>%  -->
<!--   str_replace("Alight", ".Alight") %>%  -->
<!--   str_replace("Average", ".Average")  -->
<!-- colnames(circ) = cn # then reshape using gather! -->
<!-- ``` -->

<!-- or (with some string working (see data cleaning)): -->

<!-- ```{r, eval = FALSE} -->
<!-- circ = circ %>%  -->
<!--   rename_all(.funs = str_replace, -->
<!--             pattern = "(orange|purple|green|banner)", -->
<!--             replacement = "\\1.") -->
<!-- ``` -->

## Lab Part 2

[Website](https://sisbid.github.io/Module1/)


## Reshaping data from long (tall) to wide (fat): tidyr

In `tidyr`, the `spread` function spreads rows into columns.  Now we have a long data set, but we want to separate the Average, Alightings and Boardings into different columns:

```{r}
# have to remove missing days
wide = long %>% filter(!is.na(date))
wide = wide %>% spread(type, number)
head(wide)
```


## Lab Part 3

[Website](https://sisbid.github.io/Module1/)
