---
title: "Manipulating Data in R"
author: "Andrew Jaffe"
date: "July 13, 2016"
output:
  ioslides_presentation:
    css: styles.css
  beamer_presentation: default
---

```{r, echo = FALSE}
library(knitr)
opts_chunk$set(comment = "")
```


## Reshaping Data 

In this module, we will show you how to:

1. Reshaping data from long (tall) to wide (fat)
2. Reshaping data from wide (fat) to long (tall)

## Setup

We will show you how to do each operation in base R then show you how to use the `dplyr` or `tidyr` package to do the same operation (if applicable).  

See the "Data Wrangling Cheat Sheet using `dplyr` and `tidyr`":

* https://www.rstudio.com/wp-content/uploads/2015/02/data-wrangling-cheatsheet.pdf

## Data used: Charm City Circulator {.smaller}

http://www.aejaffe.com/summerR_2016/data/Charm_City_Circulator_Ridership.csv

```{r}
circ = read.csv("../data/Charm_City_Circulator_Ridership.csv", 
                as.is = TRUE)
head(circ, 2)
```

## Creating a Date class from a character date

```{r, message= FALSE}
library(lubridate) # great for dates!
library(dplyr) # mutate/summarise functions
circ = mutate(circ, date = mdy(date))
sum( is.na(circ$date) ) # all converted correctly
head(circ$date)
class(circ$date)
```

## Making column names a little more separated

We will use `str_replace` from `stringr` to put periods in the column names.

```{r}
library(stringr)
cn = colnames(circ)
cn = cn %>% 
  str_replace("Board", ".Board") %>% 
  str_replace("Alight", ".Alight") %>% 
  str_replace("Average", ".Average") 
colnames(circ) = cn
cn
```



```{r, echo = FALSE}
ex_wide = data.frame(id = 1:2,
                     visit1 = c(10, 5),
                     visit2 = c(4, 6),
                     visit3 = c(3, NA)
                     )
ex_long = data.frame(id = c(rep(1, 3), rep(2, 2)),
                     visit = c(1:3, 1:2),
                     value = c(10, 4, 3, 5, 6))
```

## Reshaping data from wide (fat) to long (tall)

See http://www.cookbook-r.com/Manipulating_data/Converting_data_between_wide_and_long_format/

- Wide - multiple columns per observation 
    - e.g. visit1, visit2, visit3
```{r, echo = FALSE}
ex_wide
```

- Long - multiple rows per observation
```{r, echo = FALSE}
ex_long
```



## Reshaping data from wide (fat) to long (tall): base R

The `reshape` command exists.  It is a **confusing** function.  Don't use it.

## Reshaping data from wide (fat) to long (tall): tidyr {.smaller}

`tidyr::gather` - puts column data into rows.

We want the column names into "`var`" variable in the output dataset and the value in "`number`" variable. We then describe which columns we want to "gather:" 
```{r}
library(tidyr)
long = gather(circ, key = "var", value = "number", 
              starts_with("orange"),
              starts_with("purple"), 
              starts_with("green"),
              starts_with("banner"))
head(long, 2)
table(long$var)
```

## Reshaping data from wide (fat) to long (tall): tidyr

Now each `var` is boardings, averages, or alightings.  We want to separate these so we can have these by line.

```{r}
long = separate_(long, "var", 
                 into = c("line", "type"), 
                 sep = "[.]")
head(long, 3)
unique(long$line)
unique(long$type)
```

## Finding the First (or Last) record


```{r}
long = long %>% filter(!is.na(number) & number > 0)
first_and_last = long %>% arrange(date) %>% # arrange by date
  filter(type %in% "Boardings") %>% # keep boardings only
  group_by(line) %>% # group by line
  slice( c(1, n())) # select ("slice") first and last (n() command) lines
first_and_last %>%  head(4)
```

## Reshaping data from long (tall) to wide (fat): tidyr

In `tidyr`, the `spread` function spreads rows into columns.  Now we have a long data set, but we want to separate the Average, Alightings and Boardings into different columns:

```{r}
# have to remove missing days
wide = filter(long, !is.na(date))
wide = spread(wide, type, number)
head(wide)
```


## Reshaping data from long (tall) to wide (fat): tidyr

We can use `rowSums` to see if any values in the row is `NA` and keep if the row, which is a combination of date and line type has any non-missing data.

```{r}
# wide = wide %>%
#     select(Alightings, Average, Boardings) %>%
#     mutate(good = rowSums(is.na(.)) > 0)
namat = !is.na(select(wide, Alightings, Average, Boardings))
head(namat)
wide$good = rowSums(namat) > 0
head(wide, 3)
```

## Reshaping data from long (tall) to wide (fat): tidyr

Now we can filter only the good rows and delete the `good` column.
```{r}
wide = filter(wide, good) %>% select(-good)
head(wide)
```
