---
title: "Manipulating Data in R"
author: "Introduction to R for Public Health Researchers"
output:
  ioslides_presentation:
    css: styles.css
    widescreen: yes
---


```{r, echo = FALSE, include=FALSE}
library(knitr)
opts_chunk$set(comment = "")
library(tidyverse)
```

# The `merge` function exists, don't use it

## Joining in `dplyr`

* Merging/joining data sets together - usually on key variables, usually "id"
* `?join` - see different types of joining for `dplyr`
* Let's look at https://www.rstudio.com/wp-content/uploads/2015/02/data-wrangling-cheatsheet.pdf
* `inner_join(x, y)` - only rows that match for `x` and `y` are kept
* `full_join(x, y)` - all rows of `x` and `y` are kept 
* `left_join(x, y)` - all rows of `x` are kept even if not merged with `y`
* `right_join(x, y)` - all rows of `y` are kept even if not merged with `x`
* `anti_join(x, y)` - all rows from `x` not in `y` keeping just columns from `x`.


## Merging: Simple Data

`base` has baseline data for ids `1` to `10` and Age

```{r merging}
base <- tibble(id = 1:10, Age = seq(55,60, length=10))
head(base, 2)
```
`visits` has ids 1 to 8, then 11 (new id), and 3 visits and outcome

```{r}
visits <- tibble(id = c(rep(1:8, 3), 11), visit= c(rep(1:3, 8), 3),
                    Outcome = seq(10,50, length=25))
tail(visits, 2)
```


## Inner Join

```{r inner_join}
ij = inner_join(base, visits)
dim(ij)
tail(ij)
```


## Left Join

```{r left_join}
lj = left_join(base, visits)
dim(lj)
tail(lj)
```

## Right Join

```{r right_join}
rj = right_join(base, visits)
tail(rj, 3)
```

## Right Join: Switching arguments

```{r right_join2}
rj2 = right_join(visits, base)
tail(rj2, 3)
```
```{r right_join_arrange, echo = FALSE}
rj2 = arrange(rj2, id, visit) %>% select(id, visit, Outcome, Age)
lj = arrange(lj, id, visit) %>% select(id, visit, Outcome, Age)
```
```{r right_join_arrange_out}
identical(rj2, lj) ## after some rearranging
```

## Full Join

```{r full_join}
fj = full_join(base, visits)
tail(fj, 4)
```

## Logging the joins

The `tidylog` package can show you log outputs from `dplyr` (newly added).  You will need to install to use.

```{r, include=FALSE}
library(tidylog)
```

```{r left_join_log}
library(tidylog)
left_join(base, visits)
```

```{r include=FALSE}
unloadNamespace("tidylog")
```
## Using the `by` argument

By default - uses intersection of column names.  If `by` specified, then uses that, but if other columns with same name, adds `suffix`.

```{r use_by}
base = base %>% mutate(x = 5)
visits = visits %>% mutate(x = 4)
head(full_join(base, visits))
```

## Using the `by` argument

```{r use_by_suffix}
head(full_join(base, visits, by = "id"))
head(full_join(base, visits, by = "id", suffix = c("_base", "_visit")))
```

## Using the `by` argument if column names different

```{r use_by_diff}
base = base %>% 
  select(-x) %>% 
  mutate(myvar = 4)
visits = visits %>% 
  select(-x) %>% 
  mutate(MyVar = 4)
full_join(base, visits, by = c("id", "myvar" = "MyVar"))
```
