---
title: "Subsetting Data in R"
author: "Andrew Jaffe"
date: "July 11, 2016"
output:
  ioslides_presentation:
    css: styles.css
  beamer_presentation: default
---

```{r, echo = FALSE}
library(knitr)
opts_chunk$set(comment = "")
```

## Overview

We showed one way to read data into R using *read.csv*.  In this module, we will show you how to:

1. Select specific elements of an object by an index or logical condition
2. Renaming columns of a `data.frame`
3. Subset rows of a `data.frame`
4. Subset columns of a `data.frame`
5. Add/remove new columns to a `data.frame`
6. Order the columns of a `data.frame`
7. Order the rows of a `data.frame`

## Setup

We will show you how to do each operation in base R then show you how to use the `dplyr` package to do the same operation (if applicable).  

Many resources on how to use `dplyr` exist and are straightforward: 

* https://cran.rstudio.com/web/packages/dplyr/vignettes/
* https://stat545-ubc.github.io/block009_dplyr-intro.html
* https://www.datacamp.com/courses/dplyr-data-manipulation-r-tutorial


## Select specific elements using an index

Often you only want to look at subsets of a data set at any given time. As a review, elements of an R object are selected using the brackets (`[` and `]`).

For example, `x` is a vector of numbers and we can select the second element of `x` using the brackets and an index (2):

```{r}
x = c(1, 4, 2, 8, 10)
x[2]
```

## Select specific elements using an index

We can select the fifth or second AND fifth elements below:

```{r}
x = c(1, 2, 4, 8, 10)
x[5]
x[c(2,5)]
```

## Subsetting by deletion of entries

You can put a minus (`-`) before integers inside brackets to remove these indices from the data.

```{r negativeIndex}
x[-2] # all but the second
```
Note that you have to be careful with this syntax when dropping more than 1 element:

```{r negativeIndex2}
x[-c(1,2,3)] # drop first 3
# x[-1:3] # shorthand. R sees as -1 to 3
x[-(1:3)] # needs parentheses
```



## Select specific elements using logical operators

What about selecting rows based on the values of two variables? We use logical statements.  Here we select only elements of `x` greater than 2:

```{r}
x
x > 2
x[ x > 2 ]
```

## Select specific elements using logical operators
You can have multiple logical conditions using the following:

* `&` : AND
* `|` : OR

```{r}
x[ x > 2 & x < 5 ]
x[ x > 5 | x == 2 ]
```


## which function

The `which` functions takes in logical vectors and returns the index for the elements where the logical value is `TRUE`.  

```{r}
which(x > 5 | x == 2) # returns index
x[ which(x > 5 | x == 2) ]
x[ x > 5 | x == 2 ]
```


## Creating a `data.frame` to work with

Here we create a toy data.frame named `df` using random data:
```{r}
set.seed(2016) # reproducbility
df = data.frame(x = c(1, 2, 4, 10, 10),
                x2 = rpois(5, 10),
                y = rnorm(5),
                z = rpois(5, 6)
                )
```


# Renaming Columns

## Renaming Columns of a `data.frame`: base R

We can use the `colnames` function to directly reassign column names of `df`:
```{r}
colnames(df) = c("x", "X", "y", "z")
head(df)
colnames(df) = c("x", "x2", "y", "z") #reset
```

## Renaming Columns of a `data.frame`: base R

We can assign the column names, change the ones we want, and then re-assign
the column names:
```{r}
cn = colnames(df)
cn[ cn == "x2"] = "X"
colnames(df) = cn
head(df)
colnames(df) = c("x", "x2", "y", "z") #reset
```

## Renaming Columns of a `data.frame`: dplyr


```{r}
library(dplyr)
```

Note, when loading `dplyr`, it says objects can be "masked".  That means if you 
use a function defined in 2 places, it uses the one that is loaded in **last**.

## Renaming Columns of a `data.frame`: dplyr

For example, if we print `filter`, then we see at the bottom `namespace:dplyr`, which
means when you type `filter`, it will use the one from the `dplyr` package.

```{r}
filter
```
## Renaming Columns of a `data.frame`: dplyr

A `filter` function exists by default in the `stats` package, however.  If you want
to make sure you use that one, you use `PackageName::Function` with the colon-colon 
("`::`") operator. 

```{r}
head(stats::filter,2)
```

This is important when loading many packages, and you may have
some conflicts/masking:

## Renaming Columns of a `data.frame`: dplyr

To rename columns in `dplyr`, you use the `rename` command

```{r}
df = dplyr::rename(df, X = x2)
head(df)
df = dplyr::rename(df, x2 = X) # reset
```


# Subsetting Columns


## Subset columns of a `data.frame`: 

We can grab the `x` column using the `$` operator. 

```{r}
df$x
```

## Subset columns of a `data.frame`: 

We can also subset a `data.frame` using the bracket `[, ]` subsetting. 

For `data.frame`s and matrices (2-dimensional objects), the brackets are `[rows, columns]` subsetting.  We can grab the `x` column using the index of the column or the column name ("`x`") 

```{r}
df[, 1]
df[, "x"]
```

## Subset columns of a `data.frame`: 

We can select multiple columns using multiple column names:

```{r}
df[, c("x", "y")]
```


## Subset columns of a `data.frame`: dplyr

The `select` command from `dplyr` allows you to subset
```{r}
select(df, x)
```

## Select columns of a `data.frame`: dplyr

The `select` command from `dplyr` allows you to subset columns of 
```{r}
select(df, x, x2)
select(df, starts_with("x"))
```


# Subsetting Rows

## Subset rows of a `data.frame` with indices: 

Let's select **rows** 1 and 3 from `df` using brackets:

```{r}
df[ c(1, 3), ]
```


## Subset rows of a `data.frame`: 

Let's select the rows of `df` where the `x` column is greater than 5 or is equal to 2.  Without any index for columns, all columns are returned:

```{r}
df[ df$x > 5 | df$x == 2, ]
```

## Subset rows of a `data.frame`: 

We can subset both rows and colums at the same time:

```{r}
df[ df$x > 5 | df$x == 2, c("y", "z")]
```


## Subset rows of a `data.frame`: dplyr

The command in `dplyr` for subsetting rows is `filter`.  Try `?filter`

```{r}
filter(df, x > 5 | x == 2)
```
Note, no `$` or subsetting is necessary.  R "knows" `x` refers to a column of `df`.

## Subset rows of a `data.frame`: dplyr

By default, you can separate conditions by commas, and `filter` assumes these statements are joined by `&`

```{r}
filter(df, x > 2 & y < 0)
filter(df, x > 2, y < 0)
```


## Combining `filter` and `select`

You can combine `filter` and `select` to subset the rows and columns, respectively, of a `data.frame`:

```{r}
select(filter(df, x > 2 & y < 0), y, z)
```

In `R`, the common way to perform multiple operations is to wrap functions around each other in a nested way such as above 

## Assigning Temporary Objects

One can also create temporary objects and reassign them:

```{r}
df2 = filter(df, x > 2 & y < 0)
df2 = select(df2, y, z)
```

## Piping - a new concept

There is another (newer) way of performing these operations, called "piping".  It is becoming more popular as it's easier to read:

```{r}
df %>% filter(x > 2 & y < 0) %>% select(y, z)
```

It is read: "take df, then filter the rows and then select `y`, `z`".

# Adding/Removing Columns

## Adding new columns to a `data.frame`: base R

You can add a new column, called `newcol` to `df`, using the `$` operator:
```{r}
df$newcol = 5:1
df$newcol = df$x + 2
```



## Removing columns to a `data.frame`: base R

You can remove a column by assigning to `NULL`:
```{r}
df$newcol = NULL
```
or selecing only the columns that were not `newcol`:
```{r}
df = df[, colnames(df) != "newcol"]
```

## Adding new columns to a `data.frame`: base R

You can also "**c**olumn **bind**" a `data.frame` with a vector (or series of vectors), using the `cbind` command:
```{r}
cbind(df, newcol = 5:1)
```


## Adding columns to a `data.frame`: dplyr

The `mutate` function in `dplyr` allows you to add or replace columns of a `data.frame`:
```{r}
mutate(df, newcol = 5:1)
print({df = mutate(df, newcol = x + 2)})
```

## Removing columns to a `data.frame`: dplyr

The `NULL` method is still very common.

The `select` function can remove a column with a minus (`-`), much like removing rows:
```{r}
select(df, -newcol)
```

## Removing columns to a `data.frame`: dplyr

Remove `newcol` and `y`
```{r}
select(df, -one_of("newcol", "y"))
```

# Ordering columns

## Ordering the columns of a `data.frame`: base R

We can use the `colnames` function to get the column names of `df` and then put `newcol` first by subsetting `df` using brackets:

```{r}
cn = colnames(df)
df[, c("newcol", cn[cn != "newcol"]) ]
```

## Ordering the columns of a `data.frame`: dplyr

The `select` function can reorder columns.  Put `newcol` first, then select the rest of columns:
```{r}
select(df, newcol, everything())
```

# Ordering rows

## Ordering the rows of a `data.frame`: base R

We use the `order` function on a vector or set of vectors, in increasing order:

```{r}
df[ order(df$x), ]
```

## Ordering the rows of a `data.frame`: base R

The `decreasing` argument will order it in decreasing order:
```{r}
df[ order(df$x, decreasing = TRUE), ]
```

## Ordering the rows of a `data.frame`: base R

You can pass multiple vectors, and must use the negative (using `-`) to mix decreasing and increasing orderings (sort increasing on `x` and decreasing on `y`):
```{r}
df[ order(df$x, -df$y), ]
```

## Ordering the rows of a `data.frame`: dplyr

The `arrange` function can reorder rows  By default, `arrange` orders in ascending order:
```{r}
arrange(df, x)
```

## Ordering the rows of a `data.frame`: dplyr
Use the `desc` to arrange the rows in descending order:
```{r}
arrange(df, desc(x))
```

## Ordering the rows of a `data.frame`: dplyr
It is a bit more straightforward to mix increasing and decreasing orderings:
```{r}
arrange(df, x, desc(y))
```

## Transmutation

The `transmute` function in `dplyr` combines both the `mutate` and `select` functions. One can create new columns and keep the only the columns wanted:

```{r}
transmute(df, newcol2 = x * 3, x, y)
```
