---
title: "Data Summarization"
author: "Data Wrangling in R"
output:
  ioslides_presentation:
    css: ../styles/styles.css
    widescreen: yes
---


```{r, echo = FALSE, message=FALSE, error = FALSE}
library(knitr)
opts_chunk$set(comment = "", message = FALSE)
suppressWarnings({library(dplyr)})
library(readr)
library(tidyverse)
```

<style type="text/css">
pre { /* Code block - slightly smaller in this lecture */
    font-size: 20px;
}
</style>

## Data Summarization

* Basic statistical summarization
    * `mean(x)`: takes the mean of x
    * `sd(x)`: takes the standard deviation of x
    * `median(x)`: takes the median of x
    * `quantile(x)`: displays sample quantiles of x. Default is min, IQR, max
    * `range(x)`: displays the range. Same as `c(min(x), max(x))`
    * `sum(x)`: sum of x
    * `max(x)`: maximum value in x
    * `min(x)`: minimum value in x
    
* **all have the ** `na.rm =` **argument for missing data**

## Statistical summarization

These functions work on **vectors**:

```{r}
x <- c(1, 5, 7, 4, 2, 8)
mean(x)
```

Summarization on a `data.frame`/`tibble`:

```{r}
mtcars %>% pull(hp) %>% mean()
```

## GUT CHECK!

What kind of object do we need to run summary operators like `mean()` ?

A. A vector of numbers

B. A vector of characters

C. A dataset

## UFO Data

Let's use the UFO data again:

```{r warning=FALSE}
ufo <- read_csv("https://sisbid.github.io/Data-Wrangling/data/ufo/ufo_data_complete.csv")
ufo <- ufo %>% rename(duration_s = `duration (seconds)`)
head(ufo)
```

## Column to vector

Summarize a column as a vector with `pull()`.

```{r, message = FALSE}
ufo %>% pull(duration_s) %>% mean()
ufo %>% pull(duration_s) %>% mean(na.rm=TRUE)
ufo %>% pull(duration_s) %>% median(na.rm=TRUE)
```

## `dplyr`: `count` 

`count`: the number of rows in each group.

```{r, message = FALSE}
ufo %>% count(shape)
```

## `dplyr`: `count` 

Multiple columns listed further subdivides the count.

```{r, message = FALSE}
ufo %>% count(shape, country)
```

## `dplyr`: `count` 

Option to sort the results with `sort = TRUE`

```{r, message = FALSE}
ufo %>% count(shape, sort = TRUE)
```

## GUT CHECK!

The `count()` function can help us tally:

A. Sample size

B. Rows per each category

C. How many categories

## `dplyr`: `count` 

Instead of counting the **number of rows** in each group, `wt` computes sum() of a column for each group.

```{r, message = FALSE}
# Add up "duration_s" for each "shape" category
ufo %>% count(shape, wt = duration_s)
```

# Grouping

## Perform Operations By Groups: dplyr

Regular data...

```{r}
# No group_by()
ufo
```

## Perform Operations By Groups: dplyr

`group_by` allows you group the data set by variables/columns you specify:

```{r}
ufo_grouped <- ufo %>% group_by(shape)
ufo_grouped
```

## Summarize the grouped data

We can summarize grouped data using `summarize()`:

```{r}
ufo_grouped %>% 
  summarize(mean_value = mean(duration_s, na.rm = TRUE))
```

## Use the `pipe` to string these together!

Pipe `ufo` into `group_by`, then pipe that into `summarize`:

```{r}
ufo %>%
  group_by(shape) %>%
  summarize(mean_value = mean(duration_s, na.rm = TRUE),
            max_value = max(duration_s, na.rm = TRUE))
```

## `group_by` with `mutate` - Useful for comparisons {.codesmall}

Use `group_by` to calculate the mean value for each shape. We can use `mutate` to add it as a column.

```{r}
ufo_compare <- ufo %>%
  group_by(shape) %>%
  mutate(duration_shape_avg = mean(duration_s, na.rm = TRUE)) %>%
  select(shape, duration_s, duration_shape_avg)
ufo_compare
```

## `group_by` with `mutate` - Useful for comparisons {.codesmall}

Create a "difference" variable:

```{r}
ufo_compare %>% mutate(diff = duration_s - duration_shape_avg)
```

# Iterative summaries

## Iterative summaries: `dplyr` `summarize()` and `across()` functions {.codesmall}

Use the [`across`](https://dplyr.tidyverse.org/reference/across.html) function with `summarize()` to summarize across multiple columns of your data.

```{r}
dropouts <- read_delim("https://sisbid.github.io/Data-Wrangling/data/dropouts.txt", delim = "/")
dropouts %>%
  group_by(ETHNIC) %>%
  summarize(across(c(D9, D10, D11, D12), ~sum(.x)))
```

## Select different columns based on the data class

Use `?tidyr_tidy_select` functions like `where(is.numeric)`!

```{r}
dropouts %>% 
  group_by(ETHNIC) %>% 
  summarize(across( where(is.numeric), ~ sum(.x, na.rm = TRUE)))
```

## `summary()` Function

Using `summary()` can give you rough snapshots of each numeric column (character columns are skipped):

```{r}
summary(dropouts)
```

## Summary

- summary stats (`mean()`) work with `pull()`
- `count(x)`: what unique values do you have?
- `group_by()`: changes all subsequent functions
  - combine with `summarize()` to get statistics per group
  - combine with `across()` to select several columns
- `summary(x)`: quantile information

https://sisbid.github.io/Data-Wrangling/08_Data_Summarization/lab/data-summarization-lab.Rmd
