---
title: "Data Cleaning Part 2 Lab"
output: html_document
---

```{r}
library(tidyverse)
```


1. Read in the UFO dataset (used in the Data IO lectures) as an R object called `ufo`.
You can read directly from the web here: https://raw.githubusercontent.com/SISBID/Module1/gh-pages/data/ufo/ufo_data_complete.csv . 
You can ignore the "problems" with some rows. 

```{r}
library(readr)
ufo <- read_csv("https://raw.githubusercontent.com/SISBID/Module1/gh-pages/data/ufo/ufo_data_complete.csv")
```

Clean up the column/variable names of the `ufo` dataset to remove spaces and non-alphanumeric characters. You can use the `dplyr::rename()` function or use the `janitor::clean_names()` function. 

```{r}
library(janitor)
ufo <- clean_names(ufo)
```

2. How might one check how many UFO sightings were reported on a time scale of hours, specifically using the `duration (hours/min)` originally-named column? For sake of ease, assume it is only coded as "hour" and no other way.

```{r}

```

3. How many of the  minutes-scale observations (assuming "min" is the pattern to look for) have durations greater than 14400 seconds (or 4 hours)? 

```{r}

```

4. How many ufo sighting cities end in "port"?


```{r}

```

