---
title: "Data Cleaning Part 2 Lab Key"
output: html_document
---


```{r}
library(tidyverse)
```

1. Read in the UFO dataset (used in the Data IO lectures) as an R object called `ufo`.
You can read directly from the web here: https://raw.githubusercontent.com/SISBID/Module1/gh-pages/data/ufo/ufo_data_complete.csv . 
You can ignore the "problems" with some rows. 

```{r}
library(readr)
ufo <- read_csv("https://raw.githubusercontent.com/SISBID/Module1/gh-pages/data/ufo/ufo_data_complete.csv")
```

 Clean up the column/variable names of the `ufo` dataset to remove spaces and non-alphanumeric characters. You can use the `dplyr::rename()` function or use the `janitor::clean_names()` function. 

```{r}
library(janitor)
ufo <- clean_names(ufo)
```

2. How many UFO sightings were reported on a time scale of minutes, specifically using the `duration (hours/min)` originally-named column? (hint: use `str_detect` and `filter`, you can ignore observations like "1/2 hour" and similar ones that don't contain some version of the word "minutes").

```{r}
ufo %>% filter(str_detect(duration_hours_min, "min"))%>% nrow()
```

3. How accurate is the (formerly-named) `duration (seconds)` column? I.e. how many of the above minutes-scale observations have durations greater than 14400 seconds (or 4 hours)? 

```{r}
sub <- ufo %>% filter(str_detect(duration_hours_min, "min"),
                      duration_seconds >14400)
nrow(sub)
```

4. How many ufo sighting cities end in "port"? (hint - remember `stringr` uses vectors or variables)


```{r}
ufo %>% pull(city) %>% str_subset("port$") %>% length() # C
```


