---
title: "Data I/O Lab, Part 2"
output: html_document
---

In this lab you can use the interactive console to explore but please record your commands here. Remember anything you type here can be "sent" to the console with Cmd-Enter (OS-X) or Cntr-Enter (Windows/Linux) (But only in side the ```{r}``` areas).

```{r, message = FALSE}
library(tidyverse)
```


1. Read in a slice of the "ufo" dataset from "https://sisbid.github.io/Data-Wrangling/data/ufo/ufo_slice_6.csv" using `read_csv`, call it `ufo_slice_6`.

```{r}
ufo_slice_6 <- read_csv("https://sisbid.github.io/Data-Wrangling/data/ufo/ufo_slice_6.csv")
```

2. Use a few functions to check out the data (e.g., `head()`, `glimpse()`, `str()` etc.). Do results look different if you had used `read_delim` above?

```{r}
str(ufo_slice_6)
head(ufo_slice_6)
glimpse(ufo_slice_6)
nrow(ufo_slice_6)
```

3. Repeat 1-2 for "https://sisbid.github.io/Data-Wrangling/data/ufo/ufo_slice_7.txt" using `read_delim`, call it `ufo_slice_7`. Should any additional arguments be used?

```{r}
ufo_slice_7 <- read_delim("https://sisbid.github.io/Data-Wrangling/data/ufo/ufo_slice_7.txt")
ufo_slice_7
ufo_slice_7 <- read_delim("https://sisbid.github.io/Data-Wrangling/data/ufo/ufo_slice_7.txt", trim_ws = TRUE)
ufo_slice_7
```

