---
title: "Data I/O Lab, Part 2"
output: html_document
---

In this lab you can use the interactive console to explore but please record your commands here. Remember anything you type here can be "sent" to the console with Cmd-Enter (OS-X) or Cntr-Enter (Windows/Linux) (But only in side the ```{r}``` areas).

```{r, message = FALSE}
library(tidyverse)
```


1. Read in a slice of the "ufo" dataset from "https://sisbid.github.io/Data-Wrangling/data/ufo/ufo_slice_5.csv" using `read_csv`, call it `ufo_slice_5`.

```{r}

```

2. Use a few functions to check out the data (e.g., `head()`, `glimpse()`, `str()` etc.). Do results look different if you had used `read_delim` above?

```{r}

```

3. Repeat 1-2 for "https://sisbid.github.io/Data-Wrangling/data/ufo/ufo_slice_6.txt" using `read_delim`, call it `ufo_slice_6`. Should any additional arguments be used?

```{r}

```

## **Practice on your own**

4. Try using `read_table` and `read_delim` on "https://sisbid.github.io/Data-Wrangling/data/ufo/ufo_slice_7.txt". Which one is more appropriate?

```{r}

```
