---
title: "Functional Programming"
author: "Data Wrangling in R"
---

```{r, include=FALSE}
library(tidyverse)
library(repurrrsive)
library(tidyr)
```

## Functional Programming

"R, at its heart, is a functional programming (FP) language. This means that it provides many tools for the creation and manipulation of functions." - [Hadley Wickham](http://adv-r.had.co.nz/Functional-programming.html)

Don't need to write for-loops! - check this [video](https://www.youtube.com/watch?v=GyNqlOjhPCQ).

## Iterative work

Allows you to flexibly iterate functions to multiple elements of a data object!

Useful when you want to apply a function to:  

- lots of columns in a tibble  
- multiple tibbles  
- multiple data files  
 
## base R `apply` functions

Works really simply for all columns, but not a tibble.

```{r}
sapply(mtcars, FUN = round)
```

## Working **`across`** multiple columns

`across` allows us to perform functions on specific columns more easily. Use with `mutate` or `summarize`.

```{r}
mtcars %>% 
  mutate(across(
    c(mpg, disp, hp, drat), 
    round # no parentheses if no arguments
  ))
```

## Using `across` with arguments {.codesmall}

If you want to use arguments, then you need to use the `~`  and `.` (or `.x`) as a place holder for what you will be passing into the function.

See [documentation for purrr-shortcuts](https://adv-r.hadley.nz/functionals.html#purrr-shortcuts) and [a comparison of `.` and `.x`](https://forum.posit.co/t/function-argument-naming-conventions-x-vs-x/7764/10).

```{r}
mtcars %>% 
  mutate(across(
    c(mpg, disp, hp, drat), 
    ~ round(., digits = -1) # note ~ and .
  ))
```


# Writing functions

## Why write your own functions?

- Cut down on repetitive code (easier to fix things!)
- Organize code into manageable chunks
- Avoid running code unintentionally

## Functions in R

To create a function we can use the `function` function and specify what input the function will take and what it will do to it.  

```{r}
my_function <- function(x){x + 1}
my_function
```

```{r}
my_data  <- c(2,3,4)

my_function(x = my_data)
```

## Shortcut for functions

Alternatively we can use `\(x)`. See this [link about function shortcuts](https://medium.com/analytics-vidhya/how-to-use-the-new-function-symbol-in-r-programming-4-1-r-4-1-c7cd2ac27b98).

```{r}
my_function <- \(x){x + 1}
my_function

my_function(x = my_data)
```

## `purrr` is a super helpful package for iteration!

"Designed to make your functions purrr."

`dplyr` is designed for data frames
`purrr` is designed for vectors

The `purrr` package can be very helpful!

* https://purrr.tidyverse.org/

* https://github.com/rstudio/cheatsheets/raw/master/purrr.pdf

* https://jennybc.github.io/purrr-tutorial/ 

## `purrr` functions: `map` and `modify`

**applies function** to each element of an vector or object

- `map` will output a list
- `map_dbl` will output a vector
- `modify` will output the same object type

## `purrr` functions: `map` and `modify`

Apply `sum` to every column

```{r}
mtcars |> map(sum)
```

## `purrr` functions: `map` and `modify`

Apply `sum` to every column - nicer output

```{r}
mtcars |> map_dbl(sum)
```

## `purrr` functions: `map` and `modify`

Apply `round` to every column - get a data.frame back out

```{r}
mtcars |> modify(round)
```

## Anonymous functions - "functions on the fly"

You can supply a custom function without naming it. Useful if you only need to use it once!

```{r}
mtcars |> modify(\(x) x * 100)
```

## `modify_if` {.codesmall}

Using `modify_if()`, we can specify what columns to modify

```{r message=FALSE}
ufo1 <- read_csv("https://sisbid.github.io/Data-Wrangling/data/ufo/ufo_slice_1.csv")
```

```{r}
ufo1 %>% 
  modify_if(is.character, toupper) %>%
  head(3)
```

## Speed test! modify_if vs across

```{r}
system.time(ufo1 %>% 
              modify_if(is.character, toupper))

system.time(ufo1 %>%
               mutate(across(where(is.character), toupper)))
```

## Iterative filtering with `if_all`

Previously we filtered for patterns or conditions..

Dilemma: Seems a bit repetitive!

```{r}
mtcars %>%
  filter(cyl > 3 & cyl < 8,
         gear > 3 & gear < 8,
         carb > 3 & carb < 8)
```

## Now we can filter multiple columns!

`if_all()`: helps us filter on multiple similar conditions simultaneously!

```{r}
mtcars %>%
   filter(if_all(c(cyl, gear, carb), ~.x > 3 & .x < 8))
```

# Lists

## What is a list?

* Lists are the most flexible/"generic" data class in R
* Can be created using list()
* Can hold vectors, strings, matrices, models, list of other lists, lists upon lists!

```{r makeList}
mylist <- list(
  letters = c("A", "b", "c"),
  numbers = 1:3,
  matrix(1:25, ncol = 5),
  iris
)
```

## List Structure

```{r Lists}
head(mylist)
```

## List referencing

```{r Listsref1}
mylist[1] # returns a list
mylist["letters"] # returns a list
```

## List referencing
  
```{r Listsrefvec}  
mylist[[1]] # returns the vector 'letters'
mylist[["letters"]] # returns the vector 'letters'
```

## Why do this at all?

https://jennybc.github.io/purrr-tutorial/bk01_base-functions.html

You need a way to iterate in R in a data-structure-informed way. What does that mean?

* Iterate over elements of a list
* Iterate over rows or columns of a 2-dimensional object
* Iterate over sub data frames induced by one or more factors
* Iterate over tuples formed from the i-th element of several vectors of equal length

## Cleaning up multiple datasets {.codesmall}

Store datasets in a list!

```{r message=FALSE}
ufo1 <- read_csv("https://sisbid.github.io/Data-Wrangling/data/ufo/ufo_slice_1.csv")
ufo2 <- read_delim("https://sisbid.github.io/Data-Wrangling/data/ufo/ufo_slice_2.tsv")
ufo4 <- read_delim("https://sisbid.github.io/Data-Wrangling/data/ufo/ufo_slice_4.csv", delim = ":")
```

```{r}
ufo_datasets <- list(ufo1, ufo2, ufo4)
```

## Cleaning up multiple datasets

Clean names on all datasets at once:

```{r message=FALSE}
library(janitor)

ufo_datasets_clean <- 
  ufo_datasets %>%
  map(\(x) x |> count(country))
```

## Cleaning up multiple datasets {.codesmall}

Confirm columns have been cleaned! Look at `ufo_datasets_clean` or look at each with indexing, eg.:

```{r message=FALSE}
ufo_datasets_clean[[1]]
```

## List: `group_split()` a dataset 

We can create a list by splitting up a dataframe. We will use `mtcars`.

```{r}
head(mtcars)
```

## List: `group_split()` the dataset by `cyl` {.codesmall}

The following creates split of data for each unique cyl value:

```{r}
mtcars_split <- mtcars %>% group_by(cyl) %>% group_split()
mtcars_keys <- mtcars %>% group_by(cyl) %>% group_keys() %>% pull(cyl)
names(mtcars_split) <- mtcars_keys
glimpse(mtcars_split)
```

## List: model on each

```{r}
mtcars_split %>% 
  map(~lm(mpg ~ wt, data = .)) %>% # apply linear model to each
  map(summary) %>%
  map_dbl("r.squared")
```

## Summary

- `function(x){ }` or `\(x){ }` denotes a function. You also commonly see `~.x` inside `across`.
- `map_df` and `modify` **apply functions** to each element of an object. `map` returns a list, `modify` returns the same object type.
- The purrr package has other useful functional programming features.
- lists can be great for storing iterative work.
- `group_split` and `group_keys` can be handy with `group_by`

https://sisbid.github.io/Data-Wrangling/14_Functional_Programming/lab/functional-program-lab.Rmd
