---
title: "Functional Programming"
output: html_document
---

```{r, include = FALSE}
library(tidyverse)
```

Check out this purrr tutorial by Jenny Bryan: https://jennybc.github.io/purrr-tutorial/ls02_map-extraction-advanced.html

Additional content to read up on: 
https://jennybc.github.io/purrr-tutorial/ls01_map-name-position-shortcuts.html

0. Load the UFO dataset:

```{r}
ufo <- read_csv(
"https://sisbid.github.io/Data-Wrangling/data/ufo/ufo_data_complete.csv")
```

1. Working with the `ufo` data, use `mutate` and `across` to change the character variables (`where(is.character)`) to be uppercase with `toupper`.

```{r}

```

2. Now do the same thing as question one, but this time use `modify`.

```{r}

```

3. Which of question 1 or 2 is faster? Try piping each response into `system.time()`.

```{r}

```

4. Use `group_split()` to split up the ufo dataset by country. Assign the list to the object `ufo_list`

EG: ??? <- ??? %>% group_by(???) %>% group_split()

```{r}

```

5. Add the list keys.

EG: 
ufo_keys <- ??? %>% group_by(???) %>% group_keys() %>% pull(???)
names(ufo_list) <- ufo_keys

```{r}

```

6. For each of the datasets in ufo_list, `map` the anonymous function: `\(x) count(x, shape)`

```{r}

```
