---
title: "Functional Programming"
output: html_document
---

```{r, include = FALSE}
library(tidyverse)
```

Check out this purrr tutorial by Jenny Bryan: https://jennybc.github.io/purrr-tutorial/ls02_map-extraction-advanced.html

Additional content to read up on: 
https://jennybc.github.io/purrr-tutorial/ls01_map-name-position-shortcuts.html

0. Create a new dataset from `iris` called `iris_lab`. Use this for the lab.

```{r}
iris_lab <- iris
```

1. Working with the `iris_lab` data, use `mutate` and `across` to change the numeric variables to be rounded with the `round` function. Take a look with `head()`- do this for the rest of the questions in which you create an output of data.

```{r}

```

2. Now do the same thing as question one, but this time set the number of digits for rounding to be 1.

```{r}

```

3. Use `if_all()` to check if there are iris flowers with all measurements greater than 2 centimeters (which is the unit, if you type in ?iris you can learn more about the data). Hint - use `filter`. If any are, how many?

```{r}

```

4.  the function `format()` with the argument `scientific = TRUE` can be used to format numbers in scientific notation. For example:

```{r}
format(10000, scientific = TRUE)
```

Use `modify_if` to convert *numeric* columns of `iris_lab` to be in scientific notation. 

```{r}

```
