---
title: "Advanced IO lab"
output: html_document
editor_options: 
  chunk_output_type: console
---

## Google Sheets Lab

```{r, include=FALSE}
library(tidyverse)
library(httr)
library(jsonlite)
library(googlesheets4)
```

This is a lab to practice with Google Sheets and the `googlesheets` package.

1. We are going to use a sheet from previous years: https://docs.google.com/spreadsheets/d/1KIRtcPVn58R3_qr97WNtcOJiY4AaytHzGDzLW_3_R1s/edit?usp=sharing

```{r}
sheet_url = "https://docs.google.com/spreadsheets/d/1KIRtcPVn58R3_qr97WNtcOJiY4AaytHzGDzLW_3_R1s/edit?usp=sharing"
```


2. Use the `googlesheets4::read_sheet` function to read in the data like we discussed in class, call this object `x`.
```{r}

```

3. Using `sheet_names` function to get the sheet names from the `sheet_url`.
```{r}

```


4. Try reading in just the first two columns. See the `range` argument and look at the arguments for `cell_rows`, `cell_limits` and `cell_cols`.
```{r}

```


## JSON Lab


See GitHub Documentation at https://docs.github.com/en/rest/reference/repos

1. Use Github API to get all of Hadley Wickham's repo data (username = "hadley"), see: "https://api.github.com/users/XXXX/repos".  Use `httr:GET` and assign that output to `response` and then `httr::content(response)`.  Also run `httr:status_code(response)` to get the status code

```{r}
# response = httr::GET()
```

2.  Try `jsonlite::fromJSON(httr::content(response, as = "text", flatten = TRUE))`
```{r, error = TRUE}
out = jsonlite::fromJSON(httr::content(response, as = "text", flatten = TRUE))
class(out)
head(out)
```


3. How many stars does he have on each repository? (it's the `stargazers_count` element)
```{r}

```

4. Create a new data frame consisting of just the "owner" column of the resulting data set called `owner_data`

```{r}

```
