---
title: "Advanced IO lab key"
output: html_document
editor_options: 
  chunk_output_type: console
---

## Google Sheets Lab

```{r, include=FALSE}
library(tidyverse)
library(httr)
library(jsonlite)
library(googlesheets4)
```

This is a lab to practice with Google Sheets and the `googlesheets` package.

1. We are going to use a sheet from previous years: https://docs.google.com/spreadsheets/d/1KIRtcPVn58R3_qr97WNtcOJiY4AaytHzGDzLW_3_R1s/edit?usp=sharing

```{r}

```


2. Use the `googlesheets4::read_sheet` function to read in the data like we discussed in class, call this object `x`.

```{r}

```

3. Using `sheet_names` function to get the sheet names from the `sheet_url`.

```{r}

```


4. Try reading in just the first two columns. See the `range` argument and look at the arguments for `cell_rows`, `cell_limits` and `cell_cols`.

```{r}

```

## JSON Lab

<!-- See here for more fun practice! https://github.com/jdorfman/Awesome-JSON-Datasets -->

The following dataset lists airports in the US and details about the number of late flights over time.

1. Read in data from the following link: https://think.cs.vt.edu/corgis/datasets/json/airlines/airlines.json. Call this `jsonData`.

```{r}

```

2. Investigate the data. What are the three main items in `jsonData`?

```{r, error = TRUE}

```

3. Filter `jsonData` to include only NYC LaGuardia and Chicago O'Hare airports (Code : "LGA", "ORD") and Year : 2016. Call this `lga_ord`.

```{r}

```

4. Create a new tibble out of Airport Code (`Airport$Code`), Total Flights (`Statistics$Flights$Total`), and create a new column for proportion of delayed flights using (`Statistics$Flights$Delayed`). Call this new tibble `airport_compare`.

```{r}

```
