---
title: "Advanced IO lab key"
output: html_document
editor_options: 
  chunk_output_type: console
---

## Google Sheets Lab

```{r, include=FALSE}
library(tidyverse)
library(jsonlite)
library(googlesheets4)
```

This is a lab to practice with Google Sheets and the `googlesheets` package.

Make sure you go through the authentication process. You'll see a popup and will need to grant permissions.

```{r eval=FALSE}
gs4_auth()
```

1. We are going to use a sheet from previous years: https://docs.google.com/spreadsheets/d/1KIRtcPVn58R3_qr97WNtcOJiY4AaytHzGDzLW_3_R1s/edit?usp=sharing

```{r}
sheet_url <- "https://docs.google.com/spreadsheets/d/1KIRtcPVn58R3_qr97WNtcOJiY4AaytHzGDzLW_3_R1s/edit?usp=sharing"
```

2. Use the `read_sheet()` function to read in the data like we discussed in class, call this object `x`.

```{r}
x <- read_sheet(sheet_url)
```

3. Using `sheet_names` function to get the sheet names from the `sheet_url`.

```{r}
sheet_names(sheet_url)
```

4. Try reading in just the first two columns. Use the `range = cell_cols()` argument.

```{r}
read_sheet(sheet_url, range =  cell_cols("A:B"))
```

5. How often do you see the word "learn" in the dataset `x`? Hint: use a `stringr` function.

```{r}
x %>%
  filter(str_detect(`Why are you taking this module (free text)`, pattern = "learn"))
```

## JSON Lab

<!-- See here for more fun practice! https://github.com/jdorfman/Awesome-JSON-Datasets -->

The following dataset lists airports in the US and details about the number of late flights over time.

1. Read in data from the following link: https://think.cs.vt.edu/corgis/datasets/json/airlines/airlines.json. Call this `jsonData`.

```{r}
jsonData <- fromJSON("https://think.cs.vt.edu/corgis/datasets/json/airlines/airlines.json")
```

2. Investigate the data. What are the three main items in `jsonData`?

```{r, error = TRUE}
str(jsonData)
# Airport, Time and Statistics
dim(jsonData$Airport)
dim(jsonData$Time)
dim(jsonData$Statistics)

colnames(jsonData$Statistics)
```

3. Filter `jsonData` to include only Year : 2016. Call this `air_2016`.

```{r}
air_2016 <- jsonData %>% 
  filter(Time$Year == 2016)
```

4. How many unique Airports are there?

```{r}
air_2016$Airport %>% count()
# OR
length(unique(air_2016$Airport$Code))
```

5. Filter `air_2016` to include only NYC LaGuardia and Chicago O'Hare airports (Code : "LGA", "ORD"). Call this `lga_ord`.

```{r}
lga_ord <- jsonData %>% 
  filter(Airport$Code %in% c("LGA", "ORD"))
```

**Bonus Practice**

Create a new tibble out of Airport Code (`Airport$Code`), Total Flights (`Statistics$Flights$Total`), and create a new column for proportion of delayed flights using (`Statistics$Flights$Delayed`). Call this new tibble `airport_compare`.

```{r}
airport_list <- list(
  airport_code = lga_ord$Airport$Code,
  total_flights = lga_ord$Statistics$Flights$Total,
  delayed_proportion = lga_ord$Statistics$Flights$Delayed / lga_ord$Statistics$Flights$Total
  )
airport_compare <- as_tibble(airport_list)
airport_compare
```
