---
title: "Data I/O Lab, Part 1"
output: html_document
---

```{r, include = TRUE}
library(tidyverse)
```

This lab uses different versions of the `iris` dataset.  For each question, click the `iris` file in the Files pane to open it in an RStudio window.

You can see more about this dataset by running `?iris` in the console.

1. Read in the `iris_q1.csv` dataset into the `iris_q1` R object. What delimiter separates columns?

```{r}

```


2. Read in the `iris_q2.tsv` dataset into the `iris_q2` R object. What delimiter separates columns?
```{r}

```
3. Read in the `iris_q3.txt` dataset into the `iris_q3` R object. What delimiter separates columns?
```{r}

```
4. Read in the `iris_q4.csv` dataset into the `iris_q4` R object. What delimiter separates columns? [hint: file extension is ambiguous]
```{r}

```
5. Read in the `iris_q5.csv` dataset into the `iris_q5` R object. Some entries are supposed to be missing, e.g. have the value `NA` in the data.  
```{r}

```
6. Read in the `iris_q6.xlsx` dataset into the `iris_q6` R object. How many rows are in the dataset?
```{r}

```