---
title: "Data Summarization Lab"
output: html_document
editor_options: 
  chunk_output_type: console
---

# Data used

Circulator Lanes Dataset: the data is from https://data.baltimorecity.gov/Transportation/Charm-City-Circulator-Ridership/wwvu-583r

You can Download as a CSV in your current working directory.  Note its also available at:  http://sisbid.github.io/Data-Wrangling/data/Charm_City_Circulator_Ridership.csv

```{r}
library(tidyverse)

circ = read_csv(paste0("http://sisbid.github.io/Data-Wrangling/",
                       "data/Charm_City_Circulator_Ridership.csv"))
```

1. How many days are in the data set?  You can assume each observation/row is a different day

```{r q1}

```

2. What is the total (sum) number boardings on the green bus (`greenBoardings`)?

```{r q2}

```

3. How many days are missing daily ridership?  Use `is.na()` and either `table()`, `count()`, or `n()`.

```{r q3}

```

4. Group the data by day of the week (`day`). Which day of the week has the most average ridership (the `daily` column)?  How many days goes are observed for those days?

```{r q4}

```

5. Describe the distribution of `orangeBoardings`(use `quantile()`).

```{r q6}


```

6. Describe the distribution of `orangeBoardings`(use `quantile()`), but this time stratify by day of the week.

```{r q7}


```

7.  Bonus: Describe the results of #5 but using a boxplot visualization:

```{r}
# Hint:
# qplot(x = ???, y = ???, data = ???, geom = "boxplot")


```
