---
title: "Data Reshaping Lab Key"
author: "Data Wrangling in R"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```

1.  Read in the Circulator Data from http://sisbid.github.io/Data-Wrangling/data/Charm_City_Circulator_Ridership.csv.  Name the data set `wide`.

```{r, message = FALSE}
wide = read_csv(
  "http://sisbid.github.io/Data-Wrangling/data/Charm_City_Circulator_Ridership.csv"
)
```

2. Reshape wide using `pivot_wider`. Call this data `long.`  Make the `names_to = `
`variable`, and the value `n_people`  Make sure we gather all columns **but**
`daily`, `day`, `date`, using the `!`.

```{r}
long = wide %>% 
  pivot_longer( !c(daily, day, date), names_to = "variable", values_to = "n_people")
head(long)
```

3.  Separate the `date` column into ("month", "day", "year") using a sep of "/"
Make sure you don't remove the `date` column, call this `long2`.  Do a table of the day column.  Where did the original `day` column go?

```{r}
# We over-wrote the day column by assigning it new values (in this case, calendar "day")
long2 = long %>%
  separate(col = "date", into = c("month", "day", "year"),
           sep = "/", remove = FALSE)
table(long2$day)
```

4. Create a new variable in `long2` calling it `newdate` using the `unite` function.  Unite the year, month, day, columns using a hyphen (`-`) and then `table` the new column `newdate`

```{r}
long2 = long2 %>%
  unite(col = "newdate", year, month, day, sep = "-")
head(long2)
```

5. Use `str_replace()` on `long` (not `long2`) to replace "Board" with "_Board", "Alight" with "_Alight", and "Average" with "_Average".

```{r, error = TRUE}
long = long %>% 
  mutate(
    variable = variable %>% 
      str_replace("Board", "_Board") %>% 
      str_replace("Alight", "_Alight") %>% 
      str_replace("Average", "_Average") 
  )
head(long)
```

6.  Separate the `variable` column from `long` into ("route", "type") using a sep of "_".

```{r}
long = long %>%
  separate(col = "variable", into = c("route", "type"), sep = "_")
head(long)
```

7. Reshape `long` (not `long2`) using `pivot_wider`.  Spread the data out so the `route` becomes the new column name and the cells are filled with the `n_people` value.  Call this `wide_by_route`.

```{r}
wide_by_route = long %>% 
  pivot_wider(names_from = route, values_from = n_people)
```
