---
title: "Data Reshaping Lab Key"
author: "Data Wrangling in R"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```

1.  Read in the Circulator Data from http://sisbid.github.io/Module1/data/Charm_City_Circulator_Ridership.csv.  Name the data set `wide`.


```{r, message = FALSE}
wide = read_csv(
  paste0("http://sisbid.github.io/Module1/",
         "data/Charm_City_Circulator_Ridership.csv"))
head(wide, 2)
class(wide$date)
```


2. Reshape `wide` using `gather.`  Call this data `long.`  Make the `key`
`variable`, and the value `n_people`  Make sure we gather all columns **but**
`daily`, `day`, `date`, using things like `-day`. Table `long$day`

```{r}
long = wide %>% 
  gather(key = "variable", value = "n_people", -daily, -day, -date)
head(long$day)
```


3.  Separate the `date` column into ("month", "day", "year") using a sep of "/"
Make sure you don't remove the `date` column, call this `long2`.  Do a table of the day column.  Where did the original `day` column go?
```{r}
long2 = separate(long, col = "date", into = c("month", "day", "year"), 
                sep = "/", remove = FALSE)
table(long2$day)
```

4. Create a new variable in `long2` calling it `newdate` using the `unite` function.  Unite the year, month, day, columns using a hyphen (`-`) and then `table` `road_hyphen`
```{r}
long2 = unite(long2, col = "newdate", year, month, day, sep = "-")
head(long2)
```

```{r, error = TRUE}
long = long %>% 
  mutate(
    variable = variable %>% 
      str_replace("Board", "_Board") %>% 
      str_replace("Alight", "_Alight") %>% 
      str_replace("Average", "_Average") 
  )
```

5.  Separate the `variable` column from `long` (not `long2`) into ("route", "type") using a sep of "_"
```{r}
long = separate(long, col = "variable", into = c("route", "type"), 
                sep = "_")
```

6. Reshape `long` (not `long2`) using `spread.`  Spread the data where the `route` and we want the value in the new columns to be `n_people` .  Call this `wide_by_route`

```{r}
wide_by_route = spread(long, key = route, value = n_people)
```


