---
title: "Data Summarization Lab"
output: html_document
editor_options: 
  chunk_output_type: console
---

# Data used

Circulator Lanes Dataset: the data is from https://data.baltimorecity.gov/Transportation/Charm-City-Circulator-Ridership/wwvu-583r

You can Download as a CSV in your current working directory.  Note its also available at:  http://sisbid.github.io/Data-Wrangling/data/Charm_City_Circulator_Ridership.csv

```{r}
library(tidyverse)

circ = read_csv(paste0("http://sisbid.github.io/Data-Wrangling/",
                       "data/Charm_City_Circulator_Ridership.csv"))
```



1. How days are in the data set?  You can assume each observation/row is a different day
```{r q1}

```

2. What is the total (sum) number boardings on the green bus?

```{r q2}

```

3.  How many days are missing daily ridership?  Use `is.na` and either table, count?
   
```{r q3}

```


4. Which day of the week has the most average ridership (using `daily`)?  How many days goes are observed for those days? Group the data.

```{r q4}

```



5. (a) Numerically [hint: `quantile()`] and 
		(b) graphically [hint: `qplot(, data = circ)`]
		 describe the distribution of `orangeBoardings`
```{r q6}

```

6. Then describe as above, after stratifying by day of the week.
```{r q7}

```

7.  Describe the following output and then flip the `x` and `y` variables:
```{r}
qplot(x = orangeBoardings, y = day, data = circ, geom = "boxplot")
```


