---
title: "Data Subsetting, Part 2 - Lab Key"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In this lab you can use the interactive console to explore but please record your commands here.  Remember anything you type here can be "sent" to the console with Cmd-Enter (OS-X) or Cntr-Enter (Windows/Linux) (But only in side the ```{r}``` areas).

```{r, message = FALSE}
library(dplyr)
library(tidyverse)
```


1. In the data set `mpg`, rename the column `year` to `YEAR` using `rename`.

```{r}
mpg = mpg %>% rename(YEAR = year)
```

2. Convert the column names of `mpg` to all upper case. Use `rename_all`, and the `toupper` command (or `colnames`).

```{r}
mpg = rename_all(mpg, toupper)
head(mpg)
```



3. Re-order the rows of `mpg` by `CYL` in increasing order. Use `arrange()`. Don't forget to reassign the `mpg` data!

```{r}
mpg = mpg %>% arrange(CYL)
```


4. Create a new variable in `mpg` called `CTY2`, which  is equal to `CTY^2`, using `mutate()`.  Use piping `%>%`:

```{r}
mpg = mpg %>% mutate(CTY2 = CTY^2)
mpg
```


5. Create a new variable in `mpg` called `group_class`, which in the `CLASS` variable keeps "pickup" and "suv" as they are, but map "compact", "midsize", "minivan", "subcompact", and "2seater" to "car" using
`case_when` or `ifelse`, `%in%`, and `mutate`:

```{r}
# any of the following work because all cars are in either of the to groups of car `CLASS` options listed

mpg1 = mpg %>% 
  mutate(
    group_class = case_when(
      CLASS %in% c("compact", "midsize", "minivan",
                   "subcompact", "2seater") ~ "car",
      TRUE ~ CLASS)
  )

mpg2 = mpg %>% 
  mutate(
    group_class = case_when(
      CLASS %in% c("compact", "midsize", "minivan",
                   "subcompact", "2seater") ~ "car",
      CLASS %in% c("pickup", "suv") ~ CLASS)
  )

mpg3 = mpg %>% 
  mutate(
    group_class = case_when(
      CLASS %in% c("pickup", "suv") ~ CLASS,
      TRUE ~ "car")
  )

mpg4 = mpg %>% 
  mutate(
    group_class = if_else(
      CLASS %in% c("pickup", "suv"), CLASS, "car")
  )

mpg5 = mpg %>% 
  mutate(
    group_class = if_else(
      CLASS %in% c("compact", "midsize", "minivan",
                   "subcompact", "2seater"), "car", CLASS)
  )

identical(mpg1, mpg2)
identical(mpg1, mpg3)
identical(mpg1, mpg4)
identical(mpg1, mpg5)
```
