---
title: "Data I/O Lab, Part 2"
output: html_document
---

In this lab you can use the interactive console to explore but please record your commands here. Remember anything you type here can be "sent" to the console with Cmd-Enter (OS-X) or Cntr-Enter (Windows/Linux) (But only in side the ```{r}``` areas).

```{r, message = FALSE}
library(tidyverse)
library(readxl)
```


1. Read in the the Charm City Circulator Dataset from "https://sisbid.github.io/Data-Wrangling/data/Charm_City_Circulator_Ridership.csv" using `read_csv`, call it `circ`

```{r}
circ <- read_csv("https://sisbid.github.io/Data-Wrangling/data/Charm_City_Circulator_Ridership.csv")
```

2. Use a few functions to check out the data (e.g., `head()`, `glimpse()`, `str()` etc.).

```{r}
str(circ)
head(circ)
glimpse(circ)
nrow(circ)
```

3. Write out `circ` to a file called "Circulator.csv" using `write_csv`.

```{r}
write_csv(circ, "Circulator.csv")
```

***
Download the Excel data for the next two questions. Make sure it's in your project folder.
```{r}
curl::curl_download("https://sisbid.github.io/Data-Wrangling/data/iris/iris_q6.xlsx", "iris_q6.xlsx")
```
***

4. Read in sheet 1 of the `iris_q6.xlsx` dataset into the `iris_q4` R object. How many rows are in the dataset?

```{r}
iris_q4 <- read_excel("iris_q6.xlsx", sheet= 1)
dim(iris_q4)
```

5. Read in sheet 2 of the `iris_q6.xlsx` dataset into the `iris_q6_2` R object. How many rows are in the dataset?

```{r}
iris_q5 <- read_excel("iris_q6.xlsx", sheet= 2)
dim(iris_q5)
```
