---
title: "Databases lab"
author: "Jeff Leek"
date: "July 12, 2016"
output: html_document
---

1. Download and load the nycflights data with the command `install.packages('nycflights13')` and `library(nycflights13)`.

```{r}
is_package_installed = function(package) {
  requireNamespace(basename(package), quietly = TRUE)
}
if (!is_package_installed("nycflights13")) install.packages("nycflights13")
library(nycflights13)
```


2. Use the `pryr` package to figure out the size of the `flights` object. 

```{r}
pryr::object_size(flights)
```


3. Create a sqlite database, then add a table "flights" with the flights data from this package.

```{r}
my_flights_db <- src_sqlite("my_flights_db.sqlite3", 
create = TRUE)
flights_sqlite <- copy_to(my_flights_db, 
flights, temporary = FALSE,overwrite = TRUE)
```


4. Inspect the tables using the `src_tbls` command to make sure the copying happened correctly.

```{r}
src_tbls(my_flights_db)
my_flights_db %>% tbl("flights")
```


5. Find the average delay time for American Airlines (hint: the abbreviation is AA).

```{r}
ave_delay = my_flights_db %>%
  tbl("flights") %>%
  filter(carrier == "AA") %>%
  summarise(ave_delay = mean(dep_delay))
```



6. How long does it take to collect the results of your computation for 5? 

```{r}
system.time(my_flights_db %>%
  tbl("flights") %>%
  filter(carrier == "AA") %>%
  summarise(ave_delay = mean(dep_delay)))
```


7. Can you figure out the average delay time for each airline? 

```{r}
ave_delay = my_flights_db %>%
  tbl("flights") %>%
  group_by(carrier) %>%
  summarise(ave_delay = mean(dep_delay))
```

8. Can you add a variable for averge delay by carrier to the database? 

```{r}

### Doesn't work!! No mutate 
#ave_delay = my_flights_db %>%
#  tbl("flights") %>%
#  group_by(carrier) %>%
#  mutate(ave_delay = mean(dep_delay))

ave_delay = my_flights_db %>%
  tbl("flights") %>%
  group_by(carrier) %>%
  summarise(ave_delay = mean(dep_delay)) %>%
  collect()


### Doesn't work! Have to copy it over

# my_flights_db %>%
#  tbl("flights") %>%
#   left_join(ave_delay)


 my_flights_db %>%
  tbl("flights") %>%
   left_join(ave_delay,copy=TRUE)
  
```
