---
title: "Data Summarization Lab Key"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data used

Circulator Lanes Dataset: the data is from https://data.baltimorecity.gov/Transportation/Charm-City-Circulator-Ridership/wwvu-583r

Available on: https://sisbid.github.io/Data-Wrangling/data/Charm_City_Circulator_Ridership.csv

```{r}
library(tidyverse)

circ <- read_csv("https://sisbid.github.io/Data-Wrangling/data/Charm_City_Circulator_Ridership.csv")
```

1. Each row is a different day. How many days are in the data set?

```{r q1}
nrow(circ)
dim(circ)
circ %>% 
  nrow()
```

2. What is the total (sum) number of boardings on the green bus (`greenBoardings` column)?

```{r q2}
sum(circ$greenBoardings, na.rm = TRUE)

circ %>% pull(greenBoardings) %>% sum(na.rm = TRUE)

count(circ, wt = greenBoardings)
```

3. Group the data by day of the week (`day`). Find the mean daily ridership (`daily` column). (hint: use `group_by` and `summarize` functions)

```{r q3}
circ %>% 
  group_by(day) %>% 
  summarize(mean = mean(daily, na.rm = TRUE))
```

4. Take the median of `orangeBoardings`(use `median()`), grouping by day of the week.

```{r q4}
circ %>% 
  group_by(day) %>% 
  summarize(median = median(orangeBoardings, na.rm = TRUE))
```

## **Practice on your own**

5. Group by day of the week, and then find the mean ridership `summarize(across())` all numeric columns using `where(is.numeric)`.

```{r q5}
circ %>% 
  group_by(day) %>% 
  summarize(across(where(is.numeric), ~mean(.x, na.rm=T)))
```
