---
title: "Data I/O Lab, Part 2"
output: html_document
---

In this lab you can use the interactive console to explore but please record your commands here.  Remember anything you type here can be "sent" to the console with Cmd-Enter (OS-X) or Cntr-Enter (Windows/Linux) (But only in side the ```{r}``` areas).

```{r, message = FALSE}
library(tidyverse)
```


1. Read in the the Charm City Circulator Dataset from "https://sisbid.github.io/Data-Wrangling/data/Charm_City_Circulator_Ridership.csv" using `read_csv`, call it `circ`

```{r}

```

2. Use a few functions to check out the data (e.g., `str()`, `head()`, `glimpse()` etc.).

```{r}

```

3. Write out `circ` to a file called "Circulator.csv" using `write_csv`.

```{r}

```

4. Write out `circ` to a file called "Circulator.rds" using `write_rds`.

```{r}

```

5. Read in the `https://sisbid.github.io/Data-Wrangling/data/iris_q5.csv` dataset into the `iris_q5` R object. Some entries read "-999" but are supposed to be missing (the value `NA` in the data).

```{r}

```

6. Read in sheet 1 of the `iris_q6.xlsx` dataset into the `iris_q6_1` R object. How many rows are in the dataset?

```{r}

```

7. Read in sheet 2 of the `iris_q6.xlsx` dataset into the `iris_q6_2` R object. How many rows are in the dataset?

```{r}

```
