---
title: "Data Summarization Lab Key"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Data used

Circulator Lanes Dataset: the data is from https://data.baltimorecity.gov/Transportation/Charm-City-Circulator-Ridership/wwvu-583r

You can Download as a CSV in your current working directory.  Note its also available at:  http://sisbid.github.io/Module1/data/Charm_City_Circulator_Ridership.csv

```{r}
library(tidyverse)

circ = read_csv(paste0("http://sisbid.github.io/Module1/",
                       "data/Charm_City_Circulator_Ridership.csv"))
```



1. How days are in the data set?  You can assume each observation/row is a different day
```{r q1}
nrow(circ)
dim(circ)
circ %>% 
  nrow()
```

2. What is the total (sum) number boardings on the green bus?

```{r q2}
sum(circ$greenBoardings, na.rm = TRUE)
```

3.  How many days are missing daily ridership?  Use `is.na` and either table, count?
   
```{r q3}
table(is.na(circ$daily), useNA = "ifany")

circ %>% 
  count(is.na(daily))

circ %>% 
  mutate(missing_daily = is.na(daily)) %>% 
  group_by(missing_daily) %>% 
  summarise(n())
```


4. Which day of the week has the most average ridership (using `daily`)?  How many days goes are observed for those days? Group the data.

```{r q4}
circ %>% 
  group_by(day) %>% 
  summarise(n = n(),
            non_missing = sum(!is.na(daily)),
            mean = mean(daily, na.rm = TRUE)) %>% 
  arrange(desc(mean)) 

circ %>% 
  group_by(day) %>% 
  summarise(n = n(),
            non_missing = sum(!is.na(daily)),
            mean = mean(daily, na.rm = TRUE)) %>% 
  arrange(desc(mean)) %>% 
  ungroup() %>% 
  slice(1) %>% 
  magrittr::extract("day")
```



5. (a) Numerically [hint: `quantile()`] and 
		(b) graphically [hint: `qplot(, data = circ)`]
		 describe the distribution of `orangeBoardings`
```{r q6}
quantile(circ$orangeBoardings, na.rm = TRUE)
circ %>% 
  summarise(quantile(orangeBoardings, na.rm = TRUE))

qplot(x = orangeBoardings, data = circ, geom = "histogram")
qplot(x = orangeBoardings, data = circ, geom = "density")
```

6. Then describe as above, after stratifying by day of the week.
```{r q7}
quantile(circ$orangeBoardings, na.rm = TRUE)
circ %>% 
  group_by(day) %>% 
  summarise(quantile(orangeBoardings, na.rm = TRUE))
```

7.  Describe the following output and then flip the `x` and `y` variables:
```{r}
qplot(x = orangeBoardings, y = day, data = circ, geom = "boxplot")
qplot(x = day, y = orangeBoardings, data = circ, geom = "boxplot")
```


