---
title: "Data Cleaning"
author: "Andrew Jaffe (key)"
date: "July 12, 2016"
output: html_document
---

### A. "Real Property Taxes" Dataset 

Download the "real property taxes" dataset from the website (via OpenBaltimore), the data is located here (note you don\'t need to unzip it to read it into R):
http://sisbid.github.io/Module1/data/real_property_tax.csv.gz

1. Read the Property Tax data into R and call it the variable `tax`

```{r}
library(readr)
library(stringr)
suppressPackageStartupMessages(library(dplyr))
tax = read_csv("http://sisbid.github.io/Module1/data/real_property_tax.csv.gz")
# write_csv(tax, gzfile("table.csv.gz))
```

Note the `read_csv` function is treating the empty string `""` as missing whereas the `read.csv` function is not.

```{r}
tax2 = read.csv("../data/real_property_tax.csv.gz",as.is=TRUE)
table(complete.cases(tax))
table(complete.cases(tax2))
ind = which(!complete.cases(tax))[1:3]
tax[ind,]
tax2[ind,]
```

```{r}
tax3 = read.csv("../data/real_property_tax.csv.gz",as.is=TRUE,na.string="")
```

2. How many addresses pay property taxes? 

```{r}
nrow(tax)
dim(tax)
```

3. What is the total city and state tax paid?

```{r}
head(tax$cityTax)
cityTax = tax$cityTax %>% 
  str_replace(fixed("$"), "") %>%
  as.numeric
stateTax = tax$stateTax %>% 
  str_replace(fixed("$"), "") %>%
  as.numeric
head(cityTax)

head(tax$cityTax[ is.na(cityTax) ])
table(tax$cityTax[ is.na(cityTax) ])

head(tax$stateTax[ is.na(stateTax) ])
table(tax$stateTax[ is.na(stateTax) ])

tax$cityTax = cityTax
tax$stateTax = stateTax

sum(tax$cityTax, na.rm = TRUE)
sum(tax$cityTax, na.rm = TRUE)/1e6

sum(tax$stateTax, na.rm = TRUE)
sum(tax$stateTax, na.rm = TRUE)/1e6
```

Combined would be `r round((sum(tax$stateTax, na.rm = TRUE) + sum(tax$cityTax, na.rm = TRUE))/1e6)` milion dollars paid.


4. How many observations/properties are in each residence code?

```{r}
table(tax$resCode)
```

5. What is the 75th percentile of city and state tax paid by residence code?

```{r}
quantile(tax$cityTax[tax$resCode=="PRINCIPAL RESIDENCE"],0.75,na.rm=TRUE)
quantile(tax$cityTax[tax$resCode=="NOT A PRINCIPAL RESIDENCE"],0.75,na.rm=TRUE)
quantile(tax$stateTax[tax$resCode=="PRINCIPAL RESIDENCE"],0.75,na.rm=TRUE)
quantile(tax$stateTax[tax$resCode=="NOT A PRINCIPAL RESIDENCE"],0.75,na.rm=TRUE)
```

Or we can use something like `tapply()`

```{r}
tapply(tax$cityTax, tax$resCode, quantile, 0.75,na.rm=TRUE)
tapply(tax$stateTax, tax$resCode, quantile, 0.75,na.rm=TRUE)
```

Or in `dplyr` (more to come next lecture)

```{r}
tax %>% group_by(resCode) %>% 
  summarise(quantile(stateTax, 0.75,na.rm=TRUE),
            quantile(cityTax, 0.75,na.rm=TRUE)) 
```

6. Subset the data to only retain those houses that are principal residences and describe the distribution of property taxes on these residences.

```{r}
quantile(tax$cityTax[tax$resCode == "PRINCIPAL RESIDENCE"], na.rm=TRUE)
quantile(tax$stateTax[tax$resCode == "PRINCIPAL RESIDENCE"], na.rm=TRUE)
```

7. Convert the 'lotSize' variable to a numeric square feet variable. Tips:
* Look at the data 
* Assume hyphens represent inches within square foot meassuremnts 
* Assume decimals within acreage measurements
* 1 acre = 43560 square feet
* Look at the data 

```{r}
tax$lotSize = str_trim(tax$lotSize) # trim to be safe
lot = tax$lotSize # for checking later
```

First lets take care of acres

```{r}
aIndex= c(grep("AC.*", tax$lotSize),
            grep(" %", tax$lotSize, fixed=TRUE))
head(aIndex)
head(lot[aIndex])

acre = tax$lotSize[aIndex] # temporary variable
## find and replace character strings
acre = gsub(" AC.*","",acre)
acre = gsub(" %","",acre)
table(!is.na(as.numeric(acre)))

head(acre[is.na(as.numeric(acre))],50)

## lets clean the rest
acre = gsub("-",".",acre,fixed=TRUE) # hyphen instead of decimal
head(acre[is.na(as.numeric(acre))])
table(!is.na(as.numeric(acre)))

acre = gsub("ACRES","", acre, fixed=TRUE)
head(acre[is.na(as.numeric(acre))])

# take care of individual mistakes
acre = gsub("O","0", acre, fixed=TRUE) # 0 vs O
acre = gsub("Q","", acre, fixed=TRUE) # Q, oops
acre = gsub(",.",".", acre, fixed=TRUE) # extra ,
acre = gsub(",","", acre, fixed=TRUE) # extra ,
acre = gsub("L","0", acre, fixed=TRUE) # leading L
acre[is.na(as.numeric(acre))]

acre2 = as.numeric(acre)*43560 
sum(is.na(acre2)) # all but 3
```

Now let's convert all of the square feet variables

```{r}
fIndex = grep("X", tax$lotSize)
ft = tax$lotSize[fIndex]

ft = gsub("&", "-", ft, fixed=TRUE)
ft = gsub("IMP ONLY ", "", ft, fixed=TRUE)
ft = gsub("`","1",ft,fixed=TRUE)

ft= sapply(str_split(ft, " "), first)

# wrapper for string split and sapply
#### ss = function(x, pattern, slot=1,...) sapply(strsplit(x,pattern,...), "[", slot)

width = sapply(str_split(ft,"X"), first)
length = sapply(str_split(ft,"X"), nth, 2) 

## width
widthFeet = as.numeric(sapply(str_split(width, "-"), first))
widthInch = as.numeric(sapply(str_split(width, "-"),nth,2))/12
widthInch[is.na(widthInch)] = 0 # when no inches present
totalWidth = widthFeet + widthInch # add together

# length
lengthFeet = as.numeric(sapply(str_split(length, "-"),first))
lengthInch = as.numeric(sapply(str_split(length, "-",2),nth,2))/12
lengthInch[is.na(lengthInch)] = 0 # when no inches present
totalLength = lengthFeet + lengthInch

# combine together for square feet
sqrtFt = totalWidth*totalLength 
ft[is.na(sqrtFt)] # what is left?
```

And now we combine everything together:

```{r}
tax$sqft = rep(NA)
tax$sqft[aIndex] = acre2
tax$sqft[fIndex] = sqrtFt
mean(!is.na(tax$sqft))

# already in square feet, easy!!
sIndex=c(grep("FT", tax$lotSize), 
         grep("S.*F.", tax$lotSize))
sf = tax$lotSize[sIndex] # subset temporary variable

sqft2 = sapply(str_split(sf,"( |SQ|SF)"),first)
sqft2 = as.numeric(gsub(",", "", sqft2)) # remove , and convert
tax$sqft[sIndex] = sqft2
table(is.na(tax$sqft)) 
## progress!

#what remains?
lot[is.na(tax$sqft)]
```

### B. "Baltimore Salary 2015" Dataset 

Download the "Salary 2015" Dataset from the website (via OpenBaltimore), which is located here:
http://sisbid.github.io/Module1/data//Baltimore_City_Employee_Salaries_FY2015.csv

8. Make an object called health.sal using the salaries data set,	with only agencies of those with "fire" (or any forms), if any, in the name

```{r}
Sal = read_csv("http://sisbid.github.io/Module1/data//Baltimore_City_Employee_Salaries_FY2015.csv")
```
9. Make a data set called trans which contains only agencies that contain "TRANS".

```{r}
trans = Sal[grep("trans", Sal$Agency,ignore.case=TRUE),]
```

10. What is/are the profession(s) of people who have "abra" in their name for Baltimore's Salaries?

```{r}
names(Sal)[1] = "Name"
table(Sal$JobTitle[grep("abra", Sal$Name, ignore.case=TRUE)])
```

11. What is the distribution of annual salaries look like? What is the IQR?

```{r}
Sal$AnnualSalary = as.numeric(gsub(pattern = "$", replacement="", 
                              Sal$AnnualSalary, fixed=TRUE))
quantile(Sal$AnnualSalary)
```

12. Convert HireDate to the `Date` class - plot Annual Salary vs Hire Date

```{r}
library(lubridate)
Sal$HireDate = mdy(Sal$HireDate)
plot(Sal$AnnualSalary ~ Sal$HireDate)
```

13. Create a smaller dataset that only includes the Police Department, Fire Department and Sheriff's Office. How many employees are in this new dataset?

```{r}
new = Sal[grep("fire|police|sheriff", Sal$JobTitle, ignore.case=TRUE),]
nrow(new)
length(table(new$Agency))
```
