---
title: "Data I/O Lab, Part 1"
output: html_document
---

```{r, include = TRUE}
library(readr)
library(readxl)
```

This lab uses different versions of the `iris` dataset.  For each question, click the `iris` file in the Files pane to open it in an RStudio window.

You can see more about this dataset by running `?iris` in the console.

1. Read in the `iris_q1.csv` dataset into the `iris_q1` R object. What delimiter separates columns?

```{r}
iris_q1 = read_csv("../data/iris/iris_q1.csv")
head(iris_q1)
```

2. Read in the `iris_q2.tsv` dataset into the `iris_q2` R object. What delimiter separates columns?

```{r}
iris_q2 = read_delim("../data/iris/iris_q2.tsv", delim = "\t")
head(iris_q2)
```

3. Read in the `iris_q3.txt` dataset into the `iris_q3` R object. What delimiter separates columns?

```{r}
iris_q3 = read_delim("../data/iris/iris_q3.txt", delim = " ")
head(iris_q3)
```

4. Read in the `iris_q4.csv` dataset into the `iris_q4` R object. What delimiter separates columns? [hint: file extension is ambiguous]

```{r}
iris_q4 = read_delim("../data/iris/iris_q4.csv", delim = ":")
head(iris_q4)
```

5. Read in the `iris_q5.csv` dataset into the `iris_q5` R object. Some entries read "-999" but are supposed to be missing (the value `NA` in the data).  

```{r}
iris_q5 = read_csv("../data/iris/iris_q5.csv", na = "-999")
head(iris_q5)
```

6. Read in sheet 2 of the `iris_q6.xlsx` dataset into the `iris_q6` R object. How many rows are in the dataset?

```{r}
iris_q6 = read_excel("../data/iris/iris_q6.xlsx", sheet= 2)
dim(iris_q6)
```
