---
title: "Data Summarization Lab Key"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data used

Circulator Lanes Dataset: the data is from https://data.baltimorecity.gov/Transportation/Charm-City-Circulator-Ridership/wwvu-583r

Available on: http://sisbid.github.io/Data-Wrangling/data/Charm_City_Circulator_Ridership.csv

```{r}
library(tidyverse)

circ <- read_csv("http://sisbid.github.io/Data-Wrangling/data/Charm_City_Circulator_Ridership.csv")
```

1. How many days are in the data set?  You can assume each observation/row is a different day

```{r q1}
nrow(circ)
dim(circ)
circ %>% 
  nrow()
```

2. What is the total (sum) number of boardings on the green bus (`greenBoardings`)?

```{r q2}
sum(circ$greenBoardings, na.rm = TRUE)

circ %>% pull(greenBoardings) %>% sum(na.rm = TRUE)
```

3. How many days are missing daily ridership?  Use `is.na()` and `sum()`. 
   
```{r q3}
daily <- circ %>% pull(daily)
sum(is.na(daily))

# Can also
circ %>% 
  count(is.na(daily))
```

4. Group the data by day of the week (`day`).

```{r q4}
circ %>% 
  group_by(day) %>% 
  summarise(mean = mean(daily, na.rm = TRUE),
            n = n())
```

## **Extra practice:**

5. Describe the distribution of `orangeBoardings`(use `quantile()`). The `quantile()` function works like `mean()`, `sum()`, etc.

```{r q6}
circ %>% 
  summarise(quantile(orangeBoardings, na.rm = TRUE))
```

6. Describe the distribution of `orangeBoardings`(use `quantile()`), but this time stratify by day of the week.

```{r q7}
circ %>% 
  group_by(day) %>% 
  summarise(quantile(orangeBoardings, na.rm = TRUE))
```
