---
title: "Data Summarization Lab"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data used

Circulator Lanes Dataset: the data is from https://data.baltimorecity.gov/Transportation/Charm-City-Circulator-Ridership/wwvu-583r

Available on: https://sisbid.github.io/Data-Wrangling/data/Charm_City_Circulator_Ridership.csv

```{r}
library(tidyverse)

circ <- read_csv("https://sisbid.github.io/Data-Wrangling/data/Charm_City_Circulator_Ridership.csv")
```

1. How many days are in the data set?  You can assume each observation/row is a different day (hint: get the number of rows).

```{r q1}

```

2. What is the total (sum) number of boardings on the green bus (`greenBoardings` column)?

```{r q2}

```

3. How many days are missing daily ridership (`daily` column)?  Use `is.na()` and `sum()`. 
   
```{r q3}

```

4. Group the data by day of the week (`day`). Next, find the mean daily ridership (`daily` column) and the sample size. (hint: use `group_by` and `summarize` functions)

```{r q4}

```

## **Extra practice:**

5. What is the median of `orangeBoardings`(use `median()`).

```{r q6}

```

6. Take the median of `orangeBoardings`(use `median()`), but this time stratify by day of the week.

```{r q7}

```
