---
title: "Data Cleaning"
author: "Andrew Jaffe (key)"
date: "July 11, 2017"
output: html_document
---

### A. "Real Property Taxes" Dataset 

Download the "real property taxes" dataset from the website (via OpenBaltimore), the data is located here (note you don\'t need to unzip it to read it into R):
http://sisbid.github.io/Module1/data/Real_Property_Taxes.csv.gz

1. Read the Property Tax data into R and call it the variable `tax`

```{r}
library(readr)
library(stringr)
suppressPackageStartupMessages(library(dplyr))
tax = read_csv("http://sisbid.github.io/Module1/data/Real_Property_Taxes.csv.gz")
# write_csv(tax, gzfile("table.csv.gz))
```

Note the `read_csv` function is treating the empty string `""` as missing whereas the `read.csv` function is not.

```{r}
tax2 = read.csv("../data/Real_Property_Taxes.csv.gz",as.is=TRUE)
table(complete.cases(tax))
table(complete.cases(tax2))
ind = which(!complete.cases(tax))[1:3]
tax[ind,]
tax2[ind,]
```

```{r}
tax3 = read.csv("../data/Real_Property_Taxes.csv.gz",as.is=TRUE,na.string="")
```

2. How many addresses pay property taxes? 

```{r}
nrow(tax)
dim(tax)
```

3. What is the total city and state tax paid?

```{r}
head(tax$CityTax)
tax$CityTax = tax$CityTax %>% 
  str_replace(fixed("$"), "") %>%
  str_replace(",", "") %>%
  as.numeric
tax$StateTax = tax$StateTax %>% 
  str_replace(fixed("$"), "") %>%
  str_replace(",", "") %>%
  as.numeric

sum(tax$CityTax, na.rm = TRUE)
sum(tax$CityTax, na.rm = TRUE)/1e6

sum(tax$StateTax, na.rm = TRUE)
sum(tax$StateTax, na.rm = TRUE)/1e6
```

Combined would be `r round((sum(tax$StateTax, na.rm = TRUE) + sum(tax$CityTax, na.rm = TRUE))/1e6)` milion dollars paid.


4. Convert the 'LotSize' variable to a numeric square feet variable. Tips:
* Look at the data 
* Assume hyphens represent inches within square foot meassuremnts 
* Assume decimals within acreage measurements
* 1 acre = 43560 square feet
* Look at the data 

```{r}
tax$LotSize = str_trim(tax$LotSize) # trim to be safe
lot = tax$LotSize # for checking later
```

First lets take care of acres

```{r}
aIndex= c(grep("AC.*", tax$LotSize),
            grep(" %", tax$LotSize, fixed=TRUE))
head(aIndex)
head(lot[aIndex])

acre = tax$LotSize[aIndex] # temporary variable
## find and replace character strings
acre = gsub(" AC.*","",acre)
acre = gsub(" %","",acre)
table(!is.na(as.numeric(acre)))

head(acre[is.na(as.numeric(acre))],50)

## lets clean the rest
acre = gsub("-",".",acre,fixed=TRUE) # hyphen instead of decimal
head(acre[is.na(as.numeric(acre))])
table(!is.na(as.numeric(acre)))

acre = gsub("ACRES","", acre, fixed=TRUE)
head(acre[is.na(as.numeric(acre))])

# take care of individual mistakes
acre = gsub("O","0", acre, fixed=TRUE) # 0 vs O
acre = gsub("Q","", acre, fixed=TRUE) # Q, oops
acre = gsub(",.",".", acre, fixed=TRUE) # extra ,
acre = gsub(",","", acre, fixed=TRUE) # extra ,
acre = gsub("L","0", acre, fixed=TRUE) # leading L
acre[is.na(as.numeric(acre))]

acre2 = as.numeric(acre)*43560 
sum(is.na(acre2)) # all but 3
```

Now let's convert all of the square feet variables

```{r}
fIndex = grep("X", tax$LotSize)
ft = tax$LotSize[fIndex]

ft = gsub("&", "-", ft, fixed=TRUE)
ft = gsub("IMP ONLY ", "", ft, fixed=TRUE)
ft = gsub("`","1",ft,fixed=TRUE)

ft= sapply(str_split(ft, " "), first)

# wrapper for string split and sapply
#### ss = function(x, pattern, slot=1,...) sapply(strsplit(x,pattern,...), "[", slot)

width = sapply(str_split(ft,"X"), first)
length = sapply(str_split(ft,"X"), nth, 2) 

## width
widthFeet = as.numeric(sapply(str_split(width, "-"), first))
widthInch = as.numeric(sapply(str_split(width, "-"),nth,2))/12
widthInch[is.na(widthInch)] = 0 # when no inches present
totalWidth = widthFeet + widthInch # add together

# length
lengthFeet = as.numeric(sapply(str_split(length, "-"),first))
lengthInch = as.numeric(sapply(str_split(length, "-",2),nth,2))/12
lengthInch[is.na(lengthInch)] = 0 # when no inches present
totalLength = lengthFeet + lengthInch

# combine together for square feet
sqrtFt = totalWidth*totalLength 
ft[is.na(sqrtFt)] # what is left?
```

And now we combine everything together:

```{r}
tax$sqft = rep(NA)
tax$sqft[aIndex] = acre2
tax$sqft[fIndex] = sqrtFt
mean(!is.na(tax$sqft))

# already in square feet, easy!!
sIndex=c(grep("FT", tax$LotSize), 
         grep("S.*F.", tax$LotSize))
sf = tax$LotSize[sIndex] # subset temporary variable

sqft2 = sapply(str_split(sf,"( |SQ|SF)"),first)
sqft2 = as.numeric(gsub(",", "", sqft2)) # remove , and convert
tax$sqft[sIndex] = sqft2
table(is.na(tax$sqft)) 
## progress!

#what remains?
lot[is.na(tax$sqft)]
```
