---
title: "Data Input/Output Lab Key"
date: "July 12, 2018"
output: html_document
---


```{r, intialSetup}
library(readr)
```

1. Download the UFO Dataset as a text file: http://sisbid.github.io/Module1/data/ufo/ufo_data_complete.txt and read in the text file as a new R object. Confirm the number of rows is the same as the CSV-based object

```{r, tabRead}
ufo2 = read_delim("http://sisbid.github.io/Module1/data/ufo/ufo_data_complete.txt",
                    delim = "\t")
```

2. Rename columns to ensure they have no spaces or non-alphanumeric characters. 

```{r, rename}
colnames(ufo2) = c("DateTime", "City", "State", "Country", "Shape",
                   "Duration_sec","Duration_hourMin","Comments",
                   "PostedDate", "Latitude","Longitude")
```

3. Use the [`lubridate` package](https://lubridate.tidyverse.org/) to change the class of the `$datetime` column to the date class.

```{r, date}
library(lubridate)
ufo2$PostedDate = mdy(ufo2$PostedDate)
ufo2$DateTime = mdy_hm(ufo2$DateTime)

```


4. Capitalize the states in the `$state` column

```{r, caps}
ufo2$State= toupper(ufo2$State)
head(ufo2$State)
```

5. Coerce the longitude and lattitude columns to the `numeric` class

```{r, numeric}
ufo2$Latitude = as.numeric(ufo2$Latitude)
ufo2$Longitude = as.numeric(ufo2$Longitude)
```

6. Write out this updated `data.frame` as a new CSV file

```{r, export, eval=FALSE}
write.csv(ufo2, file="new_ufo_data.csv",row.names=FALSE)
```

7. Download the Excel-formatted UFO Dataset: http://sisbid.github.io/Module1/data/ufo/ufo_data_complete.xlsx and import the dataset as a new R object. Confirm it has the same number of lines as the CSV- and Tab-delimited-based R objects

```{r, excel, eval=FALSE}
library(readxl)
ufo_excel = read_excel("../data/ufo/ufo_data_complete.xlsx")
ufo_excel = read_excel("../data/ufo/ufo_data_complete.xlsx",
                        col_type="text")

```
