---
title: "Data I/O Lab, Part 1"
output: html_document
---

In this lab you can use the interactive console to explore but please record your commands here.  Remember anything you type here can be "sent" to the console with Cmd-Enter (OS-X) or Cntr-Enter (Windows/Linux) (But only in side the ```{r}``` areas).

```{r, message = FALSE}
library(dplyr)
library(tidyverse)
library(naniar)
library(visdat)
```


1. Read in the the Charm City Circulator Dataset from "http://sisbid.github.io/Data-Wrangling/data/Charm_City_Circulator_Ridership.csv" using `read_csv`, call it `circ`
```{r}
circ <- read_csv("http://sisbid.github.io/Data-Wrangling/data/Charm_City_Circulator_Ridership.csv")
```

2. Print the number of rows of `circ`
```{r}
nrow(circ)
```

3. Visualize the missing data patterns of `circ` with `vis_miss`:
```{r}
vis_miss(circ)
```

4. Determine the percent missing in each column using `miss_var_summary`

```{r}
miss_var_summary(circ)
```

5. Determine the number missing in each column using `gg_miss_var`:
```{r}
gg_miss_var(circ)
```

6. Write out `circ` to a file called "Circulator.csv" using `write_csv`:
```{r}
write_csv(circ, "Circulator.csv")
```

7. Write out `circ` to a file called "Circulator.rds" using `write_rds`:

```{r}
write_rds(circ, "Circulator.rds")
```