---
title: "Functional Programming"
output: html_document
---

```{r, include = FALSE}
library(tidyverse)
```

Check out this purrr tutorial by Jenny Bryan: https://jennybc.github.io/purrr-tutorial/ls02_map-extraction-advanced.html

Additional content to read up on: 
https://jennybc.github.io/purrr-tutorial/ls01_map-name-position-shortcuts.html

0. Create a new dataset from iris called iris_lab. Use this for the lab.
```{r}
iris_lab <- iris
```

1. Working with the iris (automatically in r), use `mutate` and `across` to change the numeric variables to be rounded with the `round` function.
```{r}

```

2. Now do the same thing as question one, but this time set the number of digits for rounding to be 1.

```{r}

```

3. Create a function for adding 2 to a value and dividing by 6. Call the function `new_function` and use x as the input. Apply this function to the numeric columns of the iris data.

```{r}

```

3. Now create a function that includes x and y inputs. Call the function `div_function` and use it to divide x by y and then add 3 to the outcome. Hint start with `function(x, y){}`. fill in the curly brackets.
```{r}

```

4. Use `div_function` and `map2_dbl` to calculate values with x as `Sepal.Length` and y as `Sepal.Width` from the `iris` data.
```{r}
```


5. Use `if_all()` to check if there are iris flowers with sepal and petal widths and lengths (all the iris numeric variables) greater than 2 centimeters but less than g centimeters (which is the unit, if you type in ?iris you can learn more about the data). Hint - use `filter`. If any are, how many?
```{r}

```


6.  Use `modify_if` to convert numeric columns of `iris` to be character. Convert this to be a tibble and reassign `iris_lab`.
```{r}

```






