---
title: "Functional Programming"
output: html_document
---

```{r, include = FALSE}
library(tidyverse)
```

Check out this purrr tutorial by Jenny Bryan: https://jennybc.github.io/purrr-tutorial/ls02_map-extraction-advanced.html

Additional content to read up on: 
https://jennybc.github.io/purrr-tutorial/ls01_map-name-position-shortcuts.html

0. Create a new dataset from `iris` called `iris_lab`. Use this for the lab.

```{r}
iris_lab <- iris
```

1. Working with the `iris_lab` data, use `mutate` and `across` to change the numeric variables to be rounded with the `round` function. Take a look with `head()`- do this for the rest of the questions in which you create an output of data.

```{r}
iris_lab %>% 
  mutate(across(.cols = !Species, round)) %>%
  head()
# OR 
iris_lab %>% 
  mutate(across(.cols = where(is.numeric), round)) %>%
  head()
```

2. Now do the same thing as question one, but this time set the number of digits for rounding to be 1.

```{r}
iris_lab %>% 
  mutate(across(.cols = !Species, ~ round(.x, digits = 1))) %>%
  head()
```

3. Create a function for adding 2 to a value and dividing by 6. Call the function `new_function` and use x as the input. Apply this function to the numeric columns of the `iris_lab` data. Hint: start with `function(x){}`. Fill in the curly brackets.

```{r}
new_function <- function(x){(x+2)/6}
iris_lab %>% 
  mutate(across(.cols = !Species, ~ new_function(.x))) %>%
  head()
```

4. Now create a function that includes x and y inputs. Call the function `div_function` and use it to divide x by y and then add 3 to the outcome. Hint: start with `function(x, y){}`. Fill in the curly brackets.

```{r}
div_function <- function(x, y){(x/y)+ 3}
```

5. Use `if_all()` to check if there are iris flowers with sepal and petal widths and lengths (all the `iris_lab` numeric variables) greater than 2 centimeters but less than 6 centimeters (which is the unit, if you type in ?iris you can learn more about the data). Hint - use `filter`. If any are, how many?

```{r}
iris_lab %>%
filter(if_all(starts_with(c("Petal", "Sepal")), ~.x > 2 & .x < 6))  %>%
  head()
```

6.  the function `format()` with the argument `scientific = TRUE` can be used to format numbers in scientific notation. For example:

```{r}
format(10000, scientific = TRUE)
```

Use `modify_if` to convert *numeric* columns of `iris_lab` to be in scientific notation. 

```{r}
iris_lab <- iris_lab %>%
  modify_if(is.numeric, ~ format(.x, scientific = TRUE))

head(iris_lab)
```
