---
title: "Google Sheets"
output: html_document
---

```{r, include=FALSE}
library(tidyverse)
library(httr)
library(jsonlite)
library(googlesheets4)
```

This is a lab to practice with Google Sheets and the `googlesheets` package.

1. We are going to use a sheet from previous years: https://docs.google.com/spreadsheets/d/1KIRtcPVn58R3_qr97WNtcOJiY4AaytHzGDzLW_3_R1s/edit?usp=sharing

```{r}
sheet_url = "https://docs.google.com/spreadsheets/d/1KIRtcPVn58R3_qr97WNtcOJiY4AaytHzGDzLW_3_R1s/edit?usp=sharing"
```


2. Use the `googlesheets4::read_sheet` function to read in the data like we discussed in class, call this object `x`.
```{r}

```

2. Using `sheet_names` function to get the sheet names from the `sheet_url`.
```{r}

```


4. Try reading in just the first two columns. See the `range` argument and look at the arguments for `cell_rows`, `cell_limits` and `cell_cols`.
```{r}

```

5. How many students are in the class (or at least have added their information to the Sheet)?
```{r}

```
6. Create a logical variable for whether students have more than 2 years of R experience
```{r}

```
7. Convert the logical variable from \#6 into a factor with labels "Skilled" and "Learning" 
```{r}

```
8. Create a new spreadsheet that is private only to you.
```{r}

```
9. Use `gs_auth` to authorize your computer to access that sheet. 
```{r}

```
10. Read the data from your private spreadsheet. 
```{r}

