---
title: "json lab"
output: html_document
---

```{r, include=FALSE}
library(tidyverse)
library(httr)
library(jsonlite)
```

See GitHub Documentation at https://docs.github.com/en/rest/reference/repos

1. Use Github API to get all of Hadley Wickham's repo data (username = "hadley"), see: "https://api.github.com/users/XXXX/repos".  Use `httr:GET` and assign that output to `response` and then `httr::content(response)`.  Also run `httr:status_code(response)` to get the status code

```{r}
response = httr::GET("https://api.github.com/users/hadley/repos")
httr::content(response)
httr::status_code(response)
```

2.  Try `jsonlite::fromJSON(httr::content(response, as = "text", flatten = TRUE))`
```{r, error = TRUE}
out = jsonlite::fromJSON(httr::content(response, as = "text", flatten = TRUE))
class(out)
head(out)
```

3. How many stars does he have on each repository? (it's the `stargazers_count` element)
```{r}
out$stargazers_count
```

4. Create a new data frame consisting of just the "owner" column of the resulting data set called `owner_data`

```{r}
owner_data = out$owner
```