---
title: "Manipulating Data in R Lab"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(readr)
library(dplyr)
library(ggplot2)
library(tidyr)
library(stringr)
```

# Part 1
1.  Read in the Wide Bike lanes data: https://sisbid.github.io/Module1/labs/Bike_Lanes_Wide.csv .  Name the data set `wide`.

```{r}
wide = read_csv("https://sisbid.github.io/Module1/labs/Bike_Lanes_Wide.csv")
```

2. Reshape `wide` using `gather.`  Call this data `long.`  Make the `key`
`lanetype`, and the value `the_length.`  Make sure we gather all columns **but**
`name`, using `-name`. Note the `NA`s here:

```{r}
long = wide %>% 
  gather(key = "lanetype", value = "the_length", -name)
head(long)
```



# Part 2

This exercise is motivated by https://anythingbutrbitrary.blogspot.com/2012/08/manipulating-data-frames-using-sqldf.html

3. read in the roads and crashes CSVs.  Call them `crash` and `road`
`crash`: https://sisbid.github.io/Module1/labs//crashes.csv
`road`: https://sisbid.github.io/Module1/labs//roads.csv

```{r}
crash = read_csv("https://sisbid.github.io/Module1/labs//crashes.csv")
road = read_csv("https://sisbid.github.io/Module1/labs//roads.csv")
head(crash)
head(road)
```

4.  Replace (using `str_replace`) any hyphens (`-`) with a space in `crash$Road`. Call this data `crash2`. Table the `Road` variable.  
```{r}
crash2 = crash %>% mutate(Road = str_replace(Road, "-", " "))
table(crash2$Road)
```

5. How many observations are in each dataset?
```{r}
dim(crash)
dim(road)
```

6.  Separate the Road column (using `separate`) into (`type` and `number`) in
`crash2.`  Reassign this to `crash2.` Table `crash2$type`
```{r}
crash2 = separate(crash2, col = "Road", into = c("type", "number"))
table( crash2$type)
```

Create a new variable calling it road_hyphen using the `unite` function.  Unite the type and number columns using a hyphen (`-`) and then `table` `road_hyphen`
```{r}
crash2 = unite(crash2, col = "road_hyphen", type, number ,sep = "-")
table( crash2$road_hyphen)
```

7. Which and how many years were data collected?
```{r}
unique(crash$Year)
length(unique(crash$Year))
```

# Part 3
8.  Read in the Bike lanes data: https://sisbid.github.io/Module1/labs/Bike_Lanes.csv.
Name the data set `bike`, 
```{r}
bike = read_csv("https://sisbid.github.io/Module1/labs/Bike_Lanes.csv")
```

9.  Keep rows where the record is not missing `type` and not missing `name` and re-assign the output to `bike`.

```{r}
bike = filter(bike, !is.na(type) & !is.na(name))
```

10.  Summarize and group the data by grouping `name` and `type`  and take the `sum` of the length (reassign `length` variable).  Call this data set `sub`

```{r}
sub = bike %>% 
  group_by(name, type) %>% 
  summarize(length = sum(length))
```

11. Reshape `sub` using `spread.`  Spread the data where the `key` is type and we want the value in the new columns to be `length` - the bike lane length.  
Call this `wide`.  

```{r}
wide = spread(sub, key = type, value = length)
```

Look at the column names of `wide` - what are they? (they also have spaces)

# Part 4

12. Join data to retain only complete data, (using an inner join) e.g. those observations with road lengths and districts.  Merge without using `by` argument, then merge using `by = "Road"`. call the output `merged.`  How many observations are there?

```{r}
merged = inner_join(crash, road)
merged = inner_join(crash, road, by = "Road")
dim(merged)
```

13. Join data using a `full_join.`  Call the output `full.`   How many observations are there?

```{r}
full = full_join(crash, road)
nrow(full)
```

14. Do a left join of the `road` and `crash`.  ORDER matters here!  How many observations are there?

```{r}
left = left_join(road, crash)
nrow(left)
```

15. Repeat above with a `right_join` with the same order of the arguments. How many observations are there?

```{r}
right = right_join(road, crash)
nrow(right)
```
