---
title: "Merging and Joining Lab Key"
author: "Data Wrangling in R"
output: html_document
---

## Part 1

Read in the data and use functions of your choice to preview it.

```{r, message = FALSE}
library(tidyverse)

crash <- read_csv("https://sisbid.github.io/Data-Wrangling/labs/crashes.csv")
road <- read_csv("https://sisbid.github.io/Data-Wrangling/labs/roads.csv")
head(crash)
head(road)
```

1. Join data to retain only complete data, (using an inner join) e.g. those observations with road lengths and districts.  Merge without using `by` argument, then merge using `by = "Road"`. call the output `merged.`  How many observations are there?

```{r}
merged <- inner_join(crash, road)
merged <- inner_join(crash, road, by = "Road")
nrow(merged)
```

2. Join data using a `full_join.`  Call the output `full.`   How many observations are there?

```{r}
full <- full_join(crash, road)
nrow(full)
```

3. Do a left join of the `road` and `crash`.  ORDER matters here!  How many observations are there?

```{r}
left <- left_join(road, crash)
nrow(left)
```

4. Repeat above with a `right_join` with the same order of the arguments. How many observations are there?

```{r}
right <- right_join(road, crash)
nrow(right)
```

5. What `road` data is missing from `crash`?

```{r}
roads1 <- road %>% pull(Road)
roads2 <- crash %>% pull(Road)
setdiff(roads1, roads2) # This value is in `road` but not `crash`

# Could also search for NAs created by the join
full %>% filter(is.na(N_Crashes))

anti_join(road, crash)
```

5. What `crash` data is missing from `road``?

```{r}
roads1 <- road %>% pull(Road)
roads2 <- crash %>% pull(Road)
setdiff(roads2, roads1) # These values are in `crash` but not `road`

# Could also search for NAs created by the join. Would be good to summarize with `count`
full %>% filter(is.na(District)) %>% count(Road)

anti_join(crash, road)
```
