---
title: "Merging Lab"
author: "Data Wrangling in R"
output: html_document
---


(This exercise is motivated by https://anythingbutrbitrary.blogspot.com/2012/08/manipulating-data-frames-using-sqldf.html)

```{r, message = FALSE}
library(tidyverse)

crash = read_csv("https://sisbid.github.io/Data-Wrangling/labs/crashes.csv")
road = read_csv("https://sisbid.github.io/Data-Wrangling/labs/roads.csv")
head(crash)
head(road)
```

1. Join data to retain only complete data, (using an inner join) e.g. those observations with road lengths and districts.  Merge without using `by` argument, then merge using `by = "Road"`. call the output `merged.`  How many observations are there?

```{r}


```

2. Join data using a `full_join.`  Call the output `full.`   How many observations are there?

```{r}


```

3. Do a left join of the `road` and `crash`.  ORDER matters here!  How many observations are there?

```{r}


```

4. Repeat above with a `right_join` with the same order of the arguments. How many observations are there?

```{r}


```
