---
title: "Merging Lab Key"
author: "Data Wrangling in R"
output: html_document
---

This exercise is motivated by https://anythingbutrbitrary.blogspot.com/2012/08/manipulating-data-frames-using-sqldf.html

1. read in the roads and crashes CSVs.  Call them `crash` and `road`
`crash`: https://sisbid.github.io/Module1/labs//crashes.csv
`road`: https://sisbid.github.io/Module1/labs//roads.csv

```{r}
crash = read_csv("https://sisbid.github.io/Module1/labs/crashes.csv")
road = read_csv("https://sisbid.github.io/Module1/labs/roads.csv")
head(crash)
head(road)
```


2. Join data to retain only complete data, (using an inner join) e.g. those observations with road lengths and districts.  Merge without using `by` argument, then merge using `by = "Road"`. call the output `merged.`  How many observations are there?

```{r}
merged = inner_join(crash, road)
merged = inner_join(crash, road, by = "Road")
dim(merged)
```

3. Join data using a `full_join.`  Call the output `full.`   How many observations are there?

```{r}
full = full_join(crash, road)
nrow(full)
```

4. Do a left join of the `road` and `crash`.  ORDER matters here!  How many observations are there?

```{r}
left = left_join(road, crash)
nrow(left)
```

5. Repeat above with a `right_join` with the same order of the arguments. How many observations are there?

```{r}
right = right_join(road, crash)
nrow(right)
```
