---
title: "Merging Lab"
author: "Data Wrangling in R"
output: html_document
---

## Part 1

(This exercise is motivated by https://anythingbutrbitrary.blogspot.com/2012/08/manipulating-data-frames-using-sqldf.html)

```{r, message = FALSE}
library(tidyverse)

crash = read_csv("https://sisbid.github.io/Data-Wrangling/labs/crashes.csv")
road = read_csv("https://sisbid.github.io/Data-Wrangling/labs/roads.csv")
head(crash)
head(road)
```

1. Join data to retain only complete data, (using an inner join) e.g. those observations with road lengths and districts.  Merge without using `by` argument, then merge using `by = "Road"`. call the output `merged.`  How many observations are there?

```{r}


```

2. Join data using a `full_join.`  Call the output `full.`   How many observations are there?

```{r}


```

3. Do a left join of the `road` and `crash`.  ORDER matters here!  How many observations are there?

```{r}


```

4. Repeat above with a `right_join` with the same order of the arguments. How many observations are there?

```{r}


```

## Part 2

5. Which highways do not have road data? Summarize by the total count of `N_Crashes`.

```{r}


```

**Bonus**

6. You have an intern who has been pouring over the raw data and found a few mistakes in the `crash` dataset. They have made a spreadsheet for you containing only the corrected entries. Modify the original tibble with the following:
- A column (`Corrected`) indicating if a particular entry has a corresponding correction in `corrections`.
- If the row has a correction, take the corrected value
- Keep the original columns (Year, Road, N_Crashes, Volume) plus the column indicating whether the data is corrected or not (`Corrected`).

*hint: take a look at the two datasets - are you sure they're joining correctly?*

```{r}
corrections = read_csv("crashes_corrections.csv")

```
