---
title: "Merging Lab Key"
author: "Data Wrangling in R"
output: html_document
---

## Part 1

Read in the data and use functions of your choice to preview it.

```{r, message = FALSE}
library(tidyverse)

crash <- read_csv("https://sisbid.github.io/Data-Wrangling/labs/crashes.csv")
road <- read_csv("https://sisbid.github.io/Data-Wrangling/labs/roads.csv")
head(crash)
head(road)
```

1. Join data to retain only complete data, (using an inner join) e.g. those observations with road lengths and districts.  Merge without using `by` argument, then merge using `by = "Road"`. call the output `merged.`  How many observations are there?

```{r}
merged <- inner_join(crash, road)
merged <- inner_join(crash, road, by = "Road")
nrow(merged)
```

2. Join data using a `full_join.`  Call the output `full.`   How many observations are there?

```{r}
full <- full_join(crash, road)
nrow(full)
```

3. Do a left join of the `road` and `crash`.  ORDER matters here!  How many observations are there?

```{r}
left <- left_join(road, crash)
nrow(left)
```

4. Repeat above with a `right_join` with the same order of the arguments. How many observations are there?

```{r}
right <- right_join(road, crash)
nrow(right)
```

## **Bonus Practice**

5. Which highways do not have road data? Do this in a "tidy" format. Summarize by the total count of `N_Crashes` per highway. Hint: Use `anti_join()` and `group_by()`.

```{r}
no_dat <- crash %>% 
  anti_join(road) %>% 
  group_by(Road) %>%
  summarize(total_crashes = sum(N_Crashes))
no_dat
```

6. You have an intern who has been pouring over the raw data and found a few mistakes in the `N_Crashes` column of the `crash` dataset. They have made a spreadsheet for you containing only the corrected entries. Modify the original tibble with the following:
- A column (`Corrected`) indicating if a particular entry has a corresponding correction in `corrections`.
- If the row has a correction, take the corrected value
- Keep the original columns (Year, Road, N_Crashes, Volume) plus the column indicating whether the data is corrected or not (`Corrected`).

*hint: take a look at the two datasets - are you sure they're joining correctly?*

```{r}
corrections <- read_csv("https://sisbid.github.io/Data-Wrangling/labs/crashes_corrections.csv")

corrections$ROAD <- corrections$ROAD %>% str_replace("INTERSTATE", "Interstate")

crash %>% 
  full_join(corrections, by = c("Year" = "YEAR", "Road" = "ROAD"), suffix = c(".orig", ".checked")) %>%
  mutate(Corrected = ifelse(is.na(N_Crashes.checked), 0, 1)) %>%
  mutate(N_Crashes = ifelse(Corrected == 0, N_Crashes.orig, N_Crashes.checked)) %>%
  select(Year, Road, N_Crashes, Volume.orig, Corrected) %>%
  rename(Volume = Volume.orig)
```
