---
title: "R markdown lab"
output: html_document
---

1. Try compiling this document using the "Knit HTML" button. What files are produced?
2. Edit the output to be "word_document" and recompile. What files are produced?
3. What does the label "chunk1" do in the following code chunk?


```{r chunk1}
x = rnorm(100)
plot(x)
```

### This is a primary header. Add a secondary header with ## on the line below and recompile.


4. Change fig.height to something larger/smaller and see what happens

```{r chunk2, fig.height=5}
x = rnorm(100)
plot(x)
```

5. Add a chunk option of fig.align="center" to center this plot

```{r chunk3, fig.align="center" }
x = rnorm(100)
plot(x)
```

6. Add a chunk option of echo=FALSE and see what happens

```{r chunk4, echo=FALSE }
x = rnorm(100)
plot(x)
```


7. Check out the docs at: http://rmarkdown.rstudio.com/ and look under formats. See if you can create and compile an R markdown presentation. 

8. Install the `devtools package` by copy pasting install.packages("devtools") into your console. Then load the package using `libray(devtools)` and the command `session_info()` in this chunk and look at the output.

```{r session_info}
#install.packages("devtools")
library(devtools)
session_info()
```






