---
title: "R markdown lab"
author: "Jeff Leek"
date: "July 3, 2015"
output: html_document
---

1. Try compiling this document using the "Knit HTML" button. What files are produced?
2. Edit the output to be "pdf_document" and recompile. What files are produced?
3. What does the label "chunk1" do in the following code chunk?

```{r chunk1}
x = rnorm(100)
plot(x)
```

# This is a primary header. Add a secondary header with ## on the line below and recompile.


5. Change fig.height to something larger/smaller and see what happens

```{r chunk2, fig.height=5}
x = rnorm(100)
plot(x)
```

6. Add a chunk option of fig.align="center" to center this plot

```{r chunk3, }
x = rnorm(100)
plot(x)
```

6. Add a chunk option of echo=FALSE and see what happens

```{r chunk4, }
x = rnorm(100)
plot(x)
```

7. Uncomment this code and recompile the document. Then add a chunk command cache=TRUE and recompile. What files are created? Try compiling one more time. 

```{r longtime, cache=TRUE}
Sys.sleep(10)
```


8. Add the command devtools::session_info() to this chunk and look at the output

```{r session_info}

```



