---
title: "R markdown lab"
output: html_document
---

1. Try compiling this document using the "Knit HTML" button. What files are produced?
2. Edit the output to be "word_document" and recompile. What files are produced?
3. Try running the chunk below by pressing the green play button on the far right. What happens?

```{r chunk1}
x = rnorm(100)
plot(x)
```

### This is a primary header. Add a secondary header with ## on the line below and recompile.


4. Try pressing the run previous chunk button - the button that is to the left of the play button. What happens?

```{r chunk2}
x = rnorm(100)
plot(x)
```

5. Add a chunk option of fig.width = 10 to this plot.

```{r chunk3, fig.width = 10}
x = rnorm(100)
plot(x)
```

6. Add a chunk option of echo=FALSE and see what happens.

```{r chunk4, echo=FALSE }
x = rnorm(100)
plot(x)
```


7. Check out the docs at: http://rmarkdown.rstudio.com/ and look under formats. 

8. Install the `devtools package` by copy pasting install.packages("devtools") into your console. Then load the package using `libray(devtools)` and the command `session_info()` in this chunk and look at the output.

```{r session_info}
#install.packages("devtools")
library(devtools)
session_info()
```






