---
title: "R markdown lab"
author: "Jeff Leek"
date: "July 3, 2015"
output: html_document
---

1. Try compiling this document using the "Knit HTML" button. What files are produced?
2. Edit the output to be "pdf_document" and recompile. What files are produced?
3. Edit the output to be "word_document" and recompile. What files are produced?
4. What does the label "chunk1" do in the following code chunk?


```{r chunk1}
x = rnorm(100)
plot(x)
```

### This is a primary header. Add a secondary header with ## on the line below and recompile.


5. Change fig.height to something larger/smaller and see what happens

```{r chunk2, fig.height=5}
x = rnorm(100)
plot(x)
```

6. Add a chunk option of fig.align="center" to center this plot

```{r chunk3, fig.align="center" }
x = rnorm(100)
plot(x)
```

7. Add a chunk option of echo=FALSE and see what happens

```{r chunk4, echo=FALSE }
x = rnorm(100)
plot(x)
```

8. Uncomment this code and recompile the document. Then add a chunk command cache=TRUE and recompile. What files are created? Try compiling one more time. 

```{r longtime, cache=TRUE}
Sys.sleep(10)
```

9. Check out the docs at: http://rmarkdown.rstudio.com/ and look under formats. See if you can create and compile an R markdown presentation. 

10. Add the command devtools::session_info() to this chunk and look at the output

```{r session_info}
devtools::session_info()
```




