---
title: "Subsetting Data Lab"
author: "Data Wrangling in R"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In this lab you can use the interactive console to explore but please record your commands here.  

Remember anything you type here can be "sent" to the console with: Cmd-Enter (OS-X) or Ctrl-Enter (Windows/Linux).

### Part 1

1. Check to see if you have the `mtcars` dataset (which is included in the `datasets` package)

```{r, import}
library(tibble)
data(mtcars)
mtcars = rownames_to_column(mtcars, var = "car")
```

2. What class is `mtcars`?

```{r, class}
class(mtcars)
```

3. How many observations (rows) and variables (columns) are in the `mtcars` dataset?

```{r, nrow}
library(dplyr)
nrow(mtcars)
mtcars %>% nrow

mtcars %>% ncol
ncol(mtcars)
```
4. Select the subset of rows that get more than 20 miles per gallon (mpg) of fuel efficiency - how many are there?

```{r, filter}
mtcars %>% filter(mpg > 20) 
mtcars %>% filter(mpg > 20) %>% nrow
sum(mtcars$mpg > 20)
```

5. Select the subset of rows that get less than 16 miles per gallon (mpg) of fuel efficiency and have more than 100 horsepower (hp) - how many are there?

```{r, filter2}
mtcars %>% filter(mpg < 16, hp > 100) 
mtcars %>% filter(mpg < 16, hp > 100) %>% nrow
sum(mtcars$mpg < 16 & mtcars$hp > 100)
```

mtcars %>% filter(mpg < 16, hp > 100) 


### Part 2

6. Create a subset of the data that only contains the columns: `wt`, `qsec`, and `hp` for only the cars that have 8 cylinder, and assign this object to `carsSub` - what are the dimensions of this dataset?

```{r, subset}
carsSub = mtcars %>% filter(cyl == 8) %>% select(car, wt, qsec, hp)
carsSub %>% dim
dim(carsSub)
```

7. Convert the column names of `carsSub` to all upper case

```{r, convert}
colnames(carsSub) = toupper(colnames(carsSub))
```

8. Re-order the rows of `carsSub` by weight in increasing order

```{r, reorder}
carsSub %>% arrange(WT)
```

9. Copy mtcars into a variable called `cars` and rename the `mpg` column in `cars` to `MPG`

```{r, rename}
cars = mtcars
cars = rename(cars, MPG=mpg)
```

10. Subset the columns from mtcars that end in "p" and call the new data frame `pvars`

```{r}
pvars = mtcars %>% select(car, ends_with("p"))
pvars
```