---
title: "subsetting Lab"
author: "Andrew Jaffe"
date: "July 11, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
```

In this lab you can use the interactive console to explore but please record your commands here.  

Remember anything you type here can be "sent" to the console with: Cmd-Enter (OS-X) or Ctrl-Enter (Windows/Linux).

1. Check to see if you have the `mtcars` dataset (which is included in the `datasets` package)

```{r}
dim(mtcars)
head(mtcars)
## rename rownames as first column
mtcars$name = rownames(mtcars)
mtcars = select(mtcars, name, everything())
rownames(mtcars) = NULL
# mtcars = datasets::mtcars # in case you mess up
```

2. What class is `mtcars`?

```{r}
class(mtcars)
```

The class of `mtcars` is `r class(mtcars)`.

3. How many observations (rows) and variables (columns) are in the `mtcars` dataset?

```{r}
dim(mtcars)
nrow(mtcars)
ncol(mtcars)
str(mtcars)
```

There are `r nrow(mtcars)` cars.

4. Select the subset of rows that get more than 20 miles per gallon (mpg) of fuel efficiency - how many are there?

```{r}
nrow(filter(mtcars, mpg > 20))
mtcars %>% filter(mpg > 20) %>% nrow
nrow(mtcars[ mtcars$mpg > 20, ])
sum(mtcars$mpg > 20)
length(which(mtcars$mpg > 20))
```

5. Select the subset of rows that get less than 16 miles per gallon (mpg) of fuel efficiency and have more than 100 horsepower (hp) - how many are there?

```{r}
mtcars %>% filter(mpg < 16 & hp > 100) %>% nrow
```

6. Create a subset of the data that only contains the columns: wt, qsec, and hp for only the cars that have 8 cylinder, and assign this object to `carsSub` - what are the dimensions of this dataset?

```{r}
carsSub = mtcars %>% 
  filter(cyl == 8) %>% 
  select(wt, qsec, hp)
carsSub = mtcars[
    mtcars$cyl == 8, c("wt", "qsec", "hp")]
```

7. Convert the column names of `carsSub` to all upper case

```{r}
cn = colnames(carsSub)
cn = toupper(cn)
colnames(carsSub) = cn
colnames(carsSub) = colnames(carsSub) %>% toupper

colnames(carsSub) = toupper(colnames(carsSub))
```

8. Re-order the rows of `carsSub` by weight in increasing order

```{r}
arrange(carsSub, WT)
carsSub[ order(carsSub$WT),]
```

9. Copy mtcars into a variable called `cars` and rename the `mpg` column in `cars` to `MPG`

```{r}
cars = mtcars
# cars$MPG = cars$mpg
cars = rename(cars, MPG = mpg)
```

10. Subset the columns from mtcars that end in "p" and call the new data frame `pvars`

```{r}
pvars = select(mtcars, ends_with("p"))
head(pvars)
```