---
title: "tidyr lab"
output: html_document
---

1. Load the `readr` package and read in the data on hospitalizations using
`seps <- read_csv("http://www.mm-c.me/mdsi/hospitals93to98.csv")`

```{r}
seps <- read_csv("http://www.mm-c.me/mdsi/hospitals93to98.csv")
```


2. Look at the data using `View`

```{r}
View(seps)
```


3. Gather the years into a single column

```{r}
seps %>% 
  gather(year,value,FY1993:FY1998)
```


4. Now use spread to create two columns out of the `Field` variable one for `PatientDays` and one for `Separations`. 

```{r}
seps %>%
  gather(year,value,FY1993:FY1998) %>%
  spread(Field,value)
```


5. Separate the `Year` variable into a column that shows the abbreviation and a column that shows the year. 

```{r}
seps %>%
  gather(year,value,FY1993:FY1998) %>%
  spread(Field,value) %>%
  separate(col=year,into = c("abb","year"),sep=2)
```

6. Now reunite these two columns, but now include a hyphen between the abbreviation and the year. 

```{r}
seps %>%
  gather(year,value,FY1993:FY1998) %>%
  spread(Field,value) %>%
  separate(col=year,into = c("abb","year"),sep=2) %>%
  unite(newyear,abb,year,sep="-")
```


This lab was adopted from this nice tutorial on tidyr: https://rstudio-pubs-static.s3.amazonaws.com/282405_e280f5f0073544d7be417cde893d78d0.html
