---
title: "Web + API Key"
output: html_document
editor_options: 
  chunk_output_type: console
---

<!-- rdom::rdom(url = "http://bioconductor.org/packages/release/BiocViews.html#___Software", filename = "data/bioc-software.html") -->
### Web scraping 

Load the html file and extract the data table
```{r}
library(rvest)
# htmlfile = read_html("../data/R-wiki.html")
# nds = html_nodes(htmlfile, xpath='//*[@class="wikitable"]')
htmlfile = read_html("../data/bioc-software.html")
nds = html_nodes(htmlfile, xpath='.//*[@class="dataTable"]')
dat = html_table(nds)
dat = as.data.frame(dat[[2]])
head(dat)
```

Paste the last column and make the word cloud

```{r}
library(wordcloud)
txt = paste(dat[,3],collapse=" ")
wordcloud(txt,max.words=50)
```

### APIs

You can modify the url like so to get _P53_ info:

```{r, error = TRUE}
library(jsonlite)
p53_dat = fromJSON("http://rest.ensembl.org/lookup/id/ENSG00000157764?expand=1;content-type=application/json")
```

To look up the BRAF gene information we would. 

1. Look at the docs under GET species lookup http://rest.ensembl.org/documentation/info/symbol_lookup. 
2. We'd look at the Example Request: `/lookup/symbol/homo_sapiens/BRCA2?content-type=application/json;expand=1` and modify it with our gene like so: 
`/lookup/symbol/homo_sapiens/BRAF?content-type=application/json;expand=1`
3. We'd create the url using the base url like this: `http://rest.ensembl.org//lookup/symbol/homo_sapiens/BRAF?content-type=application/json;expand=1`
4. We'd request the information with `fromJSON`
```{r, error = TRUE}
braf_info = fromJSON("http://rest.ensembl.org//lookup/symbol/homo_sapiens/BRAF?content-type=application/json;expand=1")
braf_info$id
```

